---
title: "dfdiffs::compareDataDemo"
author: "Martin Frigaard"
output:
  html_document: 
    highlight: espresso
    theme: darkly
    toc: true
    toc_float: yes
    df_print: paged
    code_folding: hide
    
runtime: shiny
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(fs)
library(reactable)
library(rmarkdown)
library(shiny)
library(shinythemes)
library(bs4Dash)
# Global options
knitr::opts_chunk$set(
  # cache = FALSE # cache data
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "tiny", # size of the text
  fig.path = "images", # location of figure files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
# reactable.theme ----
options(reactable.theme = reactableTheme(
          color = "hsl(235, 9%, 87%)",
          backgroundColor = "hsl(208,100%, 17%)",
          borderColor = "hsl(235, 9%, 22%)",
          stripedColor = "hsl(235, 12%, 22%)",
          highlightColor = "hsl(235, 12%, 24%)",
          inputStyle = list(backgroundColor = "hsl(235, 9%, 25%)"),
          selectStyle = list(backgroundColor = "hsl(235, 9%, 25%)"),
          pageButtonHoverStyle = list(backgroundColor = "hsl(235, 9%, 25%)"),
          pageButtonActiveStyle = list(backgroundColor = "hsl(235, 9%, 28%)")))
```

# The Goal 

The goal of `compareDataDemo` is to create a demo module for comparing two tables to get 1) new data, 2) deleted data, and 3) changed data in the `prodfdiffs` app. 

The application is broken into four modules: `upload`, `select`, `compare`, and `tables`. 

# Upload

The two functions for the upload modules are loaded below: 

## UI

```{r uploadDataUI}
source("uploadDataUI.R")
```

## Server

```{r uploadDataServer}
source("uploadDataServer.R")
```

These two files are not shown because they are displayed and covered in detail in previous documentation. The `uploadData` module returns a list of reactive datasets to be used in the `selectData` module. 

# Select 

The `uploadData` module 'passes' the data to the `selectData` module, which displays the data for the user to review and, if necessary, reduce. The `uploadData` module can import up to four data files, but the `selectData`  trims that down to two datasets (and the selected columns). 

We do this by including only two boxes in the select tab, and giving the list of columns from each uploaded dataset. 

## UI

```{r selectDataUI}
source("selectDataUI.R")
```

## Server 

```{r selectDataServer}
source("selectDataServer.R")
```

# Comparisons 

We're going to import two datasets to simulate the application's compare functions. We'll be using the Master dataset from the [Lahman baseball database](https://github.com/cdalzell/Lahman):

Our 'previous' or 'old' dataset is the 2015 Master dataset. 

```{r master15}
master15 <- readxl::read_excel("data/xlsx/lahman_compare.xlsx", 
                                sheet = "master-2015")
glimpse(master15)
```

Our 'current' or 'new' dataset is the 2020 People dataset 

```{r master20}
master20 <- readxl::read_excel("data/xlsx/lahman_compare.xlsx", 
                                sheet = "master-2020")
glimpse(master20)
```

## New Records 

We'll test the `create_new_data()` function on the 2015 vs. 2020 Master data. 

```{r create_new_data}
# run new data function
new_master_15_20 <- create_new_data(newdf = master20, 
  olddf = master15, 
  by = "playerID",
  cols = c("playerID", "nameGiven", "height", "bats", "throws"))
reactable(new_master_15_20,
  defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE)
```

In order to check these data, we can test the unique `playerID`s in and see if they're in the `master15` dataset. 

```{r new_playerID}
# unique ids
new_playerID <- unique(new_master_15_20$playerID)
# filter new ids (should be zero)
glimpse(filter(master15, 
  playerID %in% new_playerID))
```

And we can see **none** of these `playerID`s are in the previous dataset.

## Deleted Records 

We're going to check the deleted records using the `create_deleted_data()` function. 

```{r create_deleted_data}
deleted_master_15_20 <- create_deleted_data(
  newdf = master20, 
  olddf = master15, 
  by = "playerID",
  cols = c("playerID", "nameGiven", "height", "bats", "throws"))
reactable(deleted_master_15_20,
    defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE)
```

This tells us there were 625 deleted records between 2015 and 2020 in the master dataset. 

## Modified Records 

Below we test our modified function, `create_modified_data()`, which returns a list of two datasets. 

```{r create_modified_data}
modified_master_15_20 <- create_modified_data(
  newdf = master20, 
  olddf = master15, 
  by = "playerID",
  cols = c("playerID", "nameGiven", "height", "bats", "throws"))
names(modified_master_15_20)
```

The `diffs_byvar` table contains each variable modified (`Variable name`), the count of changes ( `Modified values`) and the count of missing (`Missing Values`).

```{r modified_master_15_20-diffs_byvar}
reactable(data = modified_master_15_20$diffs_byvar,
  defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE)
```

The diffs table contains the modified variable (`Variable Name`), the joining column (`playerID`), and the modified values (`Current Value` & `Previous Value`). 

```{r modified_master_15_20-diffs}
reactable(data = modified_master_15_20$diffs,
  defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE)
```

All of these tables will be presented in the `compareDataUI()`


# Compare 

The three functions for comparing any two datasets in the `dfdiffs` package: 

1. **`create_new_data(newdf, olddf, by = NULL, by_col = NULL, cols = NULL)`**: takes two datasets and returns a dataset of, "*what is here now that wasn't here before?*

2. **`create_deleted_data(newdf, olddf, by = NULL, by_col = NULL, cols = NULL)`**: takes two datasets and creates a dataset of, "*What was here before that isn't here now?*  

3. **`create_modified_data(newdf, olddf, by = NULL, by_col = NULL, cols = NULL)`**: takes two datasets and creates a dataset of, "*What was here before that isn't here now?*  

These are all stored in the helpers.R file (sourced below):

```{r helpers}
source("helpers.R")
```

## UI  

The UI

```{r compareDataUI}
compareDataUI <- function(id) {
  tagList(
    fluidRow(
      bs4Dash::sortable(
        width = 12,
        bs4Dash::box(
          solidHeader = TRUE,
          status = "danger",
          width = 12,
          title = "Join columns",
          fluidRow(
            column(
              width = 6,
          ## OUTPUT |-- (intersecting_cols) ------
          strong(em("The following columns are in both tables:")),
          reactableOutput(
            outputId = NS(
              namespace = id,
              id = "intersecting_cols"))
            ),
            column(
              width = 6,
          ## OUTPUT |-- (new_join_col) ------
        strong(em("Select the column (or columns) for joining the two datasets")),
          selectInput(inputId = NS(
              namespace = id,
              id = "new_join_col"), 
              label = "Select Join Columns",             
              choices = c("<PREVIOUS COLUMN A>", "<PREVIOUS COLUMN B>"),
              multiple = TRUE,
              selected = c("<PREVIOUS COLUMN A>"))
              )
            )
          )
        ),
        bs4Dash::box(
          width = 12,
          title = "Run Comparison",
          solidHeader = TRUE,
          status = "primary",
          h5(em(".")),
          fluidRow(
            sortable(
              width = 12,
              column(
                width = 6,
                ## OUTPUT |-- (dev_display_x) ------
                strong("dev_display_x"),
                verbatimTextOutput(
                  outputId = NS(
                    namespace = id,
                    id = "dev_display_x"
                  )
                )
              )
            )
          ),
          fluidRow(
            sortable(
              width = 6,
              column(
                width = 6,
                ## OUTPUT |-- (dev_display_y) ------
                strong("dev_display_y"),
                verbatimTextOutput(
                  outputId = NS(
                    namespace = id,
                    id = "dev_display_y"
                  )
                )
              )
            )
          )
        )
      )
    )
}
```


## Server 

```{r compareDataServer}
compareDataServer <- function(id, data_selected) {
  moduleServer(id = id, module = function(input, output, session) {

    # |-- REACTIVE data_prev() --------------------------------------
    data_prev <- eventReactive( data_selected$prev_data(), {
      data_prev <- data_selected$prev_data()
      return(data_prev)
    })

    # |-- REACTIVE data_curr() --------------------------------------
    data_curr <- eventReactive( data_selected$curr_data(), {
      data_curr <- data_selected$curr_data()
      return(data_curr)
    })

    # |-- OUTPUT (dev_display_x) --------------------------------------
    output$dev_display_x <- renderPrint({
      print(str(data_prev()))
    })

    # |-- OUTPUT (dev_display_y) --------------------------------------
    output$dev_display_y <- renderPrint({
      print(str(data_curr()))
    })
    # |-- REACTIVE col_intersect() --------------------------------------
    col_intersect <- reactive({
      prev_cols <- names(data_prev())
      curr_cols <- names(data_curr())
      intersecting_cols <- intersect(x = prev_cols, y = curr_cols)
      col_intersect <- tibble::tibble(Columns = intersecting_cols)
      return(col_intersect)
    })

    # |-- OUTPUT (intersecting_cols) --------------------------------------
    output$intersecting_cols <- renderReactable({
      reactable(
        col_intersect(),
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
        theme = reactableTheme(
        color = "hsl(0, 0%, 100%)",
        backgroundColor = "hsl(235, 48%, 40%)",
        borderColor = "hsl(233, 9%, 22%)",
        stripedColor = "hsl(233, 12%, 22%)",
        highlightColor = "hsl(233, 12%, 24%)",
        inputStyle = list(backgroundColor = "hsl(0,0%,39%)"),
        selectStyle = list(backgroundColor = "hsl(0,0%,39%)"),
        pageButtonHoverStyle = list(backgroundColor = "hsl(0,0%,39%)"),
        pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)"))
      )
    })



  })
}
```


## Demo

```{r bmrn_theme}
bmrn_theme <- function() {
  fresh::create_theme(
    # theme vars  -------------------------------------------------------------
    fresh::bs4dash_vars(
      navbar_light_color = "#353d98", # purple
      navbar_light_active_color = "#353d98", # purple
      navbar_light_hover_color = "#f26631" # orange
    ),
    # # theme yiq -------------------------------------------------------------
    fresh::bs4dash_yiq(
      contrasted_threshold = 255,
      text_dark = "#0a0a0a", # dark_gray_s10
      text_light = "#f5f5f5" # gray_t10
    ),
    # theme layout ---------------------------------------------------------
    fresh::bs4dash_layout(
      main_bg = NULL, # #ececec
      font_size_root = 12
    ),
    # theme sidebar_light -------------------------------------------------
    fresh::bs4dash_sidebar_light(
      header_color = "#ccd5dd", # dark_blue_t9
      bg = "#eaebf4", # background of entire side-bar
      color = "#002E56", # text color (no hover)
      hover_color = "#ee304e", # text color on hover
      hover_bg = "#353D98", # color on hover
      active_color = "#f26631", # color is actually the 'primary' status color
      submenu_bg = "#f5f5f5", # purple
      submenu_color = "#002444",
      submenu_hover_color = "#353D98" # purple
    ),
    # # theme sidebar_dark -------------------------------------------------
    fresh::bs4dash_sidebar_dark(
      header_color = "#ccd5dd",
      bg = "#1a1e4c",
      color = "#EE304E", # text color (no hover)
      hover_bg = "#aeb1d5", # color on hover
      hover_color = "#EE304E", # text color on hover
      active_color = "#f26631" # color is actually the 'primary' status color
    ),
   # theme status -------------------------------------------------
    fresh::bs4dash_status(
      dark = "#323232",
      light = "#A0A0A0",
      warning = "#F26631", # orange
      primary = "#00509C", # blue
      secondary = "#353D98", # purple
      success = "#A9218E", # violet
      danger = "#EE304E", # red
      info = "#A0A0A0" #orange
    ),
    # theme color -------------------------------------------------
  fresh::bs4dash_color(
    gray_900 = "#1f245b",
    gray_800 = "#646464",
    lightblue = "#6696c3", 
    blue = "#00509C" 
    # gray_600 = "#353D98",
    # yellow = "#F26631",
    # gray_900 = "#15183c",
    # gray_800 = "#646464",
    # green = "#A9218E",
    # navy = "#002E56",
    # cyan = "#A0A0A0",
    # gray_800 = "#646464",
    # red = "#EE304E",
    # white = "#272c30"
   )
  )
}
```


```{r compareDataDemo}
# set theme ---------------------------------------------------------------
compare_theme <- bmrn_theme()

compareDataDemo <- function() {
  ui <- bs4Dash::dashboardPage(
    skin = 'light',
    freshTheme = compare_theme,
    title = "compareDataDemo",
    header = bs4Dash::dashboardHeader(
      title = "compareDataDemo",
      status = "secondary"),
    # sidebar (menuItem) ------------------------------------------------------
    sidebar = bs4Dash::dashboardSidebar(
      skin = "light",
      collapsed = TRUE,
      bs4Dash::sidebarMenu(
        id = "sidebarmenu",
        # bs4Dash::sidebarHeader("Data upload demo"),
        menuItem("Upload Data",
          tabName = "upload_data_tab",
          icon = icon("file-upload")
        ),
        menuItem("Select Data",
          tabName = "select_data_tab",
          icon = icon("columns")
        ),
        menuItem("Compare Data",
          tabName = "compare_data_tab",
          icon = icon("compress-alt")
        ),
        menuItem("Tables",
          tabName = "table_output_tab",
          icon = icon("table")
        )
      )
    ),
    # dashboardBody (tabItem) ------------------------------------------------
    body = bs4Dash::dashboardBody(
      tabItems(
        tabItem(
          tabName = "upload_data_tab",
          ## uploadDataUI -----
          uploadDataUI(id = "upload_data"),
        ),
        tabItem(
          tabName = "select_data_tab",
          ## selectDataUI -----
          selectDataUI(id = "select_data")
        ),
        tabItem(
          tabName = "compare_data_tab",
          ## compareDataUI -----
          compareDataUI("compare_data")
        ),
        tabItem(
          tabName = "table_output_tab",
          ## tableOutputUI -----
          # tableOutputUI("tables")
        )
      )
    ),
    controlbar = bs4Dash::dashboardControlbar(),
    footer = bs4Dash::dashboardFooter()
  )
  server <- function(input, output, session) {
    # upload data ------------------------------------------------
    upload_data_list <- uploadDataServer(id = "upload_data")
    # select data ------------------------------------------------
    select_data_list <- selectDataServer(
      id = "select_data",
      data_upload = upload_data_list
    )
    # compare data ------------------------------------------------
    compareDataServer(
      id = "compare_data",
      data_selected = select_data_list)
  }
  # tableOutputServer ------------------------------------------------
  shinyApp(
    ui = ui, server = server, options = list(height = 1000, width = 800)
  )
}
```

# App

```{r run-compareDataDemo, echo=FALSE}
compareDataDemo()
```
