---
title: "compareDataDemo(): compare module (dev)"
author: "Martin Frigaard"
output:
  html_document: 
    highlight: espresso
    theme: darkly
    toc: true
    toc_float: yes
    df_print: paged
    code_folding: hide
    
runtime: shiny
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(fs)
library(reactable)
library(rmarkdown)
library(shiny)
library(shinythemes)
library(bs4Dash)
# Global options
knitr::opts_chunk$set(
  # cache = FALSE # cache data
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "tiny", # size of the text
  fig.path = "images", # location of figure files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
```

# Purpose

The goal of `compareDataDemo` is to create a demo module for comparing two tables to get 1) new data, 2) deleted data, and 3) changed data in the `dfdiffs` app. View the previous modules below: 

1. [UploadDataDemo](https://mjfrigaard.shinyapps.io/uploadDataDemo/): this module is used to upload two datasets for comparison. 

2. [SelectDataDemo](https://mjfrigaard.shinyapps.io/selectDataDemo/): after uploading data into the application, this module works like a filter--selecting a data source and the columns from that source (*optional*).

# Helpers 

The functions for comparing any two datasets in the `dfdiffs` package: 

1. **`create_new_data(compare = , base = , by = , by_col = , cols = )`**: takes two datasets and returns a dataset of, "*what is here now that wasn't here before?*

2. **`create_deleted_data(compare = , base = , by = , by_col = , cols = )`**: takes two datasets and creates a dataset of, "*What was here before that isn't here now?*  

3. **`create_modified_data(compare = , base = , by = , by_col = , cols = )`**: takes two datasets and creates a dataset of, "*What has changed?*  

4. **`create_changed_data(compare = , base = , by = , by_col = , cols = )`**: takes two datasets and creates a dataset of, "*What has changed?*  

These are all stored in the `helpers.R` file (sourced below):

```{r helpers}
source("helpers.R")
```

# Comparisons 

We're going to import two datasets to simulate the application's compare functions. We'll be using the Master dataset from the [Lahman baseball database](https://github.com/cdalzell/Lahman):

Our 'base' or 'target' dataset is the 2015 Master dataset. 

```{r master15}
master15 <- readxl::read_excel("data/xlsx/lahman_compare.xlsx", 
                                sheet = "master-2015")
```

Our 'compare' or 'new' dataset is the 2020 People dataset 

```{r master20}
master20 <- readxl::read_excel("data/xlsx/lahman_compare.xlsx", 
                                sheet = "master-2020")
```

```{r export, eval=FALSE, echo=FALSE}
select(master15, 
  playerID, contains("name"), 
  contains("birth"), contains("death")) |> 
  write_csv("data/csv/m15.csv")

select(master20, 
  playerID, contains("name"), 
  contains("birth"), contains("death")) |> 
  write_csv(file = "data/csv/m20.csv")
```

## Uploaded data display 

Both base and compare datasets will be differentiated by color in the application using the `reactable::reactableTheme()` function.

### Base display

The base color will use the `#761763` background color (based on the BioMarin violet (`#a9218e`) :


```{r base_react_theme}
base_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#761763",
          borderColor = "#646464",
          stripedColor = "#3A3B45",
          highlightColor = "#a9218e",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
reactable::reactable(master15, 
  theme = base_react_theme
  )
```

### Compare display

The compare color will use the `#2f3688` background color (based on the BioMarin purple (`#353d98`) :

```{r compare_react_theme}
compare_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#2f3688",
          borderColor = "#646464",
          stripedColor = "#3A3B45",
          highlightColor = "#353d98",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
reactable::reactable(master20, 
  theme = compare_react_theme
  )
```


## New Records 

We'll test the `create_new_data()` function on the 2015 vs. 2020 Master data. 

```{r create_new_data, class.source='fold-show'}
# run new data function
new <- create_new_data(master20, master15, 
  by = "playerID",
  cols = c("playerID", "nameFirst", "nameLast", "nameGiven", "birthCity"))
```

We'll also apply the `new_react_theme` to this display: 


```{r reactable-new}
new_react_theme <- reactableTheme(
          color = "#00509C",
          backgroundColor = "#FFFFFF",
          borderColor = "#A0A0A0",
          stripedColor = "#3A3B45",
          highlightColor = "#eeeeee",
          inputStyle = list(backgroundColor = "#eeeeee"),
          selectStyle = list(backgroundColor = "#eeeeee"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
reactable(new,
  defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE, 
  theme = new_react_theme)
```

In order to check these data, we can test the unique `playerID`s in and see if they're in the `master15` dataset. 

```{r new_playerID, class.source='fold-show'}
# unique ids
new_playerID <- unique(new$playerID)
# filter new ids (should be zero) 
reactable(filter(master15, 
  playerID %in% new_playerID),
  defaultPageSize = 10,
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE, 
  theme = new_react_theme)
```



And we can see **none** of these `playerID`s are in the base dataset.

## Deleted Records 

We're going to check the deleted records using the `create_deleted_data()` function. 

```{r create_deleted_data, class.source='fold-show'}
deleted <- create_deleted_data(
  compare = master20, 
  base = master15, 
  by = "playerID",
  cols = c("playerID", "nameFirst", "nameLast", "nameGiven", "birthCity"))
```

We'll also apply the `deleted_react_theme` to this display: 

```{r reactable-deleted}
deleted_react_theme <- reactableTheme(
          color = "#d62b46",
          backgroundColor = "#FFFFFF",
          borderColor = "#A0A0A0",
          stripedColor = "#3A3B45",
          highlightColor = "#eeeeee",
          inputStyle = list(backgroundColor = "#eeeeee"),
          selectStyle = list(backgroundColor = "#eeeeee"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )

reactable(deleted,
    defaultPageSize = 10, 
    theme = deleted_react_theme,
    resizable = TRUE,
    highlight = TRUE,
    compact = TRUE,
    wrap = FALSE,
    bordered = TRUE,
    filterable = TRUE)
```

This tells us there were 625 deleted records between 2015 and 2020 in the master dataset. 

## Modified Records 

Below we test our modified function, `create_modified_data()`, which returns a list of two datasets. 

```{r create_modified_data, class.source='fold-show'}
modified <- create_modified_data(
  compare = master20, 
  base = master15, 
  by = "playerID",
  cols = c("playerID", "nameFirst", "nameLast", "nameGiven", "birthCity"))
names(modified)
```

The `diffs_byvar` table contains each variable modified (`Variable name`), the count of changes ( `Modified values`) and the count of missing (`Missing Values`). We'll also apply the `deleted_react_theme` to these displays: 

```{r modified-diffs_byvar}
changed_react_theme <- reactableTheme(
          color = "#c15127",
          backgroundColor = "#FFFFFF",
          borderColor = "#646464",
          stripedColor = "#3A3B45",
          highlightColor = "#eeeeee",
          inputStyle = list(backgroundColor = "#eeeeee"),
          selectStyle = list(backgroundColor = "#eeeeee"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
reactable(data = modified$diffs_byvar,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```

The diffs table contains the modified variable (`Variable Name`), the joining column (`playerID`), and the modified values (`comapre Value` & `base Value`). 

```{r modified-diffs}
reactable(data = modified$diffs,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```

## Changed Records 

```{r changed, class.source='fold-show'}
changed <- create_changed_data(
  compare = master20, 
  base = master15, 
  by = "playerID", 
  cols = c("playerID", "nameGiven", "height", "bats", "throws"))
names(changed)
```

The `base_diffs` table contains `playerIDs in BASE that are not in COMPARE`

```{r base_diffs}
reactable(data = changed$base_diffs,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```

The `base_diffs` table contains `playerIDs in COMPARE that are not in BASE`

```{r comp_diffs}
reactable(data = changed$comp_diffs,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```

The `num_diffs` table contains counts of all the variables with differences. 

```{r num_diffs}
reactable(data = changed$num_diffs,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```

The `var_diffs` table contains row-by-row differences.  

```{r var_diffs}
reactable(data = changed$var_diffs,
  defaultPageSize = 10,
  theme = changed_react_theme, 
  resizable = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  bordered = TRUE,
  filterable = TRUE)
```



All of these tables will be presented in the `compareDataUI()`

# Upload

The application is broken into four modules: `upload`, `select`, `compare`, and `tables`. The two functions for the upload modules are loaded below: 

## UI

```{r uploadDataUI}
source("uploadDataUI.R")
```

## Server

```{r uploadDataServer}
source("uploadDataServer.R")
```

These two files are not shown because they are displayed and covered in detail in base documentation. The `uploadData` module returns a list of reactive datasets to be used in the `selectData` module. 

# Select 

The `uploadData` module 'passes' the data to the `selectData` module, which displays the data for the user to review and, if necessary, reduce. The `uploadData` module can import up to four data files, but the `selectData`  trims that down to two datasets (and the selected columns). 

## UI

```{r selectDataUI}
source("selectDataUI.R")
```

## Server 

```{r selectDataServer}
source("selectDataServer.R")
```


# Compare 

The `compareData` module has 

## UI  

The UI

```{r compareDataUI, class.source='fold-hide'}

```


## Server 

```{r compareDataServer, class.source='fold-hide'}

```


# compareDataDemo()

```{r compareDataDemo, class.source='fold-hide'}
compareDataUI <- function(id) {
  tagList(
    bs4Dash::sortable(
      width = 12,
      bs4Dash::box(
        solidHeader = TRUE,
        status = "warning",
        width = 12,
        title = strong("Join columns"),
        fluidRow(
          column(
            width = 5,
            ## OUTPUT |-- (intersecting_cols) ------
            h5(
              strong(
                em("The following columns are in both tables:")
              )
            ),
            br(),
            reactableOutput(
              outputId = NS(
                namespace = id,
                id = "intersecting_cols"
              )
            )
          ),
          column(
            width = 6,
            ## INPUT |-- (by) ------
            h5(
              strong(
                em("Select Join Columns")
              )
            ),
            selectizeInput(
              inputId = NS(
                namespace = id,
                id = "by"
              ),
              label = "Select the column (or columns) to create a unique id",
              choices = c("", NULL),
              multiple = TRUE,
              selected = c("", NULL)
            ),
          em("Leave blank for a row-by-row comparison"),
            ## INPUT |-- (by_col) ------
            br(), br(),
            h5(
              strong(
                em("Join Column Name")
              )
            ),
            textInput(
              inputId = NS(namespace = id, id = "by_col"),
              label = "New join column name",
              value = NULL
            ),
          em("If no name is provided, the join column will be named", code("join")),
          )
        )
      )
    ),
    bs4Dash::sortable(
      bs4Dash::box(
        width = 12,
        title = strong("The ", code("base"), " data"),
        solidHeader = FALSE,
        status = "primary",
        fluidRow(
          column(
            width = 12,
            ## OUTPUT |-- (base_join_col_display) ------
            # br(),
            # h5(strong(
            #   em("The ", code("base"), " data are displayed below"))),
            # br(),
            reactableOutput(
              outputId = NS(
                namespace = id,
                id = "base_join_col_display"
              )
            )
          )
        )
      )
    ),
    bs4Dash::sortable(
      bs4Dash::box(
        width = 12,
        title = strong("The ", code("compare"), " data"),
        solidHeader = FALSE,
        status = "secondary",
        fluidRow(
          column(
            width = 12,
            ## OUTPUT |-- (comp_join_col_display) ------
            # br(),
            # h5(strong(
            #   em("The ", code("compare"), " data are displayed below"))),
            # br(),
            reactableOutput(
              outputId = NS(
                namespace = id,
                id = "comp_join_col_display"
              )
            )
          )
        )
      )
    ),
    bs4Dash::sortable(
      bs4Dash::box(
        width = 12,
        title = strong("New Data"),
        solidHeader = FALSE,
        status = "success",
        fluidRow(
          sortable(
            width = 12,
            column(
              width = 12,
              ## OUTPUT |-- (go_new_data) ------
              h5("The new data between", code("base"),
                " and ", code("compare"), " are below:"),
              actionButton(
                inputId = NS(
                  namespace = id,
                  id = "go_new_data"
                ),
                label = strong("Get new data!"),
                status = "info"
              ),
              br(), br(),
              reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "new_data_display"
                )
              )
            )
          )
        )
      )
    ),
    bs4Dash::sortable(
      bs4Dash::box(
        width = 12,
        title = strong("Deleted Data"),
        solidHeader = FALSE,
        status = "success",
        fluidRow(
          sortable(
            width = 12,
            column(
              width = 12,
              ## OUTPUT |-- (dev_display_y) ------
              h5("The deleted data between", code("base"),
                " and ", code("compare"), " are below:"),
              actionButton(
                inputId = NS(
                  namespace = id,
                  id = "go_deleted_data"
                ),
                status = "info",
                label = strong("Get deleted data!")
              ),
              br(),
              reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "deleted_data_display"
                )
              )
            )
          )
        )
      )
    ),
    bs4Dash::sortable(
      bs4Dash::box(
        width = 12,
        title = "DEV",
        solidHeader = TRUE,
        status = "info",
        h5(em("DEV")),
        fluidRow(
          sortable(
            width = 12,
            column(
              width = 6,
              ## OUTPUT |-- (dev_display_x) ------
              code("input$by_col"),
              verbatimTextOutput(
                outputId = NS(
                  namespace = id,
                  id = "dev_display_x"
                )
              )
            ),
            column(
              width = 6,
              ## OUTPUT |-- (dev_display_y) ------
              code("input$by"),
              verbatimTextOutput(
                outputId = NS(
                  namespace = id,
                  id = "dev_display_y"
                )
              )
            )
          )
        )
      )
    )
  )
}
compareDataServer <- function(id, data_selected) {
  moduleServer(id = id, module = function(input, output, session) {

    # |-- REACTIVE data_base() --------------------------------------
    data_base <- eventReactive(data_selected$base_data(), {
      data_base <- data_selected$base_data()
      return(data_base)
    })

    # |-- REACTIVE data_comp() --------------------------------------
    data_comp <- eventReactive(data_selected$comp_data(), {
      data_comp <- data_selected$comp_data()
      return(data_comp)
    })

    # |-- OUTPUT (dev_display_x) --------------------------------------
    output$dev_display_x <- renderPrint({
      print(input$by_col)
    })

    # |-- OUTPUT (dev_display_y) --------------------------------------
    output$dev_display_y <- renderPrint({
      print(input$by)
    })

    # |-- REACTIVE col_intersect() --------------------------------------
    col_intersect <- reactive({
      base_cols <- names(data_base())
      comp_cols <- names(data_comp())
      intersecting_cols <- intersect(x = base_cols, y = comp_cols)
      col_intersect <- tibble::tibble(Columns = intersecting_cols)
      return(col_intersect)
    })

    # |-- OUTPUT (intersecting_cols) --------------------------------------
    output$intersecting_cols <- renderReactable({
      reactable(
        col_intersect(),
        resizable = TRUE,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        defaultPageSize = 5,
        theme = reactableTheme(
        color = "#2a3079",
        borderColor = "#e5eaee",
        stripedColor = "#f6f8fa",
        highlightColor = "#f0f5f9",
        cellPadding = "8px 12px"
        )
      )
    })

    ##  UPDATE |-- input$by   ---------
    observeEvent(col_intersect(), {
      data_choices <- col_intersect()$Columns
      updateSelectizeInput(
        inputId = "by",
        choices = data_choices,
        selected = NULL
      )
    })

    ##  REACTIVE |--  base_join_col_data  ---------
    base_join_col_data <- reactive({
       # no by col, no new name
      if (length(input$by) == 0 & nchar(input$by_col) == 0) {
        base_join_col <- data_base()
      # by col, no by_col
      } else if (length(input$by) > 0 & nchar(input$by_col) == 0) {
        base_join_col <- create_join_column(
          df = data_base(),
          by_colums = input$by,
          new_by_column_name = "join"
        )
        # no by col, new name
      } else if (length(input$by) == 0 & nchar(input$by_col) != 0) {
        base_join_col <- data_base()
        # by col and new col name
      } else if (length(input$by) > 0 & nchar(input$by_col) != 0) {
        base_join_col <- create_join_column(
          df = data_base(),
          by_colums = input$by,
          new_by_column_name = as.character(input$by_col)
        )
        # no by col, no new name
      } else {
        base_join_col <- data_base()
      }
    })

    ##  REACTIVE |--  comp_join_col_data ---------
    comp_join_col_data <- reactive({
      # no by col, no new name
      if (length(input$by) == 0 & nchar(input$by_col) == 0) {
        comp_join_col <- data_comp()
      # by col, no by_col
      } else if (length(input$by) > 0 & nchar(input$by_col) == 0) {
        comp_join_col <- create_join_column(
          df = data_comp(),
          by_colums = input$by,
          new_by_column_name = "join"
        )
        # no by col, new name
      } else if (length(input$by) == 0 & nchar(input$by_col) != 0) {
        comp_join_col <- data_comp()
        # by col and new col name
      } else if (length(input$by) > 0 & nchar(input$by_col) != 0) {
        comp_join_col <- create_join_column(
          df = data_comp(),
          by_colums = input$by,
          new_by_column_name = as.character(input$by_col)
        )
        # no by col, no new name
      } else {
        comp_join_col <- data_comp()
      }
    })



    # |-- OUTPUT (base_join_col_display) --------------------------------------
    output$base_join_col_display <- renderReactable({
      reactable(
        data = base_join_col_data(),
        resizable = TRUE,
        defaultPageSize = 5,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE,
        theme = base_react_theme
      )
    })

    # |-- OUTPUT (comp_join_col_display) --------------------------------------
    output$comp_join_col_display <- renderReactable({
      reactable(
        data = comp_join_col_data(),
        resizable = TRUE,
        defaultPageSize = 5,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE,
        theme = compare_react_theme
      )
    })

    ##  REACTIVE |--  new_data ---------
    new_data <- reactive({
      # no by_col, no by
      if (nchar(input$by_col) == 0 & length(input$by) == 0) {
          new <- create_new_data(
                  compare = data_comp(),
                  base = data_base())
          # by_col, no by
      } else if (nchar(input$by_col) > 0 & length(input$by) == 0) {
          new <- create_new_data(
                  compare = data_comp(),
                  base = data_base())
          # no by_col, by
      } else if (nchar(input$by_col) == 0 & length(input$by) > 0) {
          new <- create_new_data(
                  compare = data_comp(),
                  base = data_base(),
                  by = as.character(input$by))
          # by_col, by
      } else if (nchar(input$by_col) > 0 & length(input$by) > 0) {
          new <- create_new_data(
                  compare = data_comp(),
                  base = data_base(),
                  by = as.character(input$by),
                  by_col = as.character(input$by_col))
          # all else
      } else {
          new <- create_new_data(
                  compare = data_comp(),
                  base = data_base())
      }
        return(new)
    })

    # |-- OUTPUT (new_data_display) --------------------------------------
    observeEvent(input$go_new_data, {
    output$new_data_display <- renderReactable({
      reactable(
        data = new_data(),
        resizable = TRUE,
        defaultPageSize = 5,
        highlight = TRUE,
        compact = TRUE,
        wrap = FALSE,
        bordered = TRUE,
        filterable = TRUE,
        theme = new_react_theme
      )
    })

    })


    ##  REACTIVE |--  deleted_data ---------

    ##  REACTIVE |--  new_data ---------


  })
}
compareDataDemo <- function() {
  ui <- bs4Dash::dashboardPage(
    skin = "light",
    freshTheme = compare_theme,
    title = "compareDataDemo",
    header = bs4Dash::dashboardHeader(
      title = "compareDataDemo",
      status = "secondary"
    ),
    # sidebar (menuItem) ------------------------------------------------------
    sidebar = bs4Dash::dashboardSidebar(
      skin = "light",
      collapsed = TRUE,
      bs4Dash::sidebarMenu(
        id = "sidebarmenu",
        # bs4Dash::sidebarHeader("Data upload demo"),
        menuItem("Upload Data",
          tabName = "upload_data_tab",
          icon = icon("file-upload")
        ),
        menuItem("Select Data",
          tabName = "select_data_tab",
          icon = icon("columns")
        ),
        menuItem("Compare Data",
          tabName = "compare_data_tab",
          icon = icon("compress-alt")
        ),
        menuItem("Tables",
          tabName = "table_output_tab",
          icon = icon("table")
        )
      )
    ),
    # dashboardBody (tabItem) ------------------------------------------------
    body = bs4Dash::dashboardBody(
      tabItems(
        tabItem(
          tabName = "upload_data_tab",
          ## uploadDataUI -----
          uploadDataUI(id = "upload_data"),
        ),
        tabItem(
          tabName = "select_data_tab",
          ## selectDataUI -----
          selectDataUI(id = "select_data")
        ),
        tabItem(
          tabName = "compare_data_tab",
          ## compareDataUI -----
          compareDataUI("compare_data")
        ),
        tabItem(
          tabName = "table_output_tab",
          ## tableOutputUI -----
          # tableOutputUI("tables")
        )
      )
    ),
    controlbar = bs4Dash::dashboardControlbar(),
    footer = bs4Dash::dashboardFooter()
  )
  server <- function(input, output, session) {
    # upload data ------------------------------------------------
    upload_data_list <- uploadDataServer(id = "upload_data")
    # select data ------------------------------------------------
    select_data_list <- selectDataServer(
      id = "select_data",
      data_upload = upload_data_list
    )
    # compare data ------------------------------------------------
    compareDataServer(
      id = "compare_data",
      data_selected = select_data_list
    )
  }
  # tableOutputServer ------------------------------------------------
  shinyApp(
    ui = ui, server = server, options = list(height = 1000, width = 800)
  )
}
```

# App

```{r run-compareDataDemo, echo=FALSE}
compareDataDemo()
```
