



## `diffdf::diffdf()`

```{r args, eval=FALSE}
diffdf(base, compare, keys = NULL, suppress_warnings = FALSE, 
    strict_numeric = TRUE, strict_factor = TRUE, file = NULL, 
    tolerance = sqrt(.Machine$double.eps), scale = NULL)
```

```{r }
diffdf::diffdf(base = InitialData, compare = ChangedData)
```


```{r}
diffdf::diffdf(base = InitialData, compare = ChangedData, 
  keys = c("subject_id", "record"))
```

```{r diffdf_output}
diffdf_output <- diffdf::diffdf(
  base = master15 , 
  compare = master20, 
  keys = "playerID")
names(diffdf_output)
```


We can see this comparison gives us a larger list of names in the returned object, including `ExtRowsBase` and `ExtRowsComp`. 

#### Rows in `BASE` that are not in `COMPARE`

```{r ExtRowsBase}
ext_rows_base <- diffdf_output$ExtRowsBase
key <- "playerID"
base_msg <- attr(ext_rows_base,"message")
msg_rep <- str_replace_all(base_msg, "rows", paste0(key, "s"))
msg_rem <- str_remove_all(msg_rep, "There are | !!")
ext_rows_base <- purrr::set_names(
  x = tibble::as_tibble(ext_rows_base), nm = msg_rem)
```

```{r ext_rows_base}
head(ext_rows_base)
```

#### Rows in `COMPARE` that are not in `BASE`

```{r ExtRowsComp}
ext_rows_comp <- diffdf_output$ExtRowsComp
comp_msg <- attr(ext_rows_comp,"message")
msg_rep <- stringr::str_replace_all(comp_msg, "rows", paste0(key, "s"))
msg_rem <- stringr::str_remove_all(msg_rep, "There are | !!")
ext_rows_comp <- purrr::set_names(
  x = tibble::as_tibble(ext_rows_comp), nm = msg_rem)
```


```{r ext_rows_comp}
head(ext_rows_comp)
```

## extract_diffs()

We can capture the output from `diffdf()` and return the differences in a more 'table' friendly output. 

```{r extract_diffs, echo=TRUE}
extract_diffs <- function(base, compare, extract, keys = NULL) {
  
  diff_lst <- diffdf::diffdf(base = base , compare = compare, keys = keys)
  # get names 
  diff_lst_nms <- base::names(diff_lst)
  # extract nums
  numdiffs_lst <- diff_lst[stringr::str_detect(diff_lst_nms, "NumDiff")]
  # better names
  num_diffs <- purrr::map_df(.x = numdiffs_lst, .f = janitor::clean_names)
  # extract vars
  vardiffs_lst <- diff_lst[stringr::str_detect(diff_lst_nms, "VarDiff")]
  # convert to character
  chr_vardiffs_lst <- purrr::map(.x = vardiffs_lst,
                                      dplyr::mutate_all, base::as.character)
  # better names
  var_diffs <- purrr::map_df(.x = chr_vardiffs_lst, .f = janitor::clean_names)
  
  diffs <- list(
    "num_diffs" = num_diffs, 
    "var_diffs" = var_diffs
  )
  #   diffs <- diff_lst
  #   
  # 
  #   

  #   
  # } else {
  #   
  #   diffs <- diff_lst
  #   
  # }
  return(diffs)
}
```

## Test `extract_diffs()`

```{r test-extract_diffs-InitialData-ChangedData}
extract_diffs(base = InitialData, compare = ChangedData, extract = "num")
extract_diffs(base = InitialData, compare = ChangedData, extract = "var")
```

We will test this with a little larger datasets,

### extract = "base"

```{r master_base_diffs, echo=TRUE}
master_base_diffs <- extract_diffs(
  base = master15, 
  compare = master20, 
  keys = "playerID",
  extract = "base")
head(master_base_diffs)
```

### extract = "comp"

```{r master_comp_diffs, echo=TRUE}
master_comp_diffs <- extract_diffs(
  base = master15, 
  compare = master20, 
  keys = "playerID",
  extract = "comp")
head(master_comp_diffs)
```

### extract = "num"

```{r master_num_diffs}
master_num_diffs <- extract_diffs(
  base = master15, 
  compare = master20, 
  keys = "playerID",
  extract = "num")
head(master_num_diffs)
```


### extract = "var"

```{r master_var_diffs, echo=TRUE}
master_var_diffs <- extract_diffs(
  base = master15, 
  compare = master20, 
  keys = "playerID",
  extract = "var")
head(master_var_diffs)
```

