---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dfdiffs


<!-- badges: start -->
<!-- badges: end -->

The goal of `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   

The `dfdiffs` Shiny app-package is built with [`bslib`](https://rstudio.github.io/bslib/). Its comparison engine is pure base R, but it owes a debt to the [`diffdf`](https://gowerc.github.io/diffdf/) package: the tolerance-based numeric comparison in `compare_values()` is a direct port of `diffdf`'s own formula.

You can access a development version of the application [here](https://mjfrigaard.shinyapps.io/compareDataApp/).

## Installation

You can install the development version of dfdiffs from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mjfrigaard/dfdiffs")
```

```{r pkg}
library(dfdiffs)
```

## Packages

`dfdiffs` depends on:

```{r dep-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(bslib)
library(data.table)
library(ggplot2)
library(haven)
library(janitor)
library(openxlsx)
library(purrr)
library(reactable)
library(readxl)
library(shiny)
library(tibble)
```

`dfdiffs`'s own comparison engine (`create_new_data()`, `create_deleted_data()`, `create_changed_data()`, `create_modified_data()`, and the helpers behind them) is built on base R alone, so it doesn't require `dplyr`, `tidyr`, `tidyselect`, or `diffdf`. Those packages (plus `arsenal`) remain in `Suggests` only because a few vignettes demonstrate them directly, side-by-side, as points of comparison.

The Shiny app previews and displays data with [`reactable`](https://glin.github.io/reactable/) (an interactive JS widget). GitHub renders this README as plain markdown and strips the JS that a `reactable` table needs, so the tables below use `knitr::kable()`, which GitHub can render natively.

## Package functions

`dfdiffs` has a function for each of the questions posed above, and each function comes with a pair of example datasets that demonstrate how it works (which we'll cover below).

## *What rows are here now that weren't here before?*

To check for new data, we'll use `T1Data` and `T2Data`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
NewData <- dfdiffs::NewData
```

### Timepoint 1 data (original)

`T1Data` represents data collected at the first timepoint (T1).

```{r T1Data}
T1Data |> knitr::kable()
```

### Timepoint 2 data (new)

`T2Data` is the 'new' dataset, representing data collected at the second timepoint (T2).

```{r T2Data}
T2Data |> knitr::kable()
```

### `create_new_data()`

The `create_new_data()` function returns the 'new data' (i.e., the rows that are here now but weren't here before).

```{r}
create_new_data(
  compare = T2Data, 
  base = T1Data) |>
  knitr::kable()
```

We can check this against the `NewData` dataset, which should match the output from `create_new_data()`.

```{r NewData}
NewData |> knitr::kable()
```

## *What rows were here before that aren't here now?*

To test for deleted data, we'll compare `CompleteData` and `IncompleteData`, then check the result against `DeletedData`.

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
DeletedData <- dfdiffs::DeletedData
```

### A complete dataset

`CompleteData` represents a 'complete' set of data. 

```{r CompleteData}
CompleteData |> knitr::kable()
```


### An incomplete dataset

This is a dataset with rows removed from `CompleteData`.

```{r IncompleteData}
IncompleteData |> knitr::kable()
```

### `create_deleted_data()`

Running `create_deleted_data()` checks for rows that were deleted between `CompleteData` and `IncompleteData`.

```{r create_deleted_data, echo=TRUE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) |>
  knitr::kable()
```

### The deleted data 

The output above is identical to the data stored in `DeletedData`.

```{r DeletedData}
DeletedData |> knitr::kable()
```

## *What values have been changed?*

To answer this question, we have two options: `create_changed_data()` and `create_modified_data()`. Both build on the same `diff_values_by_key()`/`compare_values()` engine and differ only in the shape of their output: `create_changed_data()` returns `$num_diffs`/`$var_diffs` (used by `compare_data()`), and `create_modified_data()` returns `$diffs_byvar`/`$diffs` (used by `mod_compare` and `create_comparison_report()`).

To check for changes between two datasets, we'll use `InitialData` and `ChangedData`.

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
```

### Initial data 

```{r InitialData}
InitialData |> knitr::kable()
```

### Changed data

```{r ChangedData}
ChangedData |> knitr::kable()
```

### `create_changed_data()` 

`create_changed_data()` creates a list of tables. 

```{r changed}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
names(changed)
```

#### Counts of changes (`num_diffs`)

The counts of changes by variable are stored in `num_diffs`.

```{r num_diffs}
changed$num_diffs |> knitr::kable()
```

#### Changes by row (`var_diffs`)

The changes by row are stored in `var_diffs`.

```{r var_diffs}
changed$var_diffs |> knitr::kable()
```

### `create_modified_data()` 

The `create_modified_data()` function also creates a list of tables. 

```{r modified}
modified <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
names(modified)
```

#### Counts of changes (`diffs_byvar`)

The counts of changes by variable are stored in `diffs_byvar`.

```{r diffs_byvar}
modified$diffs_byvar |> knitr::kable()
```

#### Changes by row 

The changes by row are stored in `diffs`.

```{r diffs}
modified$diffs |> knitr::kable()
```

