---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dfdiffs


<!-- badges: start -->
<!-- badges: end -->

The goal of `dfdiffs` is to is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   

The `dfdiffs` package and application wouldn't be possible without the previous work from the authors of the [`arsenal`](https://mayoverse.github.io/arsenal/reference/arsenal.html) and [`diffdf`](https://gowerc.github.io/diffdf/) packages.

You can access a development version of the application [here.](https://mjfrigaard.shinyapps.io/compareDataApp/) 

## Installation

You can install the development version of dfdiffs from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mjfrigaard/dfdiffs")
```

```{r pkg}
library(dfdiffs)
```

## Packages 

Dependencies

```{r dep-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
library(janitor) 
library(arsenal) 
library(diffdf)  
library(gt)
library(labelled)
library(gtsummary)
```

## Package functions 

We have functions for answering each of the questions posed above. Each function has a pair of datasets to demonstrate how they work (which we'll cover below).

## *What rows are here now that weren't here before?*

To check new data, we're going to use `T1Data` and `T2Data`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
NewData <- dfdiffs::NewData
```

### Timepoint 1 data (original)

These data represent data taken at T1. 

```{r T1Data}
T1Data |> gt::gt()
```

### Timepoint 2 data (new)

This is a 'new' dataset representing T2.

```{r T2Data}
T2Data |> gt::gt()
```

### `create_new_data()`

The `create_new_data()` function shows us the 'new data' (i.e. what is here now that wasn't here before?)

```{r}
create_new_data(
  compare = T2Data, 
  base = T1Data) |> 
  gt::gt()
```

We can check this against the `NewData` dataset (which should match the output from `create_new_data()`)

```{r NewData}
NewData |> gt::gt()
```

## *What rows were here before that aren't here now?*

To test for the deleted data, we use the `CompleteData`, `IncompleteData`, and check these with `DeletedData`. 

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
DeletedData <- dfdiffs::DeletedData
```

### A complete dataset

`CompleteData` represents a 'complete' set of data. 

```{r CompleteData}
CompleteData |> gt::gt()
```


### An incomplete dataset

This is a dataset with rows removed from `CompleteData`.

```{r IncompleteData}
IncompleteData |> gt::gt()
```

### `create_deleted_data()`

When we run the `create_deleted_data()`, we check for the deleted rows between `IncompleteData` and `CompleteData`.

```{r create_deleted_data, echo=TRUE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) |> 
  gt::gt()
```

### The deleted data 

This is identical to the data stored in `DeletedData`

```{r DeletedData}
DeletedData |> gt::gt()
```

## *What values have been changed?*

To answer this question, we have two options: `create_changed_data()` and `create_modified_data()`.

+ `create_changed_data()` relies on the `diffdf()` function from the [`diffdf` package](https://gowerc.github.io/diffdf/reference/diffdf.html) package. 

+ `create_modified_data()` relies on the `comparedf()` function from the [`arsenal` package](https://mayoverse.github.io/arsenal/reference/comparedf.html) package.

To check for changes between two datasets, we use the `InitialData` and `ChangedData`. 

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
```

### Initial data 

```{r InitialData}
InitialData |> gt::gt()
```

### Changed data

```{r ChangedData}
ChangedData |> gt::gt()
```

### `create_changed_data()` 

`create_changed_data()` creates a list of tables. 

```{r changed}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
names(changed)
```

#### Counts of changes (`num_diffs`)

The counts of changes by variable are stored in `num_diffs`.

```{r num_diffs}
changed$num_diffs |> gt::gt()
```

#### Changes by row (`var_diffs`)

The changes by row are stored in `var_diffs`.

```{r var_diffs}
changed$var_diffs |> gt::gt()
```

### `create_modified_data()` 

The `create_modified_data()` function also creates a list of tables. 

```{r modified}
modified <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
names(modified)
```

#### Counts of changes (`diffs_byvar`)

The counts of changes by variable are stored in `diffs_byvar`.

```{r diffs_byvar}
modified$diffs_byvar |> gt::gt()
```

#### Changes by row 

The changes by row are stored in `diffs`.

```{r diffs}
modified$diffs |> gt::gt()
```

