

```{r , eval=FALSE, echo=FALSE}
    # multiple 'by', no 'by_col', no 'cols' ----
    } else if (length(by) > 1 & is.null(by_col) & is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = "JOIN")
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = "JOIN")
    # check
    # mod_data <- list(newdf_join, olddf_join)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join,
                                y = olddf_join,
                                by = "JOIN"))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                    dplyr::contains("JOIN"),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # multiple 'by', new 'by_col', no 'cols' -----
  } else if (length(by) > 1 & !is.null(by_col) & is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = {{by_col}})
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = {{by_col}})
    # check
    # mod_data <- list(newdf_join, olddf_join)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join,
                                y = olddf_join,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # multiple 'by', 'by_col', multiple 'cols' ----
  } else if (length(by) > 1 & !is.null(by_col) & !is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = {{by_col}})
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = {{by_col}})
    newdf_join_cols <- dplyr::select(newdf_join, {{by_col}}, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf_join, {{by_col}}, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # single 'by', no 'by_col', no 'cols' ----
  } else if (length(by) == 1 & is.null(by_col) & is.null(cols)) {
    newdf_join_cols <- dplyr::relocate(newdf, {{by}}, everything())
    olddf_join_cols <- dplyr::relocate(olddf, {{by}}, everything())
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # single 'by', 'by_col', no 'cols' ----
  } else if (length(by) == 1 & !is.null(by_col) & is.null(cols)) {
    names(newdf)[names(newdf) == by] <- by_col
    names(olddf)[names(olddf) == by] <- by_col
    newdf_join_cols <- newdf
    olddf_join_cols <- olddf
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # single 'by', 'by_col', multiple 'cols' -----
  } else if (length(by) == 1 & !is.null(by_col) & !is.null(cols)) {
    names(newdf)[names(newdf) == by] <- by_col
    names(olddf)[names(olddf) == by] <- by_col
    # newdf_join_cols <- newdf
    # olddf_join_cols <- olddf
    newdf_join_cols <- dplyr::select(newdf, {{by_col}}, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf, {{by_col}}, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # no 'by', no 'by_col', multiple 'cols' ----
  } else if (is.null(by) & is.null(by_col) & !is.null(cols)) {
    newdf_join_cols <- dplyr::select(newdf, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # single 'by_col', no 'by', no 'cols' ----
  } else if (length(by_col) == 1 & is.null(by) & is.null(cols)) {
    # only 'by' provided
    newdf_join_cols <- dplyr::mutate(newdf,
                                        ID = row_number(),
                                        ID = as.character(ID))
    olddf_join_cols <- dplyr::mutate(olddf,
                                        ID = row_number(),
                                        ID = as.character(ID))
    names(newdf_join_cols)[names(newdf_join_cols) == "ID"] <- by_col
    names(olddf_join_cols)[names(olddf_join_cols) == 'ID'] <- by_col
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    # comparedf_list
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols),
                                by = {{by_col}})
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                      contains("row.names"),
                    `Variable name` = var.x,
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    names(diffs_names)[names(diffs_names) == "..row.names.."] <- by_col

    mod_data <- tibble::as_tibble(diffs_names)
  }

  return(mod_data)
}
```

## Use 

Below we will test the `create_modified_data()` function under various conditions. 

### No arguments provided 

If only datasets are provided, a row-by-row comparison is made on all columns. 

```{r nothing, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData)
```

```{r nothing-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData)
)
```

### Multiple `by` columns

If we need to create a new `JOIN` column from multiple columns, we can provide them to `by`.

```{r multiple-by, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"))
```

```{r multiple-by-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"))
)
```

### Multiple `by` columns, new `by_col`

If we need to create a new joining column from multiple columns and we want to provide a name for this new column, we can do this with `by` and `by_col`:

```{r multiple-by-new-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col")
```

```{r multiple-by-new-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col")
)
```

### Multiple `by` columns, new `by_col`, and `cols`

If we need to create a new joining column from multiple columns, and we want to provide a name for this new column, *and* we want to only compare a few columns, we can do this with `by`, `by_col`, and `cols`. 

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col", 
  cols = c("text_value_a", "text_value_b"))
```

```{r multiple-by-new-by_col-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col", 
  cols = c("text_value_a", "text_value_b"))
)
```

### Single `by` column 

Single `by` column (assuming a join column exists), with no `by_col` or `cols`. 

```{r single-by-var, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var")
```

```{r single-by-var-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var")
)
```


### Single `by` column, new `by_col` name

Single `by` column (assuming a join column exists), with no `by_col` or `cols`. 

```{r single-by-var-new-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col")
```

```{r single-by-var-new-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col")
)
```


### Single `by` column, new `by_col` name, and `cols` 

```{r single-by-col-new-by_cols-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col",
  cols = c("text_value_a", "text_value_b")
  )
```

```{r single-by-col-new-by_cols-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col",
  cols = c("text_value_a", "text_value_b")
  ))
```


### Multiple `cols` 

If no `by` column is provided, and no `by_col`, but `cols` are provided, the function will perform a row-by-row comparison on the selected columns. 

```{r no-by-no-new-by_col-multiple-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  cols = c("text_value_a", "text_value_b"))
```

```{r no-by-no-new-by_col-multiple-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  cols = c("text_value_a", "text_value_b")))
```


### Only `by_col` 

If only a `by_col` is provided, the function will create an `ID` for each row in the datasets, then perform the comparison using this joining variable. 

```{r only-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  by_col = "new_join_col")
```

```{r only-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  by_col = "new_join_col"))
```



```{r old-fun, eval=FALSE, include=FALSE, echo=FALSE}
function(newdf, olddf, by = NULL, cols = NULL, return_cols = NULL) {
      # join col x
      data_x_join <- create_new_column(
                          data = newdf, 
                          cols = all_of(by), 
                          new_name = by) 
      # var names x
      data_x <- dplyr::rename_with(
        .data = data_x_join, 
        .fn =  base::paste0, 
        .cols = !all_of(by), "_x")
      # join col y
      data_y_join <- create_new_column(
                         data = olddf, 
                         cols = all_of(by), 
                         new_name = by) 
      # var names y
      data_y <- dplyr::rename_with(
        .data = data_y_join, 
        .fn =  base::paste0, 
        .cols = !all_of(by), "_y")
      # mod_table
      mod_table <- run_modified_data(newdf = newdf, 
                                     olddf = olddf, 
                                     by = by, 
                                     cols = cols)
    
      # join back to data_x
      data_x_mod <- dplyr::inner_join(
                             x = mod_table, 
                             y = data_x, 
                             by = all_of(by))
      # join back to data_y
      data_y_mod <- dplyr::inner_join(
                           x = mod_table, 
                           y = data_y, 
                           by = all_of(by))
      
      # join back to data_x and data_y
      data_both_mod <- dplyr::inner_join(x = data_x_mod, 
                               y = data_y, 
                               by = all_of(by))
  
}
```




```{r comparedf, eval=FALSE, include=FALSE}
arsenal::comparedf(x = , y = , by = , by.x = , by.y = , control = )
```




# Loading data 

Below we load three datasets from the [`Lahman` package](https://cran.r-project.org/web/packages/Lahman/index.html)

## 2010 Lahamn 'Master'

```{r lahman_2010_master_path}
lahman_2010_master_path <- list.files(
  path = "../inst/extdata/csv/2010-lahman", 
  full.names = TRUE, 
  pattern = "^Master")
master_2010_raw <- load_flat_file(path = lahman_2010_master_path)
master_2010 <- mutate(master_2010_raw, 
  debut = lubridate::mdy_hms(debut),
  finalGame = lubridate::mdy_hms(finalGame)
  )
glimpse(master_2010)
```

## 2015 Lahamn 'Master'

```{r master_2015}
lahman_2015_master_path <- list.files(
  path = "../inst/extdata/csv/2015-baseballdatabank", 
  full.names = TRUE, pattern = "^Master")
master_2015 <- load_flat_file(path = lahman_2015_master_path)
glimpse(master_2015)
```

## 2020 Lahamn 'People'

The name for this dataset changed from `Master` to `People`.

```{r people_2020}
lahman_2020_people_path <- list.files(
  path = "../inst/extdata/csv/2020-baseballdatabank/", 
  full.names = TRUE, pattern = "^People")
people_2020 <- load_flat_file(path = lahman_2020_people_path)
glimpse(people_2020)
```

# New data: "*What rows are here now that weren't here before?*"

```{r create_new_data-args, eval=FALSE}
dfdiffs::create_new_data(newdf = , olddf = , by = , by_col = , cols = )
```


We can test this function by looking at the new data between 2010, 2015, and 2020 datasets. 

## New data between 2010 - 2020

The new data between 2010 and 2020 is created below. 

```{r new_2010thru2020}
new_2010thru2020 <- create_new_data(newdf = people_2020, 
                                    olddf = master_2010, 
                                    by = "playerID")
```

The most straightforward way to check if this function is working is to get the `playerID`s that are in `people_2020` but not in `master_2010`. 

```{r new_2010thru2020_playerIDs}
new_2010thru2020_playerIDs <- unique(new_2010thru2020$playerID)
length(new_2010thru2020_playerIDs)
```

Now we can check to see if these names are the identical number of rows filtered to the `playerID`s in the 2020 data:

```{r identical-new_2010thru2020-new_2010thru2020_playerIDs}
identical(x = nrow(new_2010thru2020), 
  y = nrow(filter(people_2020, playerID %in% new_2010thru2020_playerIDs)))
```


We can also check the `debut` variable in the `new_2010thru2020` data. The earliest and latest dates for `debut` in `new_2010thru2020` are below:

```{r min-max-new_2010thru2020-debut}
min(new_2010thru2020$debut, na.rm = TRUE)
max(new_2010thru2020$debut, na.rm = TRUE)
```


If we filter each dataset to for the earliest date, we find the player who was added to the 2020 data:

```{r 1873-player}
people_2020_1873 <- filter(people_2020, debut == lubridate::as_date("1873-04-14"))
master_2010_1873 <- filter(master_2010, debut == lubridate::as_date("1873-04-14"))
player_debut_1873 <- setdiff(x = people_2020_1873$playerID, 
  y = master_2010_1873$playerID)
player_debut_1873
```

Our missing player is `r player_debut_1873`. If we use this player to check the 2010 and 2020 data, we can see this player is in the 2020 dataset...

```{r check-2020-player_debut_1873}
knitr::kable(filter(people_2020, playerID == player_debut_1873) %>% 
  select(playerID, debut))
```

...and not the 2010 data:

```{r check-2010-player_debut_1873}
knitr::kable(filter(master_2010, playerID == player_debut_1873) %>% 
  select(playerID, debut))
```

## New data between 2010 - 2015

Below we perform similar checks for 2010 vs. 2015 data

```{r new_2010thru2015}
new_2010thru2015 <- create_new_data(newdf = master_2015, 
                                    olddf = master_2010, 
                                    by = "playerID")
```

```{r new_2010thru2015_playerIDs}
new_2010thru2015_playerIDs <- unique(new_2010thru2015$playerID)
# check length
length(new_2010thru2015_playerIDs)
```

Check to see if these names are the identical rows in the 2015 data

```{r identical-new_2010thru2015-new_2010thru2015_playerIDs}
identical(x = nrow(new_2010thru2015), 
  y = nrow(filter(master_2015, playerID %in% new_2010thru2015_playerIDs)))
```

## New data between 2015 - 2020

```{r new_2015thru2020}
new_2015thru2020 <- create_new_data(newdf = people_2020, 
                                    olddf = master_2015, 
                                    by = "playerID")
```

```{r new_2015thru2020_playerIDs}
new_2015thru2020_playerIDs <- unique(new_2015thru2020$playerID)
length(new_2015thru2020_playerIDs)
```

Check to see if these names are the identical rows in the 2020 data

```{r identical-new_2015thru2020-new_2015thru2020_playerIDs}
identical(x = nrow(new_2015thru2020), 
  y = nrow(filter(people_2020, playerID %in% new_2015thru2020_playerIDs)))
```

# Deleted Data 

Now that we've covered the creation of 'new data', we can move onto the 'deleted data'. 

## Deleted data: "*What rows were here before that aren't here now?*"

For the deleted datasets, we'll be using the `create_deleted_data()` function.  

```{r create_deleted_data-args, eval=FALSE}
dfdiffs::create_deleted_data(newdf = , olddf = , by = , by_col = , cols = )
```

## Deleted data between 2010 - 2020

We can test to see if any data have been removed between the 2020 and 2010 'Master/People' datasets.

```{r deleted_2010thru2020}
deleted_2010thru2020 <- create_deleted_data(
  newdf = people_2020, 
  olddf = master_2010, 
  by = "playerID")
```

Now if we check the `playerID`s from the `deleted_2010thru2020` data, they shouldn't be in the `people_2020` data. 

```{r identical-people_2020-deleted_2010thru2020_playerIDs}
deleted_2010thru2020_playerIDs <- unique(deleted_2010thru2020$playerID)
deleted_2010thru2020_playerIDs
```

And the number of IDs in the `master_2010` data should be identical to the number of rows in the `deleted_2010thru2020` dataset. 

```{r identical-deleted_2010thru2020-deleted_2010thru2020_playerIDs}
identical(
  x = nrow(filter(master_2010, playerID %in% deleted_2010thru2020_playerIDs)),
  y = nrow(deleted_2010thru2020))
```

## Deleted data between 2010 - 2015

We'll also test to see if any data have been removed between the 2015 and 2010 'Master/People' datasets.

```{r deleted_2010thru2015}
deleted_2010thru2015 <- create_deleted_data(
  newdf = master_2015, 
  olddf = master_2010, 
  by = "playerID")
# get unique ids
deleted_2010thru2015_playerIDs <- unique(deleted_2010thru2015$playerID)
deleted_2010thru2015_playerIDs
```
```{r identical-deleted_2010thru2015-deleted_2010thru2015_playerIDs}
# check identical
identical(
  x = nrow(filter(master_2010, playerID %in% deleted_2010thru2015_playerIDs)),
  y = nrow(deleted_2010thru2015))
```

## Deleted data between 2015 - 2020

```{r deleted_2015thru2020}
deleted_2015thru2020 <- create_deleted_data(
  newdf = people_2020, 
  olddf = master_2015, 
  by = "playerID")
# get unique ids
deleted_2015thru2020_playerIDs <- unique(deleted_2015thru2020$playerID)
deleted_2015thru2020_playerIDs
```


```{r identical-deleted_2015thru2020-deleted_2015thru2020_playerIDs}
# check identical
identical(
  x = nrow(filter(master_2015, playerID %in% deleted_2015thru2020_playerIDs)),
  y = nrow(deleted_2015thru2020))
```

## Modified data: "*What values have been changed?*"

We now have functions for returning new records and deleted records, but what about records that have been modified? Both `create_new_data()` and `create_deleted_data()` are built using `dplyr` joins, but to find records where specific values have changed row-by-row, we need to enlist the help of a few functions from other packages. 


Other functions for comparing datasets to each other are used in the `dfdiffs` package, and these have all been rewritten to return a `tibble`. 


