---
title: "dfdiffs::create_new_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dfdiffs::create_new_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

The goal of the `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   
4. What columns have been moved?    
5. What rows have been moved?    

This vignette takes us through the `create_new_data()` function, which answers the "*What rows are here now that weren't here before?*"

## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(shiny)
library(data.table)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## What rows are here now that weren't here before? 

We're going to use two test datasets to demonstrate the `dfdiffs::create_new_data()` function. 

```{r Original}
T1Data <- dfdiffs::T1Data
T1Data
```


```{r T2Data}
T2Data <- dfdiffs::T2Data
T2Data
```

We want a function that allows users to specify a joining variable, which is created by supplying the columns that constitute a unique identifier in each dataset. 

For example, in `T1Data` and `T2Data`, if the unique identifier is the combination of `subject` and `record`, we want to be able to create a new column from these columns. We can do that with `create_new_column()`. 

```{r create_new_column}
  create_new_column <- function(data, cols, new_name) {
    new_col_data <- data %>%
      tidyr::unite({{new_name}}, {{cols}}, remove = FALSE, sep = "-") %>%
      dplyr::relocate({{new_name}}, everything())
    return(new_col_data)
  }
T1DataJoin <- create_new_column(data = T1Data, 
  cols = c("subject", "record"), 
  new_name = "join_var")
T2DataJoin <- create_new_column(data = T2Data, 
  cols = c("subject", "record"), 
  new_name = "join_var")
```



