---
title: "(dev) uploadData module"
author: "Martin Frigaard"
output:
  html_document: 
    highlight: espresso
    theme: darkly
    toc: true
    toc_float: yes
    df_print: paged
    code_folding: hide
    
runtime: shiny
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(fs)
library(reactable)
library(rmarkdown)
library(shiny)
library(shinythemes)
library(bs4Dash)
# Global options
knitr::opts_chunk$set(
  # cache = FALSE # cache data
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "tiny", # size of the text
  fig.path = "images", # location of figure files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
# reactable.theme ----
options(reactable.theme = reactableTheme(
  color = "hsl(233, 9%, 87%)",
  backgroundColor = "hsl(233, 9%, 19%)",
  borderColor = "hsl(233, 9%, 22%)",
  stripedColor = "hsl(233, 12%, 22%)",
  highlightColor = "hsl(233, 12%, 24%)",
  inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
))
```

# Purpose

This is a development document (written in `rmarkdown` and deployed with `shiny`) for the `dfdiffs` package. The purpose of the dfdiffs package is to perform  comparisons between two datasets (similar to `PROC COMPARE` in SAS). 

In this file, we'll walk through the development  of the `uploadDataDemo`, which is a shiny module for uploading flat files *or* excel files with multiple sheets into a shiny app. 

# Helpers

The application depends on the functions stored in the `helpers.R` file. 

```{r helpers, class.source = 'fold-hide'}
#' Load flat data files
#'
#' @param path path to data file (with extension)
#'
#' @return return_data
#' @export load_flat_file
#' @importFrom data.table fread
#' @importFrom haven read_sas
#' @importFrom haven read_sav
#' @importFrom haven read_dta
#' @importFrom tools file_ext
#' @importFrom tibble as_tibble
#'
#' @examples # from local
#' load_flat_file(path = "inst/extdata/csv/2015-baseballdatabank/core/AllstarFull.csv")
#'
load_flat_file <- function(path) {
  ext <- tools::file_ext(path)
  data <- switch(ext,
    txt = data.table::fread(path),
    csv = data.table::fread(path),
    tsv = data.table::fread(path),
    sas7bdat = haven::read_sas(data_file = path),
    sas7bcat = haven::read_sas(data_file = path),
    sav = haven::read_sav(file = path),
    dta = haven::read_dta(file = path)
  )
  return_data <- tibble::as_tibble(data)
  return(return_data)
}

#' Upload data to app `upload_data()`
#'
#' @param path path to data file (with extension)
#' @param sheet excel sheet (if excel file)
#'
#' @return uploaded
#' @export upload_data
#' @importFrom readxl read_excel
#' @importFrom tools file_ext
#' @importFrom tibble as_tibble
#'
#'
#' @examples # not run
#' upload_data(path = "inst/extdata/app-testing/lahman_compare.xlsx",
#'             sheet = "master-2015")
#' upload_data(path = "inst/extdata/app-testing/m15.csv")
#' upload_data(path = "inst/extdata/dta/iris.dta")
#' upload_data(path = "inst/extdata/sas7bdat/iris.sas7bdat")
#' upload_data(path = "inst/extdata/sav/iris.sav")
#' upload_data(path = "inst/extdata/tsv/Batting.tsv")
#' upload_data(path = "inst/extdata/txt/Batting.txt")
upload_data <- function(path, sheet = NULL) {
  ext <- tools::file_ext(path)
  if (ext == "xlsx") {
    raw_data <- readxl::read_excel(
        path = path,
        sheet = sheet
      )
    uploaded <- tibble::as_tibble(raw_data)
  } else {
    uploaded <- load_flat_file(path = path)
  }
  return(uploaded)
}


# bmrn_fresh_theme --------------------------------------------------------
bmrn_fresh_theme <- function() {
  fresh::create_theme(
    # theme vars  -------------------------------------------------------------
    fresh::bs4dash_vars(
      navbar_light_color = "#353d98", # purple
      navbar_light_active_color = "#353d98", # purple
      navbar_light_hover_color = "#f26631" # orange
    ),
    # # theme yiq -------------------------------------------------------------
    fresh::bs4dash_yiq(
      contrasted_threshold = 255,
      text_dark = "#0a0a0a", # dark_gray_s10
      text_light = "#f5f5f5" # gray_t10
    ),
    # theme layout ---------------------------------------------------------
    fresh::bs4dash_layout(
      main_bg = NULL, # #ececec
      font_size_root = 12
    ),
    # theme sidebar_light -------------------------------------------------
    fresh::bs4dash_sidebar_light(
      header_color = "#ccd5dd", # light blue
      bg = "#eaebf4", # background of entire side-bar
      color = "#002E56", # text color (no hover)
      hover_color = "#ee304e", # text color on hover
      hover_bg = "#353D98", # color on hover
      active_color = "#f26631", # color is actually the 'primary' status color
      submenu_bg = "#f5f5f5", # purple
      submenu_color = "#002444",
      submenu_hover_color = "#353D98" # purple
    ),
    # # theme sidebar_dark -------------------------------------------------
    fresh::bs4dash_sidebar_dark(
      header_color = "#ccd5dd",
      bg = "#1a1e4c",
      color = "#EE304E", # text color (no hover)
      hover_bg = "#aeb1d5", # color on hover
      hover_color = "#EE304E", # text color on hover
      active_color = "#f26631" # color is actually the 'primary' status color
    ),
    # theme status -------------------------------------------------
    fresh::bs4dash_status(
      dark = "#323232",
      light = "#A0A0A0", # gray
      warning = "#F26631", # orange
      primary = "#A9218E", # violet = #A9218E, blue = #00509C
      secondary = "#353D98", # purple
      success = "#00509C", # blue
      danger = "#EE304E", # red
      info = "#A0A0A0" # gray
    ),
    # theme color -------------------------------------------------
    fresh::bs4dash_color(
      gray_900 = "#1f245b",
      gray_800 = "#646464",
      lightblue = "#6696c3",
      blue = "#00509C"
    )
  )
}
# base_react_theme --------------------------------------------------------
base_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#761763",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#a9218e",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
# compare_react_theme -----------------------------------------------------
comp_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#2f3688",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#353d98",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
```


# Module part 1) UI

In our `uploadDataUI()`, we want the UI to clearly display three actions: 
1) Upload a data file  

  1a) If the imported data are excel files (`.xlsx`), display the sheets in the UI  
  1b) If the data files are `.sas7bdat`, `.csv`, `.txt`, `.tsv`, 
  
2) Upload a flat file with the `upload_data()` function (stored in the `helpers.R` file)

```{r upload_data, echo=FALSE}
upload_data <- function(path, sheet = NULL) {
    # define load flat file
    load_flat_file <- function(path) {
      ext <- tools::file_ext(path)
      data <- switch(ext,
        txt = data.table::fread(path),
        csv = data.table::fread(path),
        tsv = data.table::fread(path),
        sas7bdat = haven::read_sas(data_file = path),
        sas7bcat = haven::read_sas(data_file = path),
        sav = haven::read_sav(file = path),
        dta = haven::read_dta(file = path)
      )
      return_data <- tibble::as_tibble(data)
      return(return_data)
    }
    # define upload 
  ext <- tools::file_ext(path)
  
  if (ext == "xlsx") {
    raw_data <- readxl::read_excel(
        path = path,
        sheet = sheet
      )
    uploaded <- tibble::as_tibble(raw_data)
  } else {
    # use load flat file
    uploaded <- load_flat_file(path = path)
  }
  return(uploaded)
}
```


3) Enter a name for the uploaded data files    

After uploading and naming the data sources, the data are visible to preview as a `reactable()` table. 

The input IDs in this module have the following naming conventions: 

- `base` are the objects associated with the base (target) data. 

- `comp` are the objects associated with the compare (current) data. 

## `base` (target) objects 

If the user uploads a file, the following inputs are exported as a list.

1. **`base_data`**: the `base` file uploaded with the `fileInput()`

2. **`base_name`**: the new name designated for the `base` uploaded file


## `comp` (compare) objects 

If the user uploads a file, the following inputs are exported as a list.

3. **`comp_data`**: the `compare` file uploaded with the `fileInput()`

4. **`comp_name`**: the new name designated for the `compare` uploaded file

### uploadDataUI() 

```{r uploadDataUI}
uploadDataUI <- function(id) {
  tagList(
    h3("Upload a ", strong("base"), " (i.e., target) data source "),
    fluidRow(
      sortable(
        width = 12,
        # |- upload base xlsx file ----
        box(
          maximizable = TRUE,
          collapsible = TRUE,
          collapsed = FALSE,
          closable = FALSE,
          solidHeader = TRUE,
          status = "primary",
          width = 12,
          title = tags$strong("Upload File (base)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [base_file] -------
                inputId = NS(
                  namespace = id,
                  id = "base_file"
                ),
                label = tags$strong(
                  "Accepts: ",
                  code(".sas7bdat"), code(".csv"),
                  code(".txt"), code(".tsv"), code(".xlsx")),
                  accept = c(".sas7bdat", ".csv", ".txt", ".tsv", ".xlsx")
              )
            ),
            column(
              width = 6,
              ### |-- INPUT [base_xlsx_sheets] ---------
              selectInput(
                inputId = NS(
                  namespace = id,
                  id = "base_xlsx_sheets"
                ),
                label = strong("Select sheet (if ", code(".xlsx"), " file):"),
                choices = c("", NULL)
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [base_filename] ---------
              tags$strong("Data file name:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "base_filename"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [base_new_name] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "base_new_name"
                ),
                label = strong(
                  "Provide a name to preview the", code("base"), " file:"
                )
              ),
              em("Not sure what name to use? Copy + paste the file name."),
            )
          ),
          fluidRow(
            column(
              width = 12,
              br(), br(),
              ## |-- OUTPUT [base_display_upload] ---------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "base_display_upload"
                )
              )
            )
          )
        )
      )
    ),
    ## DEV (base) -----
    fluidRow(
      sortable(
        width = 12,
        box(
          width = 12,
          status = "info",
          solidHeader = TRUE,
          closable = TRUE,
          maximizable = TRUE,
          collapsed = TRUE,
          title = "Reactive values (base)",
          strong(em("For DEV purposes only")),
          fluidRow(
            column(12,
              ## base_dev_a -----
              code("base_dev_a"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "base_dev_a"
            )
          ))),
          fluidRow(
            column(12,
              ## base_dev_b -----
              code("base_dev_b"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "base_dev_b"
            )
          ))
          ),
          fluidRow(
            column(12,
              ## base_dev_x -----
              code("base_dev_x"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "base_dev_x"
            )
          ))),
          fluidRow(
            column(12,
              ## base_dev_y -----
              code("base_dev_y"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "base_dev_y"
            )
          ))
          )
        )
      )
    ),
    h3("Upload a ", strong("compare"), " (i.e., current) data source"),
    # br(), br(),
    fluidRow(
      sortable(
        width = 12,
        # |- upload compare xlsx file ----
        box(
          maximizable = TRUE,
          collapsed = FALSE,
          solidHeader = TRUE,
          status = "secondary",
          width = 12,
          collapsible = TRUE,
          closable = FALSE,
          title = tags$strong("Upload File (compare)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [comp_file] -------
                inputId = NS(
                  namespace = id,
                  id = "comp_file"
                ),
                label = tags$strong(
                  "Accepts: ",
                  code(".sas7bdat"), code(".csv"),
                  code(".txt"), code(".tsv"), code(".xlsx")),
                  accept = c(".sas7bdat", ".csv", ".txt", ".tsv", ".xlsx")
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [comp_xlsx_sheets] ---------
              selectInput(
                inputId = NS(
                  namespace = id,
                  id = "comp_xlsx_sheets"
                ),
                label = strong("Select sheet (if ", code(".xlsx"), " file):"),
                choices = c("", NULL)
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [comp_filename] ---------
              tags$strong("Data file name:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "comp_filename"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [comp_new_name] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "comp_new_name"
                ),
                label = strong(
                  "Provide a name to preview the ", code("compare"), " file:"
                )
              ),
              em("Not sure what name to use? Copy + paste the file name."),
            )
          ),
          fluidRow(
            column(
              width = 12,
              br(), br(),
              ## |-- OUTPUT [comp_display_upload] ---------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "comp_display_upload"
                )
              )
            )
          )
        )
      )
    ),
    ## DEV -----
    fluidRow(
      sortable(
        width = 12,
        box(
          width = 12,
          status = "info",
          solidHeader = TRUE,
          closable = TRUE,
          maximizable = TRUE,
          collapsed = TRUE,
          title = "Reactive values (compare)",
          strong(em("For DEV purposes only")),
          fluidRow(
            column(12,
              ## comp_dev_a -----
              code("comp_dev_a"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "comp_dev_a"
            )
          ))),
          fluidRow(
            column(12,
              ## comp_dev_b -----
              code("comp_dev_b"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "comp_dev_b"
            )
          ))
          ),
          fluidRow(
            column(12,
              ## comp_dev_x -----
              code("comp_dev_x"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "comp_dev_x"
            )
          ))),
          fluidRow(
            column(12,
              ## comp_dev_y -----
              code("comp_dev_y"),
          verbatimTextOutput(
            outputId = NS(
              namespace = id,
              id = "comp_dev_y"
            )
          ))
          )
        )
      )
    )
  )
}
```

# Module part 2) Server

The server needs to be capable of completing the following functions: 

1. Import the excel/flat files

  1a. Files are uploaded using the  `upload_data()` function  (stored in the `helpers.R` file) 
    
2. Create a display of the names of the imported files, and--if an excel file--display the sheets of the imported excel file  

3. After a value is provided in **`_new_name_base`**/**`_new_name_comp`**, the module will display the data 

## The return list

The uploaded items must be shared with the `selectData` module. All the objects follow a similar naming convention as the UI/Server, with a `base`/`comp` suffix. All of the objects stored and returned in the list are shown below:

```{r uploadDataServer-return-list, eval=FALSE, class.source = 'fold-show'}
    return(
      list(
        # |------ base_data ----
        base_data = reactive({
          req(input$base_file)
          req(input$base_new_name)
          if (nchar(input$base_xlsx_sheets) == 0) {
            uploaded <- upload_data(path = input$base_file$datapath)
          } else {
            uploaded <- upload_data(path = input$base_file$datapath,
                                  sheet = as.character(input$base_xlsx_sheets))
          }
          return(uploaded)
        }),
        # |------ base_name ----
        base_name = reactive({
          if (nchar(input$base_new_name) != 0) {
            as.character(input$base_new_name)
          } else {
            as.character(input$base_filename)
          }
        }),
        # |------ comp_data ----
        comp_data = reactive({
          req(input$comp_file)
          req(input$comp_new_name)
          if (nchar(input$comp_xlsx_sheets) == 0) {
            uploaded <- upload_data(path = input$comp_file$datapath)
          } else {
            uploaded <- upload_data(path = input$comp_file$datapath,
                                  sheet = as.character(input$comp_xlsx_sheets))
          }
          return(uploaded)
        }),
        # |------ comp_name ----
        comp_name = reactive({
          if (nchar(input$comp_new_name) != 0) {
            as.character(input$comp_new_name)
          } else {
            as.character(input$base_filename)
          }
        })
      )
    )
```

## uploadDataServer()

```{r uploadDataServer}
uploadDataServer <- function(id) {

  moduleServer(id = id, module = function(input, output, session) {

    # |-- INPUT [base] base_xlsx_sheets -----
    observeEvent(eventExpr = input$base_file, handlerExpr = {
      if (tools::file_ext(input$base_file$name) == "xlsx") {
        choices <- readxl::excel_sheets(path = input$base_file$datapath)
      } else {
        choices <- c("", NULL)
      }
      updateSelectInput(session = session,
          inputId = "base_xlsx_sheets",
          choices = choices
          )
    })

    # |-- OUTPUT [base] xlsx file name -----
    output$base_filename <- renderPrint({
      req(input$base_file)
      base_filename <- as.character(input$base_file$name)
      paste0(
        tags$code(base_filename)
      )
    })

    base_data <- eventReactive(input$base_file, {
        if (nchar(input$base_xlsx_sheets) == 0) {
          uploaded <- upload_data(path = input$base_file$datapath)
        } else {
          uploaded <- upload_data(path = input$base_file$datapath,
                                  sheet = as.character(input$base_xlsx_sheets))
        }
        return(uploaded)
    })

    # |-- OUTPUT display [base] xlsx ----
    # require name
    observeEvent(eventExpr = input$base_new_name, handlerExpr = {
      req(input$base_file)
      req(input$base_new_name)
      output$base_display_upload <- reactable::renderReactable(
        reactable(
          data = base_data(),
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = base_react_theme
        )
      )
    })

      ## DEV OUTPUT |-- (base_dev_a) ---------
      output$base_dev_a <- renderPrint({
        print(
          paste0("input$base_filename = ", input$base_file$name)
          )
      })
      ## DEV OUTPUT |-- (base_dev_b) ---------
      output$base_dev_b <- renderPrint({
        print(
          base_data()
          )
      })
      ## DEV OUTPUT |-- (base_dev_x) ---------
      output$base_dev_x <- renderPrint({
        print(
          paste0("input$base_xlsx_sheets = ",  as.character(input$base_xlsx_sheets))
          )
      })
      ## DEV OUTPUT |-- (base_dev_y) ---------
      output$base_dev_y <- renderPrint({
        print(
          paste0("input$base_new_name = ",  as.character(input$base_new_name))
          )
      })


    # |-- INPUT [comp] comp_xlsx_sheets -----
    observeEvent(eventExpr = input$comp_file, handlerExpr = {
      if (tools::file_ext(input$comp_file$name) == "xlsx") {
        choices <- readxl::excel_sheets(path = input$comp_file$datapath)
      } else {
        choices <- c("", NULL)
      }
      updateSelectInput(session = session,
          inputId = "comp_xlsx_sheets",
          choices = choices
          )
    })

    # |-- OUTPUT [comp] xlsx file name -----
    output$comp_filename <- renderPrint({
      req(input$comp_file)
      comp_filename <- as.character(input$comp_file$name)
      paste0(
        tags$code(comp_filename)
      )
    })

    comp_data <- eventReactive(input$comp_file, {
        if (nchar(input$comp_xlsx_sheets) == 0) {
          uploaded <- upload_data(path = input$comp_file$datapath)
        } else {
          uploaded <- upload_data(path = input$comp_file$datapath,
                                  sheet = as.character(input$comp_xlsx_sheets))
        }
        return(uploaded)
    })

   # |-- OUTPUT display [comp] xlsx ----
    # require name
    observeEvent(eventExpr = input$comp_new_name, handlerExpr = {
      req(input$comp_file)
      req(input$comp_new_name)
      output$comp_display_upload <- reactable::renderReactable(
        reactable(
          data = comp_data(),
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = comp_react_theme
        )
      )
    })

      ## DEV OUTPUT |-- (comp_dev_a) ---------
      output$comp_dev_a <- renderPrint({
        print(
          paste0("input$comp_filename = ",  as.character(input$comp_file$name))
          )
      })
      ## DEV OUTPUT |-- comp_dev_y (dev) ---------
      output$comp_dev_b <- renderPrint({
        print(
          comp_data()
          )
      })
      ## DEV OUTPUT |-- comp_dev_a (dev) ---------
      output$comp_dev_x <- renderPrint({
        print(
          paste0("input$comp_xlsx_sheets = ",  as.character(input$comp_xlsx_sheets))
          )
      })
      ## DEV OUTPUT |-- comp_dev_b (dev) ---------
      output$comp_dev_y <- renderPrint({
        print(
          paste0("input$comp_new_name = ",  as.character(input$comp_new_name))
          )
      })

    # |---- return list -----
    # assign this as 'upload_data_list'
    return(
      list(
        # |------ base_data ----
        base_data = reactive({
          req(input$base_file)
          req(input$base_new_name)
          if (nchar(input$base_xlsx_sheets) == 0) {
            uploaded <- upload_data(path = input$base_file$datapath)
          } else {
            uploaded <- upload_data(path = input$base_file$datapath,
                                  sheet = as.character(input$base_xlsx_sheets))
          }
          return(uploaded)
        }),
        # |------ base_name ----
        base_name = reactive({
          # req(input$base_new_name)
          if (nchar(input$base_new_name) != 0) {
            as.character(input$base_new_name)
          } else {
            as.character(input$base_filename)
          }
        }),
        # |------ comp_data ----
        comp_data = reactive({
          req(input$comp_file)
          req(input$comp_new_name)
          if (nchar(input$comp_xlsx_sheets) == 0) {
            uploaded <- upload_data(path = input$comp_file$datapath)
          } else {
            uploaded <- upload_data(path = input$comp_file$datapath,
                                  sheet = as.character(input$comp_xlsx_sheets))
          }
          return(uploaded)
        }),
        # |------ comp_name ----
        comp_name = reactive({
          # req(input$comp_new_name)
          if (nchar(input$comp_new_name) != 0) {
            as.character(input$comp_new_name)
          } else {
            as.character(input$base_filename)
          }
        })
      )
    )

  })
}
```

## Uploaded data display 

Both base and compare datasets will be differentiated by color in the application using the `reactable::reactableTheme()` function.

### Base display

The base color will use the `#761763` background color (based on the bmrn violet (`#a9218e`) :


```{r base_react_theme}
# base_react_theme ------
base_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#761763",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#a9218e",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
```

### Compare display

The compare color will use the `#2f3688` background color (based on the bmrn purple (`#353d98`) :

```{r compare_react_theme}
# comp_react_theme ------------
comp_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#2f3688",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#353d98",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
```

## App theme 

Below we define the app theme using **`fresh::create_theme()`**:

```{r bmrn_fresh_theme}
bmrn_fresh_theme <- function() {
  fresh::create_theme(
    # theme vars  -------------------------------------------------------------
    fresh::bs4dash_vars(
      navbar_light_color = "#353d98", # purple
      navbar_light_active_color = "#353d98", # purple
      navbar_light_hover_color = "#f26631" # orange
    ),
    # # theme yiq -------------------------------------------------------------
    fresh::bs4dash_yiq(
      contrasted_threshold = 255,
      text_dark = "#0a0a0a", # dark_gray_s10
      text_light = "#f5f5f5" # gray_t10
    ),
    # theme layout ---------------------------------------------------------
    fresh::bs4dash_layout(
      main_bg = NULL, # #ececec
      font_size_root = 12
    ),
    # theme sidebar_light -------------------------------------------------
    fresh::bs4dash_sidebar_light(
      header_color = "#ccd5dd", # light blue
      bg = "#eaebf4", # background of entire side-bar
      color = "#002E56", # text color (no hover)
      hover_color = "#ee304e", # text color on hover
      hover_bg = "#353D98", # color on hover
      active_color = "#f26631", # color is actually the 'primary' status color
      submenu_bg = "#f5f5f5", # purple
      submenu_color = "#002444",
      submenu_hover_color = "#353D98" # purple
    ),
    # # theme sidebar_dark -------------------------------------------------
    fresh::bs4dash_sidebar_dark(
      header_color = "#ccd5dd",
      bg = "#1a1e4c",
      color = "#EE304E", # text color (no hover)
      hover_bg = "#aeb1d5", # color on hover
      hover_color = "#EE304E", # text color on hover
      active_color = "#f26631" # color is actually the 'primary' status color
    ),
    # theme status -------------------------------------------------
    fresh::bs4dash_status(
      dark = "#323232",
      light = "#A0A0A0", # gray
      warning = "#F26631", # orange
      primary = "#A9218E", # violet = #A9218E, blue = #00509C
      secondary = "#353D98", # purple
      success = "#00509C", # blue
      danger = "#EE304E", # red
      info = "#A0A0A0" # gray
    ),
    # theme color -------------------------------------------------
    fresh::bs4dash_color(
      gray_900 = "#1f245b",
      gray_800 = "#646464",
      lightblue = "#6696c3",
      blue = "#00509C"
    )
  )
}
```

# Module part 3) Demo

We define `uploadDataDemo()`, a function for running our application:

```{r uploadDataDemo}
# set theme ---------------------------------------------------------------
upload_theme <- bmrn_fresh_theme()

uploadDataDemo <- function() {
  ui <- bs4Dash::dashboardPage(
    freshTheme = upload_theme,
    dark = FALSE,
    title = "uploadDataDemo",
    header = dashboardHeader(title = "uploadDataDemo"),
    # sidebar (menuItem) -------------------
    sidebar = dashboardSidebar(
      sidebarMenu(
        id = "sidebarmenu",
        sidebarHeader("Data upload demo"),
        menuItem("1) Upload Data",
          tabName = "upload_data_tab",
          icon = icon("file")
        )
      )
    ),
    # dashboardBody (tabItem) --------------
    body = dashboardBody(
      tabItems(
        tabItem(
          tabName = "upload_data_tab",
          ## uploadDataUI -----
          uploadDataUI(id = "upload_data")
        )
      )
    ),
    controlbar = dashboardControlbar(),
    footer = dashboardFooter()
  )

  server <- function(input, output, session) {
    uploadDataServer(id = "upload_data")
  }

  shinyApp(
    ui = ui, server = server,
    options = list(width = 800, height = 1000)
  )
}
```

# App

You can demo this by running `uploadDataDemo()`.

## Testing

Files for testing this application can be found here:

1. [base](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/csv/2015-baseballdatabank/Master.csv): a Master [Lahman dataset](https://www.seanlahman.com/baseball-archive/statistics/) from 2015

2. [compare](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/csv/2020-baseballdatabank/People.csv): a Master [Lahman dataset](https://www.seanlahman.com/baseball-archive/statistics/) from 2020

3. Also use [this .xlsx file](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/xlsx/lahman_compare.xlsx) with the same datasets. 

## Reactive values 

Two `sortable()` boxes are displayed in the UI (`Reactive values (base)` and `Reactive values (compare)`). These are for development purposes only and will not be displayed in the final application. 

```{r run-uploadDataDemo}
uploadDataDemo()
```
