---
title: "uploadDataDemo(): upload module (dev)"
author: "Martin Frigaard"
output:
  html_document: 
    highlight: espresso
    theme: darkly
    toc: true
    toc_float: yes
    df_print: paged
    code_folding: hide
    
runtime: shiny
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(fs)
library(reactable)
library(rmarkdown)
library(shiny)
library(shinythemes)
library(bs4Dash)
# Global options
knitr::opts_chunk$set(
  # cache = FALSE # cache data
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "tiny", # size of the text
  fig.path = "images", # location of figure files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
# reactable.theme ----
options(reactable.theme = reactableTheme(
  color = "hsl(233, 9%, 87%)",
  backgroundColor = "hsl(233, 9%, 19%)",
  borderColor = "hsl(233, 9%, 22%)",
  stripedColor = "hsl(233, 12%, 22%)",
  highlightColor = "hsl(233, 12%, 24%)",
  inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
))
```

# Purpose

This is a development document (written in `rmarkdown` and deployed with `shiny`) for the `dfdiffs` package. The purpose of the dfdiffs package is to perform  comparisons between two datasets (similar to `PROC COMPARE` in SAS). 

In this file, we'll walk through the development  of the `uploadDataDemo`, which is a shiny module for uploading flat files *or* excel files with multiple sheets into a shiny app. 

# Module part 1) UI

In our `uploadDataUI()`, we want the UI to clearly display three actions: 

1) Upload a flat/excel file  
  1a) If the imported data are excel files, select the sheets to import  
  
2) Upload a flat file with the `load_flat_file()` function  
2) Enter a name for the uploaded data files    

After uploading and naming the data sources, the data are visible to preview as a `reactable()` table. 

The input IDs in this module have the following naming conventions: 

- `base` are the objects associated with the base (target) data. 

- `comp` are the objects associated with the compare (current) data. 

## Excel objects 

If the user uploads an excel file, the following inputs are used.

1. **`xlsx_file_base`**/**`xlsx_file_comp`**: the file uploaded with the `fileInput()`

2. **`xlsx_sheets_base`**/**`xlsx_sheets_comp`**: the sheets in the uploaded excel file (displayed in a dropdown)

3. **`xlsx_filename_base`**/**`xlsx_filename_comp`**: the filename for the data uploaded in `xlsx_file_base`/`xlsx_file_comp`. 

4. **`xlsx_new_name_base`**/**`xlsx_new_name_comp`**: the new name designated for the uploaded excel file

5. After entering a value in `xlsx_new_name_base`/`xlsx_new_name_comp`, the data are displayed as **`xlsx_upload_base`**/**`xlsx_upload_comp`** in a `reactableOutput()`  

## Flat file objects 

If the user uploads a flat file (`.sas7bdat`, `.csv`, etc.), the following inptus are used: 

1. **`flat_file_base`**/**`flat_file_comp`**: the file uploaded with the `fileInput()` using the `load_flat_file()`  

2. **`flat_filename_base`**/**`flat_filename_comp`**: the original filename of the uploaded data file in `flat_file_base`/`flat_file_comp`.

3. **`flat_file_new_name_base`**/**`flat_file_new_name_comp`**: the new name designated for the uploaded flat file 

4. **`flat_file_upload_base`**/**`flat_file_upload_comp`**: After entering a value in `flat_file_new_name_base`/`flat_file_new_name_comp`, the data are displayed as `xlsx_upload_base`/`xlsx_upload_comp` in a `reactableOutput()`  

### uploadDataUI() 

```{r uploadDataUI}
#' data upload (UI) for selectDataDemo
#'
#' @param id module id
#'
#' @export uploadDataUI
#'
#' @description `uploadDataUI()`/`uploadDataServer()` create the upload module
#' for the dfdiffs app.
uploadDataUI <- function(id) {
  tagList(
    h3("Upload a ", strong("base"), " (target) data source "),
    fluidRow(
      sortable(
        width = 12,
        # |- upload base xlsx file ----
        box(
          maximizable = TRUE,
          collapsible = TRUE,
          collapsed = TRUE,
          closable = FALSE,
          solidHeader = TRUE,
          status = "primary",
          width = 12,
          title = tags$strong("Upload Excel File (base)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [xlsx_file_base] -------
                inputId = NS(
                  namespace = id,
                  id = "xlsx_file_base"
                ),
                label = "Excel file input",
                accept = c(".xlsx")
              )
            ),
            column(
              width = 6,
              ### |-- INPUT [xlsx_sheets_base] ---------
              selectInput(
                inputId = NS(
                  namespace = id,
                  id = "xlsx_sheets_base"
                ),
                label = "Select sheet:",
                choices = ""
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [xlsx_filename_base] ---------
              tags$strong("Excel data file:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "xlsx_filename_base"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [xlsx_new_name_base] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "xlsx_new_name_base"
                ),
                label = strong(
                  "Provide a name for the", code("base"), " excel fil:"
                )
              ),
            )
          ),
          fluidRow(
            column(
              width = 12,
              br(), br(),
              ## |-- OUTPUT [xlsx_upload_base] ---------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "xlsx_upload_base"
                )
              )
            )
          )
        )
      )
    ),
    # |- upload base flat file ----
    fluidRow(
      sortable(
        width = 12,
        box(
          maximizable = TRUE,
          collapsed = FALSE,
          collapsible = TRUE,
          closable = FALSE,
          solidHeader = TRUE,
          status = "primary",
          width = 12,
          title = tags$strong("Upload Flat Data File (base)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [flat_file_base] -------
                inputId = NS(
                  namespace = id,
                  id = "flat_file_base"
                ),
                label = tags$strong(
                  "Accepts: ",
                  code(".sas7bdat"), code(".csv"), code(".txt"), code(".tsv")
                ),
                accept = c(".sas7bdat", ".csv", ".txt", ".tsv")
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [flat_filename_base] ---------
              tags$strong("Flat file data:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "flat_filename_base"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [flat_file_new_name_base] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "flat_file_new_name_base"
                ),
                label = strong(
                  "Provide a name for the ", code("base"), " flat file:"
                )
              )
            )
          ),
          fluidRow(
            column(
              width = 12,
              ## |-- OUTPUT [flat_file_upload_base] -------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "flat_file_upload_base"
                )
              )
            )
          )
        )
      )
    ),
    h3("Upload a ", strong("compare"), " (current) data source"),
    # br(), br(),
    fluidRow(
      sortable(
        width = 12,
        # |- upload compare xlsx file ----
        box(
          maximizable = TRUE,
          collapsed = TRUE,
          solidHeader = TRUE,
          status = "secondary",
          width = 12,
          collapsible = TRUE,
          closable = FALSE,
          title = tags$strong("Upload Excel File (compare)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [xlsx_file_comp] -------
                inputId = NS(
                  namespace = id,
                  id = "xlsx_file_comp"
                ),
                label = "Excel file upload",
                accept = c(".xlsx")
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [xlsx_sheets_comp] ---------
              selectInput(
                inputId = NS(
                  namespace = id,
                  id = "xlsx_sheets_comp"
                ),
                label = "Select sheet:",
                choices = ""
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [xlsx_filename_comp] ---------
              tags$strong("Excel Data File:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "xlsx_filename_comp"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [xlsx_new_name_comp] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "xlsx_new_name_comp"
                ),
                label = strong(
                  "Provide a name for the ", code("compare"), " excel file :"
                )
              ),
            )
          ),
          fluidRow(
            column(
              width = 12,
              br(), br(),
              ## |-- OUTPUT [xlsx_upload_comp] ---------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "xlsx_upload_comp"
                )
              )
            )
          )
        )
      )
    ),
    # |- upload compare flat file ----
    fluidRow(
      sortable(
        width = 12,
        box(
          maximizable = TRUE,
          collapsed = FALSE,
          solidHeader = TRUE,
          status = "secondary",
          width = 12,
          collapsible = TRUE,
          closable = FALSE,
          title = tags$strong("Upload Flat Data File (compare)"),
          fluidRow(
            column(
              width = 6,
              fileInput(
                ## |-- INPUT [flat_file_comp] -------
                inputId = NS(
                  namespace = id,
                  id = "flat_file_comp"
                ),
                label = tags$strong(
                  "Accepts: ",
                  code(".sas7bdat"), code(".csv"), code(".txt"), code(".tsv")
                ),
                accept = c(".sas7bdat", ".csv", ".txt", ".tsv")
              )
            )
          ),
          fluidRow(
            column(
              width = 6,
              ## |-- OUTPUT [flat_filename_comp] ---------
              tags$strong("Flat file data:"),
              shiny::htmlOutput(
                outputId = NS(
                  namespace = id,
                  id = "flat_filename_comp"
                )
              )
            ),
            column(
              width = 6,
              ## |-- INPUT [flat_file_new_name_comp] ---------
              textInput(
                inputId = NS(
                  namespace = id,
                  id = "flat_file_new_name_comp"
                ),
                label = strong(
                  "Provide a name for the ", code("compare"), " flat file:"
                )
              )
            )
          ),
          fluidRow(
            column(
              width = 12,
              ## |-- OUTPUT [flat_file_upload_comp] -------
              reactable::reactableOutput(
                outputId = NS(
                  namespace = id,
                  id = "flat_file_upload_comp"
                )
              )
            )
          )
        )
      )
    )
  )
}
```

# Module part 2) Server

The server needs to be capable of completing the following functions: 

1. Import the excel/flat files

  1a. If the uploaded data are flat files, the `load_flat_file()` function is used: 
      
```{r load_flat_file, class.source = 'fold-show'}
load_flat_file <- function(path) {
  ext <- tools::file_ext(path)
  data <- switch(ext,
    txt = data.table::fread(path),
    csv = data.table::fread(path),
    tsv = data.table::fread(path),
    sas7bdat = haven::read_sas(data_file = path),
    sas7bcat = haven::read_sas(data_file = path),
    sav = haven::read_sav(file = path),
    dta = haven::read_dta(file = path)
  )
  return_data <- as_tibble(data)
  return(return_data)
}
```


2. Display the names of the imported files, and--if an excel file--display the sheets of the imported excel file  

3. After a value is provided in **`_new_name_base`**/**`_new_name_comp`**, the module will display the data 

## The return list

The uploaded items must be shared with the `selectData` module. All the objects follow a similar naming convention as the UI/Server, but instead of a suffix, the `base`/`comp` is added to these objects as a prefix.

All of the objects stored and returned in the list are shown below:

```{r uploadDataServer-return-list, eval=FALSE, class.source = 'fold-show'}
    # |---- return list -----
    # assign this as 'upload_data_list'
    return(
      list(
        # |------ base_xlsx_data ----
        base_xlsx_data = reactive({
          req(input$xlsx_file_base)
          req(input$xlsx_sheets_base)
          req(input$xlsx_new_name_base)
          worksheet_data <- readxl::read_excel(
            path = input$xlsx_file_base$datapath,
            sheet = input$xlsx_sheets_base
          )
        }),
        # |------ base_xlsx_data_name ----
        base_xlsx_data_name = reactive({
          # req(input$xlsx_new_name_base)
          if (length(input$xlsx_new_name_base) == 1) {
            as.character(input$xlsx_new_name_base)
          } else {
            NULL
          }
        }),
        # |------ base_flat_file_data ----
        base_flat_file_data = reactive({
          req(input$flat_file_base)
          req(input$flat_file_new_name_base)
          flat_file_base <- load_flat_file(
            path = input$flat_file_base$datapath
          )
        }),
        # |------ base_flat_file_data_name ----
        base_flat_file_data_name = reactive({
          # req(input$flat_file_new_name_base)
          if (length(input$flat_file_new_name_base) == 1) {
            as.character(input$flat_file_new_name_base)
          } else {
            NULL
          }
        }),

        # |------ comp_xlsx_data ----
        comp_xlsx_data = reactive({
          req(input$xlsx_file_comp)
          req(input$xlsx_sheets_comp)
          req(input$xlsx_new_name_comp)
          comp_xlsx_data <- readxl::read_excel(
            path = input$xlsx_file_comp$datapath,
            sheet = input$xlsx_sheets_comp
          )
        }),
        # |------ comp_xlsx_data_name ----
        comp_xlsx_data_name = reactive({
          # req(input$xlsx_new_name_comp)
          if (length(input$xlsx_new_name_comp) == 1) {
            as.character(input$xlsx_new_name_comp)
          } else {
            NULL
          }
        }),
        # |------ comp_flat_file_data ----
        comp_flat_file_data = reactive({
          req(input$flat_file_comp)
          req(input$flat_file_new_name_comp)
          comp_flat_file <- load_flat_file(
            path = input$flat_file_comp$datapath
          )
        }),
        # |------ comp_flat_file_data_name ----
        comp_flat_file_data_name = reactive({
          # req(input$flat_file_new_name_comp)
          if (length(input$flat_file_new_name_comp) == 1) {
            as.character(input$flat_file_new_name_comp)
          } else {
            NULL
          }
        })
      )
    )
```

## uploadDataServer()

```{r uploadDataServer}
# uploadDataServer --------------------------------------------------------
#' data upload (Server) for selectDataDemo()
#'
#' @param id module id
#'
#' @export uploadDataServer
#'
#' @return list with data inputs
#' \describe{
#'   \item{base_xlsx_data}{reactive dataset of imported 'base' excel file}
#'   \item{base_xlsx_data_name}{name of reactive dataset of imported 'base' excel file}
#'   \item{base_flat_file_data}{reactive dataset of imported 'base' flat file}
#'   \item{base_flat_file_data_name}{name of reactive dataset of imported 'base' flat file}
#'   \item{comp_xlsx_data}{reactive dataset of imported 'compare' excel file}
#'   \item{comp_xlsx_data_name}{name of reactive dataset of imported 'compare' excel file}
#'   \item{comp_flat_file_data}{reactive dataset of imported 'compare' flat file}
#'   \item{comp_flat_file_data_name}{name of reactive dataset of imported 'compare' flat file}
#' }
#'
#' @examples # build server using uploadDataServer() and displayDataServer()
#' upload_data_list <- uploadDataServer(id = "upload_data")
#' displayDataServer(id = "display_data", data_upload = upload_data_list)
uploadDataServer <- function(id) {
  moduleServer(id = id, module = function(input, output, session) {

    # |-- INPUT [base] xlsx sheets -----
    observeEvent(eventExpr = input$xlsx_file_base, handlerExpr = {
      if (is.null(input$xlsx_file_base)) {
        return(NULL)
      } else {
        xlsx_sheets <- readxl::excel_sheets(path = input$xlsx_file_base$datapath)
        updateSelectInput(session, "xlsx_sheets_base", choices = xlsx_sheets)
      }
    })

    # |-- OUTPUT [base] xlsx file name -----
    output$xlsx_filename_base <- renderPrint({
      req(input$xlsx_file_base)
      xlsx_filename_base <- as.character(input$xlsx_file_base$name)
      paste0(
        tags$code(xlsx_filename_base)
      )
    })

    # |-- OUTPUT display [base] xlsx ----
    # require name
    observeEvent(eventExpr = input$xlsx_new_name_base, handlerExpr = {
      req(input$xlsx_file_base)
      req(input$xlsx_sheets_base)
      req(input$xlsx_new_name_base)
      worksheet_data <- readxl::read_excel(
        path = input$xlsx_file_base$datapath,
        sheet = input$xlsx_sheets_base
      )
      worksheet_names_tbl <- tibble::as_tibble(worksheet_data)
      output$xlsx_upload_base <- reactable::renderReactable(
        reactable(
          data = worksheet_names_tbl,
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = base_react_theme
        )
      )
    })

    # |-- INPUT [base] flat file -----
    flat_file_base <- reactive({
      req(input$flat_file_base)
      req(input$flat_file_new_name_base)
      flat_file_base <- load_flat_file(
        path = input$flat_file_base$datapath
      )
      return(flat_file_base)
    })

    # |-- OUTPUT [base] flat file name -----
    output$flat_filename_base <- renderPrint({
      req(input$flat_file_base)
      flat_filename_base <- as.character(input$flat_file_base$name)
      paste0(
        tags$code(flat_filename_base)
      )
    })

    # |-- OUTPUT display [base] flat file -----
    observeEvent(eventExpr = input$flat_file_new_name_base, handlerExpr = {
      output$flat_file_upload_base <- reactable::renderReactable(
        reactable(
          data = flat_file_base(),
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = base_react_theme
        )
      )
    })

    # |-- [compare] xlsx sheets -----
    observeEvent(eventExpr = input$xlsx_file_comp, handlerExpr = {
      if (is.null(input$xlsx_file_comp)) {
        return(NULL)
      } else {
        xlsx_sheets <- readxl::excel_sheets(path = input$xlsx_file_comp$datapath)
        updateSelectInput(session, "xlsx_sheets_comp", choices = xlsx_sheets)
      }
    })

    # |-- [compare] xlsx filename name -----
    output$xlsx_filename_comp <- renderPrint({
      req(input$xlsx_file_comp)
      xlsx_filename_comp <- as.character(input$xlsx_file_comp$name)
      paste0(
        tags$code(xlsx_filename_comp)
      )
    })

    # |-- display [compare] xlsx ----
    observeEvent(eventExpr = input$xlsx_new_name_comp, handlerExpr = {
      req(input$xlsx_file_comp)
      req(input$xlsx_sheets_comp)
      req(input$xlsx_new_name_comp)
      worksheet_data <- readxl::read_excel(
        path = input$xlsx_file_comp$datapath,
        sheet = input$xlsx_sheets_comp
      )
      worksheet_names_tbl <- tibble::as_tibble(worksheet_data)
      output$xlsx_upload_comp <- reactable::renderReactable(
        reactable(
          data = worksheet_names_tbl,
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = compare_react_theme
        )
      )
    })

    # |-- import [compare] flat file -----
    flat_file_comp <- reactive({
      req(input$flat_file_comp)
      req(input$flat_file_new_name_comp)
      flat_file_comp <- load_flat_file(
        path = input$flat_file_comp$datapath
      )
      return(flat_file_comp)
    })

    # |-- [compare] flat file name -----
    output$flat_filename_comp <- renderPrint({
      req(input$flat_file_comp)
      flat_filename_comp <- as.character(input$flat_file_comp$name)
      paste0(
        tags$code(flat_filename_comp)
      )
    })

    # |-- display [compare] flat file -----
    observeEvent(eventExpr = input$flat_file_new_name_comp, handlerExpr = {
      output$flat_file_upload_comp <- reactable::renderReactable(
        reactable(
          data = flat_file_comp(),
          defaultPageSize = 5,
          resizable = TRUE,
          highlight = TRUE,
          compact = TRUE,
          wrap = FALSE,
          bordered = TRUE,
          filterable = TRUE,
          theme = compare_react_theme
        )
      )
    })

    # |---- return list -----
    # assign this as 'upload_data_list'
    return(
      list(
        # |------ base_xlsx_data ----
        base_xlsx_data = reactive({
          req(input$xlsx_file_base)
          req(input$xlsx_sheets_base)
          req(input$xlsx_new_name_base)
          worksheet_data <- readxl::read_excel(
            path = input$xlsx_file_base$datapath,
            sheet = input$xlsx_sheets_base
          )
        }),
        # |------ base_xlsx_data_name ----
        base_xlsx_data_name = reactive({
          # req(input$xlsx_new_name_base)
          if (length(input$xlsx_new_name_base) == 1) {
            as.character(input$xlsx_new_name_base)
          } else {
            NULL
          }
        }),
        # |------ base_flat_file_data ----
        base_flat_file_data = reactive({
          req(input$flat_file_base)
          req(input$flat_file_new_name_base)
          flat_file_base <- load_flat_file(
            path = input$flat_file_base$datapath
          )
        }),
        # |------ base_flat_file_data_name ----
        base_flat_file_data_name = reactive({
          # req(input$flat_file_new_name_base)
          if (length(input$flat_file_new_name_base) == 1) {
            as.character(input$flat_file_new_name_base)
          } else {
            NULL
          }
        }),

        # |------ comp_xlsx_data ----
        comp_xlsx_data = reactive({
          req(input$xlsx_file_comp)
          req(input$xlsx_sheets_comp)
          req(input$xlsx_new_name_comp)
          comp_xlsx_data <- readxl::read_excel(
            path = input$xlsx_file_comp$datapath,
            sheet = input$xlsx_sheets_comp
          )
        }),
        # |------ comp_xlsx_data_name ----
        comp_xlsx_data_name = reactive({
          # req(input$xlsx_new_name_comp)
          if (length(input$xlsx_new_name_comp) == 1) {
            as.character(input$xlsx_new_name_comp)
          } else {
            NULL
          }
        }),
        # |------ comp_flat_file_data ----
        comp_flat_file_data = reactive({
          req(input$flat_file_comp)
          req(input$flat_file_new_name_comp)
          comp_flat_file <- load_flat_file(
            path = input$flat_file_comp$datapath
          )
        }),
        # |------ comp_flat_file_data_name ----
        comp_flat_file_data_name = reactive({
          # req(input$flat_file_new_name_comp)
          if (length(input$flat_file_new_name_comp) == 1) {
            as.character(input$flat_file_new_name_comp)
          } else {
            NULL
          }
        })
      )
    )
  })
}
```

## Uploaded data display 

Both base and compare datasets will be differentiated by color in the application using the `reactable::reactableTheme()` function.

### Base display

The base color will use the `#761763` background color (based on the BioMarin violet (`#a9218e`) :


```{r base_react_theme}
base_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#761763",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#a9218e",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
```

### Compare display

The compare color will use the `#2f3688` background color (based on the BioMarin purple (`#353d98`) :

```{r compare_react_theme}
compare_react_theme <- reactableTheme(
          color = "#FFFFFF",
          backgroundColor = "#2f3688",
          borderColor = "#646464",
          stripedColor = "hsl(233, 12%, 22%)",
          highlightColor = "#353d98",
          inputStyle = list(backgroundColor = "#3A3B45"),
          selectStyle = list(backgroundColor = "#3A3B45"),
          pageButtonHoverStyle = list(backgroundColor = "3A3B45"),
          pageButtonActiveStyle = list(backgroundColor = "#3A3B45")
        )
```

## App theme 

Below we define the app theme using **`fresh::create_theme()`**:

```{r bmrn_fresh_theme}
# bmrn_fresh_theme --------------------------------------------------------
bmrn_fresh_theme <- function() {
  fresh::create_theme(
    # theme vars  -------------------------------------------------------------
    fresh::bs4dash_vars(
      navbar_light_color = "#353d98", # purple
      navbar_light_active_color = "#353d98", # purple
      navbar_light_hover_color = "#f26631" # orange
    ),
    # # theme yiq -------------------------------------------------------------
    fresh::bs4dash_yiq(
      contrasted_threshold = 255,
      text_dark = "#0a0a0a", # dark_gray_s10
      text_light = "#f5f5f5" # gray_t10
    ),
    # theme layout ---------------------------------------------------------
    fresh::bs4dash_layout(
      main_bg = NULL, # #ececec
      font_size_root = 12
    ),
    # theme sidebar_light -------------------------------------------------
    fresh::bs4dash_sidebar_light(
      header_color = "#ccd5dd", # light blue
      bg = "#eaebf4", # background of entire side-bar
      color = "#002E56", # text color (no hover)
      hover_color = "#ee304e", # text color on hover
      hover_bg = "#353D98", # color on hover
      active_color = "#f26631", # color is actually the 'primary' status color
      submenu_bg = "#f5f5f5", # purple
      submenu_color = "#002444",
      submenu_hover_color = "#353D98" # purple
    ),
    # # theme sidebar_dark -------------------------------------------------
    fresh::bs4dash_sidebar_dark(
      header_color = "#ccd5dd",
      bg = "#1a1e4c",
      color = "#EE304E", # text color (no hover)
      hover_bg = "#aeb1d5", # color on hover
      hover_color = "#EE304E", # text color on hover
      active_color = "#f26631" # color is actually the 'primary' status color
    ),
    # theme status -------------------------------------------------
    fresh::bs4dash_status(
      dark = "#323232",
      light = "#A0A0A0", # gray
      warning = "#F26631", # orange
      primary = "#A9218E", # violet = #A9218E, blue = #00509C
      secondary = "#353D98", # purple
      success = "#00509C", # blue
      danger = "#EE304E", # red
      info = "#A0A0A0" # gray
    ),
    # theme color -------------------------------------------------
    fresh::bs4dash_color(
      gray_900 = "#1f245b",
      gray_800 = "#646464",
      lightblue = "#6696c3",
      blue = "#00509C"
    )
  )
}

```

# Module part 3) Demo

We define `uploadDataDemo()`, a function for running our application:

```{r uploadDataDemo}
# set theme ---------------------------------------------------------------
upload_theme <- bmrn_fresh_theme()

uploadDataDemo <- function() {
  ui <- bs4Dash::dashboardPage(
    freshTheme = upload_theme,
    dark = FALSE,
    title = "uploadDataDemo",
    header = dashboardHeader(title = "uploadDataDemo"),
    # sidebar (menuItem) ------------------------------------------------------
    sidebar = dashboardSidebar(
      sidebarMenu(
        id = "sidebarmenu",
        sidebarHeader("Data upload demo"),
        menuItem("Upload Data",
          tabName = "upload_data_tab",
          icon = icon("file")
        )
      )
    ),
    # dashboardBody (tabItem) ------------------------------------------------------
    body = dashboardBody(
      tabItems(
        tabItem(
          tabName = "upload_data_tab",
          ## uploadDataUI -----
          uploadDataUI(id = "upload_data")
        )
      )
    ),
    controlbar = dashboardControlbar(),
    footer = dashboardFooter()
  )

  server <- function(input, output, session) {
    uploadDataServer(id = "upload_data")
  }

  shinyApp(
    ui = ui, server = server,
    options = list(width = 800, height = 1000)
  )
}
```

# App

You can demo this by running `uploadDataDemo()`.

## Testing

Files for testing this application can be found here:

1. [base](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/csv/2015-baseballdatabank/Master.csv): a Master [Lahman dataset](https://www.seanlahman.com/baseball-archive/statistics/) from 2015

2. [compare](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/csv/2020-baseballdatabank/People.csv): a Master [Lahman dataset](https://www.seanlahman.com/baseball-archive/statistics/) from 2020

3. Also use [this .xlsx file](https://github.com/mjfrigaard/dfdiffs/raw/main/inst/extdata/xlsx/lahman_compare.xlsx) with the same datasets. 

```{r run-uploadDataDemo}
uploadDataDemo()
```
