

```{r , eval=FALSE, echo=FALSE}
    # multiple 'by', no 'by_col', no 'cols' ----
    } else if (length(by) > 1 & is.null(by_col) & is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = "JOIN")
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = "JOIN")
    # check
    # mod_data <- list(newdf_join, olddf_join)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join,
                                y = olddf_join,
                                by = "JOIN"))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                    dplyr::contains("JOIN"),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # multiple 'by', new 'by_col', no 'cols' -----
  } else if (length(by) > 1 & !is.null(by_col) & is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = {{by_col}})
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = {{by_col}})
    # check
    # mod_data <- list(newdf_join, olddf_join)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join,
                                y = olddf_join,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # multiple 'by', 'by_col', multiple 'cols' ----
  } else if (length(by) > 1 & !is.null(by_col) & !is.null(cols)) {
    newdf_join <- create_new_column(data = newdf,
                        cols = all_of(by), new_name = {{by_col}})
    olddf_join <- create_new_column(data = olddf,
                        cols = all_of(by), new_name = {{by_col}})
    newdf_join_cols <- dplyr::select(newdf_join, {{by_col}}, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf_join, {{by_col}}, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # single 'by', no 'by_col', no 'cols' ----
  } else if (length(by) == 1 & is.null(by_col) & is.null(cols)) {
    newdf_join_cols <- dplyr::relocate(newdf, {{by}}, everything())
    olddf_join_cols <- dplyr::relocate(olddf, {{by}}, everything())
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # single 'by', 'by_col', no 'cols' ----
  } else if (length(by) == 1 & !is.null(by_col) & is.null(cols)) {
    names(newdf)[names(newdf) == by] <- by_col
    names(olddf)[names(olddf) == by] <- by_col
    newdf_join_cols <- newdf
    olddf_join_cols <- olddf
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # single 'by', 'by_col', multiple 'cols' -----
  } else if (length(by) == 1 & !is.null(by_col) & !is.null(cols)) {
    names(newdf)[names(newdf) == by] <- by_col
    names(olddf)[names(olddf) == by] <- by_col
    # newdf_join_cols <- newdf
    # olddf_join_cols <- olddf
    newdf_join_cols <- dplyr::select(newdf, {{by_col}}, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf, {{by_col}}, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols,
                                by = {{by_col}}))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                     all_of(by_col),
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)
    # no 'by', no 'by_col', multiple 'cols' ----
  } else if (is.null(by) & is.null(by_col) & !is.null(cols)) {
    newdf_join_cols <- dplyr::select(newdf, all_of(cols))
    olddf_join_cols <- dplyr::select(olddf, all_of(cols))
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols))
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                    `Variable name` = var.x,
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    mod_data <- tibble::as_tibble(diffs_names)

    # single 'by_col', no 'by', no 'cols' ----
  } else if (length(by_col) == 1 & is.null(by) & is.null(cols)) {
    # only 'by' provided
    newdf_join_cols <- dplyr::mutate(newdf,
                                        ID = row_number(),
                                        ID = as.character(ID))
    olddf_join_cols <- dplyr::mutate(olddf,
                                        ID = row_number(),
                                        ID = as.character(ID))
    names(newdf_join_cols)[names(newdf_join_cols) == "ID"] <- by_col
    names(olddf_join_cols)[names(olddf_join_cols) == 'ID'] <- by_col
    # check
    # mod_data <- list(newdf_join_cols, olddf_join_cols)
    # comparedf_list
    comparedf_list <- summary(arsenal::comparedf(
                                x = newdf_join_cols,
                                y = olddf_join_cols),
                                by = {{by_col}})
    diffs_table <- dplyr::mutate(
              .data = comparedf_list$diffs.table,
                  dplyr::across(.cols = dplyr::contains("values"),
                        .fns = base::as.character))
    diffs_names <- dplyr::select(
                    .data = diffs_table,
                      contains("row.names"),
                    `Variable name` = var.x,
                    `Current Value` = values.x,
                    `Previous Value` = values.y)
    names(diffs_names)[names(diffs_names) == "..row.names.."] <- by_col

    mod_data <- tibble::as_tibble(diffs_names)
  }

  return(mod_data)
}
```

## Use 

Below we will test the `create_modified_data()` function under various conditions. 

### No arguments provided 

If only datasets are provided, a row-by-row comparison is made on all columns. 

```{r nothing, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData)
```

```{r nothing-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData)
)
```

### Multiple `by` columns

If we need to create a new `JOIN` column from multiple columns, we can provide them to `by`.

```{r multiple-by, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"))
```

```{r multiple-by-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"))
)
```

### Multiple `by` columns, new `by_col`

If we need to create a new joining column from multiple columns and we want to provide a name for this new column, we can do this with `by` and `by_col`:

```{r multiple-by-new-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col")
```

```{r multiple-by-new-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col")
)
```

### Multiple `by` columns, new `by_col`, and `cols`

If we need to create a new joining column from multiple columns, and we want to provide a name for this new column, *and* we want to only compare a few columns, we can do this with `by`, `by_col`, and `cols`. 

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col", 
  cols = c("text_value_a", "text_value_b"))
```

```{r multiple-by-new-by_col-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_col", 
  cols = c("text_value_a", "text_value_b"))
)
```

### Single `by` column 

Single `by` column (assuming a join column exists), with no `by_col` or `cols`. 

```{r single-by-var, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var")
```

```{r single-by-var-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var")
)
```


### Single `by` column, new `by_col` name

Single `by` column (assuming a join column exists), with no `by_col` or `cols`. 

```{r single-by-var-new-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col")
```

```{r single-by-var-new-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col")
)
```


### Single `by` column, new `by_col` name, and `cols` 

```{r single-by-col-new-by_cols-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col",
  cols = c("text_value_a", "text_value_b")
  )
```

```{r single-by-col-new-by_cols-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentDataJoin, 
  olddf = PreviousDataJoin, 
  by = "join_var", 
  by_col = "new_join_col",
  cols = c("text_value_a", "text_value_b")
  ))
```


### Multiple `cols` 

If no `by` column is provided, and no `by_col`, but `cols` are provided, the function will perform a row-by-row comparison on the selected columns. 

```{r no-by-no-new-by_col-multiple-cols, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  cols = c("text_value_a", "text_value_b"))
```

```{r no-by-no-new-by_col-multiple-cols-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  cols = c("text_value_a", "text_value_b")))
```


### Only `by_col` 

If only a `by_col` is provided, the function will create an `ID` for each row in the datasets, then perform the comparison using this joining variable. 

```{r only-by_col, eval=FALSE}
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  by_col = "new_join_col")
```

```{r only-by_col-show, echo=FALSE}
knitr::kable(
create_modified_data(
  newdf = CurrentData, 
  olddf = PreviousData,
  by_col = "new_join_col"))
```



```{r old-fun, eval=FALSE, include=FALSE, echo=FALSE}
function(newdf, olddf, by = NULL, cols = NULL, return_cols = NULL) {
      # join col x
      data_x_join <- create_new_column(
                          data = newdf, 
                          cols = all_of(by), 
                          new_name = by) 
      # var names x
      data_x <- dplyr::rename_with(
        .data = data_x_join, 
        .fn =  base::paste0, 
        .cols = !all_of(by), "_x")
      # join col y
      data_y_join <- create_new_column(
                         data = olddf, 
                         cols = all_of(by), 
                         new_name = by) 
      # var names y
      data_y <- dplyr::rename_with(
        .data = data_y_join, 
        .fn =  base::paste0, 
        .cols = !all_of(by), "_y")
      # mod_table
      mod_table <- run_modified_data(newdf = newdf, 
                                     olddf = olddf, 
                                     by = by, 
                                     cols = cols)
    
      # join back to data_x
      data_x_mod <- dplyr::inner_join(
                             x = mod_table, 
                             y = data_x, 
                             by = all_of(by))
      # join back to data_y
      data_y_mod <- dplyr::inner_join(
                           x = mod_table, 
                           y = data_y, 
                           by = all_of(by))
      
      # join back to data_x and data_y
      data_both_mod <- dplyr::inner_join(x = data_x_mod, 
                               y = data_y, 
                               by = all_of(by))
  
}
```




```{r comparedf, eval=FALSE, include=FALSE}
arsenal::comparedf(x = , y = , by = , by.x = , by.y = , control = )
```


