
# Modified datasets

We're going to cover a few more functions for comparing datasets, but in order to do that, we need to have some multiple datasets to compare. Below is the `diff_modified_data` list, which contains two datasets with various differences:

```{r diff_modified_data}
diff_modified_data <- dfdiffs::diff_modified_data
```

```{r diff_current_modified}
current_mod <- diff_modified_data$diff_current_modified
current_mod
```

```{r diff_previous_modified}
previous_mod <- diff_modified_data$diff_previous_modified
previous_mod
```

For more complete comparisons, we can use the `proc_compare_df()` function. 

```{r proc_compare_df-args, eval=FALSE}
proc_compare_df(data_x = , data_y = , by = , by_col = , cols = )
```

`proc_compare_df()` allows us to make comparisons using the [`arsenal::comparedf` function](https://cran.r-project.org/web/packages/arsenal/vignettes/comparedf.html), but with a few additional options:

- `by`: a way of uniquely identifying rows between the datasets. `by` can be a single column or a combination of columns. 

- `by_col`: this is the new name of the `by` column (if multiple are supplied). 

- `cols`: the columns from the datasets we want to compare. 

```{r compare_list_01}
compare_list_01 <- proc_compare_df(
  data_x = current_mod, data_y = previous_mod, 
  by = c("subject_id", "record"), 
  by_col = "id" ,
  cols = c("text_value_a", "text_value_b"))
knitr::kable(compare_list_01$diffs.table)
```

We can omit the `by` column, and the `proc_compare_df()` will return a named join column (`JOINCOL`). 

```{r compare_list_02}
compare_list_02 <- proc_compare_df(
  data_x = current_mod, data_y = previous_mod, 
  by = c("subject_id", "record"), 
  cols = c("text_value_a", "text_value_b"))
knitr::kable(compare_list_02$diffs.table)
```

If we omit the `cols` argument, `proc_compare_df()` compares the values in all of the columns. 

```{r compare_list_03}
compare_list_03 <- proc_compare_df(
  data_x = current_mod, data_y = previous_mod, 
  by = c("subject_id", "record"))
knitr::kable(compare_list_03$diffs.table)
```



# Create modified data 

```{r create_modified_data}
create_modified_data <- function(x, y, join_cols, join_col_name, compare_cols, return_cols = "both") {
  # define inputs
  input_data_x <- x
  input_data_y <- y
  input_by <- as.character(join_cols)
  input_by_col <- as.character(join_col_name)
  input_cols <- as.character(compare_cols)
  # create compare_list
  compare_list <- proc_compare_df(
    data_x = input_data_x, 
    data_y = input_data_y, 
    by = join_cols, 
    by_col = join_col_name,
    
    )
  # diffs_table_raw
  diffs_table_raw <- mutate(.data = compare_list$diffs.table, 
      across(.cols = everything(), .fns = as.character))
  # join col x
  data_x_join <- create_join_column(df = input_data_x,
    by_colums = input_by,
    new_by_column_name = input_by_col) 
  # var names x
  data_x <- rename_with(.data = data_x_join, 
    .fn =  paste0, 
    .cols = !all_of(input_by_col), "_x")
  # join col x
  data_y_join <- create_join_column(df = input_data_y,
    by_colums = input_by,
    new_by_column_name = input_by_col)
  # var names y
  data_y <- rename_with(.data = data_y_join, 
    .fn =  paste0, 
    .cols = !all_of(input_by_col), "_y")
  # reorganize
  diffs_table <- select(diffs_table_raw, 
    variable = var.x, 
    all_of(input_by_col),
    `X variable value` = values.x,
    `Y variable value` = values.y,
    )
  # add modified cols
  mod_table <- tibble::add_column(.data = diffs_table, 
    modified = TRUE)
  # join back to data_x
  data_x_mod <- inner_join(x = mod_table, 
                           y = data_x, 
                           by = input_by_col)
  # join back to data_y
  data_y_mod <- inner_join(x = mod_table, 
                           y = data_y, 
                           by = input_by_col)
  # join back to data_x and data_y
  data_both_mod <- inner_join(x = data_x_mod, 
                           y = data_y, 
                           by = input_by_col)
  # return selected columns
  if (return_cols == "both") {
    return(data_both_mod)
  } else if (return_cols == "x") {
    return(data_x_mod)
  } else {
    return(data_y_mod)
  }
}
```

```{r test-create_modified_data}
create_modified_data(
  x = current_mod, 
  y = previous_mod, 
  join_cols = c("subject_id", "record"), 
  join_col_name = "id",
  compare_cols = c("text_value_a", "text_value_b"), 
  return_cols = "x")
```








The output from `proc_compare_df()` is a list of data frames, just like from the original [`arsenal::comparedf` function](https://cran.r-project.org/web/packages/arsenal/vignettes/comparedf.html). 

To understand how the comparison works, we can look at the `frame.summary.table` and `diffs.table`. 

```{r compare_list_01-frame.summary.table}
knitr::kable(compare_list_01$frame.summary.table)
```

```{r compare_list_01-diffs.table}
knitr::kable(compare_list_01$diffs.table)
```







```{r arsenal-diffs.table, echo=FALSE, eval=FALSE, include=FALSE}
library(arsenal)
# create joins
current_mod_join <- mutate(current_mod, 
  join = paste0(subject_id, record))
previous_mod_join <- mutate(previous_mod, 
  join = paste0(subject_id, record))
compare_list <- summary(arsenal::comparedf(
  x = current_mod_join, 
  y = previous_mod_join, 
  by = "subject_id"))
compare_list$diffs.table %>% 
  mutate(across(.cols = everything(), .fns = as.character)) 
  # select(
  #   variable = var.x, 
  #   current_value = values.x,
  #   previous_value = values.y
  # )
```




```{r teams-data}
teams_2010_raw <- load_flat_file(
  path = "../inst/extdata/csv/2010-lahman/Teams.csv")
teams_2010 <- select(teams_2010_raw,
  yearID, lgID, teamID,
  franchID, name, attendance,
  LgWin, DivWin, WSWin)

teams_2015_raw <- load_flat_file(
  "../inst/extdata/csv/2015-baseballdatabank/Teams.csv")
teams_2015 <- select(teams_2015_raw,
  yearID, lgID, teamID,
  franchID, name, attendance,
  LgWin, DivWin, WSWin)

teams_2020_raw <- load_flat_file(
  "../inst/extdata/csv/2020-baseballdatabank/Teams.csv")
teams_2020 <- select(teams_2020_raw,
  yearID, lgID, teamID,
  franchID, name, attendance,
  LgWin, DivWin, WSWin)
```

