---
title: "change-frequency-from-proc-compare"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
    df_print: kable
vignette: >
  %\VignetteIndexEntry{change-frequency-from-proc-compare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height=5,
  fig.width=7
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(openxlsx)
library(readxl)
library(tidyverse)
library(janitor) 
library(arsenal) # comparedf
library(diffdf)  # diffdf
```

# Motivation 

In this vignette I'll cover how to calculate the frequency of changes per column for each dataset in a PROC COMPARE output file, `snapshot_compare_270301_20221116_year3_preview_noDAP.xlsx`.

We'll import the Excel file, then build a regular expression for separating the changes recorded in the `value_changes` column (the cleaned name of the `Value changes` column).

# Import data

The code below lists the sheet names in the `snapshot_compare_270301_20221116_year3_preview_noDAP.xlsx` file.

```{r snapshot_pth}
snapshot_pth <- "../inst/extdata/xlsx/snapshot_compare_270301_20221116_year3_preview_noDAP.xlsx"
snapshoty3_sheets <- readxl::excel_sheets(snapshot_pth)
snapshoty3_sheets
```


## AE 

We'll import the `"AE"` sheet and check the column names.

```{r snap_ae_raw}
snap_ae_raw <- readxl::read_excel(snapshot_pth, sheet = "AE") 
names(snap_ae_raw) |> head(10)
names(snap_ae_raw) |> tail(10)
```


Next, we'll select and rename the columns we need, clean their names, and filter the `"AE"` dataset to the rows with an `"Updated"` value in the `status` column.

```{r snap_updated}
snap_updated <- snap_ae_raw |>  
  select(
    subject_id = `Subject name or identifier`, 
    record_num = `Record number`,
    DATASET, 
    STATUS, 
    `Value changes`, 
    NEW_VALUES = NEWVALUES) %>% 
  janitor::clean_names() |> 
  filter(status == "Updated")
glimpse(snap_updated)
```


### Testing `SAEDESC1`

Changes to `SAEDESC1` contain a lot of unstructured text with multiple commas, which makes them a good test case for separating the column names from the `value_changes` column.

```{r test_SAEDESC1}
test_SAEDESC1 <- snap_updated |> 
  filter(str_detect(snap_updated$value_changes, "SAEDESC1")) |> 
  head(3)
glimpse(test_SAEDESC1)
```

### Testing the regular expression

Below is the regular expression we'll test these rows against.

```{r sas_cols_regex}
sas_cols_regex <- "\\w*[A-Z]\\w*[A-Z]\\w*="
sas_cols_regex
```


We'll use `str_view_all()` to check whether the regular expression matches the column names.

```{r test-SAEDESC1-02}
str_view_all(string = test_SAEDESC1$value_changes, 
  pattern = sas_cols_regex) 
```

## `count_change_columns()`

Now that we have a regular expression that matches the column names in `value_changes`, we'll write a function that counts the occurrences of each match.

```{r count_change_columns}
count_change_columns <- function(df) {
  sas_cols_regex <- "\\w*[A-Z]\\w*[A-Z]\\w*="
  # clean names
  df_clean_nms <- df |> janitor::clean_names()
  # value_changes
  value_change_match <- dplyr::mutate(df_clean_nms, 
         raw_col_match = stringr::str_match_all(value_changes, 
                                                sas_cols_regex)) |> 
    dplyr::relocate(raw_col_match, .after = 1)
  # unnest 
  raw_matches_long <- tidyr::unnest_longer(data = value_change_match, 
                                  col = raw_col_match)
  col_matches_long <- dplyr::mutate(.data = raw_matches_long,
      col_matches = stringr::str_remove_all(raw_col_match, "=")) |> 
    dplyr::relocate(col_matches, .after = 2)
  col_count_raw <- dplyr::count(col_matches_long, col_matches, 
                            name = "occurrence")
  col_count <- dplyr::filter(col_count_raw, !is.na(col_matches))
  count_tbl <- dplyr::rename(.data = col_count,
    Field = col_matches,
    ChangeCount = occurrence
  )
  return(count_tbl)
}
```

## Field and Change Count

We can derive `Field` and `ChangeCount` by passing `snap_ae_raw` (the data imported from the spreadsheet) to `count_change_columns()`.

```{r ae_change_freq}
ae_change_freq <- count_change_columns(df = snap_ae_raw)
ae_change_freq
```

## Graph Changes 

We can graph these changes, but we have to set the font size manually because of the long `y` axis.

```{r graph-ae-changes}
ggplot(data = ae_change_freq, 
       mapping = aes(x = ChangeCount, 
                     y = forcats::fct_reorder(.f = Field, 
                                              .x = ChangeCount))) + 
  geom_col(aes(fill = Field), show.legend = FALSE) + 
  labs(title = "AE Changes", 
       y = "Field", 
       x = "Number of Changes") + 
  theme_minimal(base_size = 6)
```

# Map to other sheets

Now we can apply `count_change_columns()` to every sheet in `snapshot_compare_270301_20221116_year3_preview_noDAP.xlsx`.

First, we remove the `Summary` and `Key variables` sheets from `snapshoty3_sheets`.

```{r form_sheets}
form_sheets <- snapshoty3_sheets[! snapshoty3_sheets %in% c('Summary', 'Key variables')]
form_sheets
```

Then we name each element in the vector with its own sheet name and store the result in `form_sheets_nmd`.

```{r form_sheets_nmd}
form_sheets_nmd <- purrr::set_names(form_sheets)
form_sheets_nmd
```

## Import all sheets

We can now import every sheet from the file at `snapshot_pth` and store them in the `all_sheets` list.

```{r all_sheets}
all_sheets <- form_sheets_nmd |> 
  map(~ readxl::read_excel(path = snapshot_pth, sheet = .x))
```


The names of `all_sheets` confirm the list contains every imported sheet.

```{r sheet-names-all_sheets}
names(all_sheets)
```

## Map `count_change_columns()` to `all_sheets`

We'll `map()` `count_change_columns()` over each element in `all_sheets` and store the results in the `all_field_counts` list.

```{r all_field_counts}
all_field_counts <- purrr::map(.x = all_sheets, .f = count_change_columns)
names(all_field_counts)
```

We can check the results by graphing the changes in a few of the forms.

### CM Changes

The graph below shows the number of changes per field in the `CM` form.

```{r graph-cm-changes}
all_field_counts$CM |> 
  ggplot(mapping = aes(x = ChangeCount, 
                     y = forcats::fct_reorder(.f = Field, 
                                              .x = ChangeCount))) + 
  geom_col(aes(fill = Field), show.legend = FALSE) + 
  labs(title = "CM Changes", 
       y = "Field", 
       x = "Number of Changes") + 
  theme_minimal(base_size = 6)
```

### QSQOL Changes

The graph below shows the number of changes per field in the `QSQOL` form.

```{r graph-qsqol-changes}
all_field_counts$QSQOL |> 
  ggplot(mapping = aes(x = ChangeCount, 
                     y = forcats::fct_reorder(.f = Field, 
                                              .x = ChangeCount))) + 
  geom_col(aes(fill = Field), show.legend = FALSE) + 
  labs(title = "QSQOL Changes", 
       y = "Field", 
       x = "Number of Changes") + 
  theme_minimal(base_size = 6)
```


### VS Changes

The graph below shows the number of changes per field in the `VS` form.

```{r graph-vs-changes}
all_field_counts$VS |> 
  ggplot(mapping = aes(x = ChangeCount, 
                     y = forcats::fct_reorder(.f = Field, 
                                              .x = ChangeCount))) + 
  geom_col(aes(fill = Field), show.legend = FALSE) + 
  labs(title = "VS Changes", 
       y = "Field", 
       x = "Number of Changes") + 
  theme_minimal(base_size = 6)
```

## Export field count tables

We can export these tables to the `inst/out/` folder as `all-field-counts.xlsx`, with one sheet per form.

```{r export}
library(openxlsx)
# create wb
openxlsx::write.xlsx(x = all_field_counts,
                file = "../inst/out/all-field-counts.xlsx")
```



