---
title: "changed-modified-data"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
    df_print: kable
vignette: >
  %\VignetteIndexEntry{changed-modified-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

# Motivation

When we receive an updated version of a dataset, `dfdiffs` helps us answer three questions:

1. What rows are here now that weren't here before?
2. What rows were here before that aren't here now?
3. *What values have been changed?*

This vignette covers the two functions that answer the third question: `create_changed_data()` and `create_modified_data()`. Both are built on the same base R engine (`diff_values_by_key()` and `compare_values()`), and they differ only in the shape of their output:

- `create_changed_data()` returns `$num_diffs` and `$var_diffs`, which are used by `compare_data()`.
- `create_modified_data()` returns `$diffs_byvar` and `$diffs`, which are used by `mod_compare` and `create_comparison_report()`.

## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(arsenal) # comparedf, for comparison
```

## Example data

We'll use two datasets from the package to test what's been modified.

### Initial Data

`InitialData` is the base (original) dataset:

```{r InitialData}
InitialData <- dfdiffs::InitialData
```

```{r InitialData-show, echo=FALSE}
knitr::kable(InitialData, caption = "InitialData") |>
  kableExtra::kable_paper()
```

### Changed Data

`ChangedData` is the dataset we'll compare against `InitialData`:

```{r ChangedData}
ChangedData <- dfdiffs::ChangedData
```

```{r ChangedData-show, echo=FALSE}
knitr::kable(ChangedData, caption = "ChangedData") |>
  kableExtra::kable_paper()
```


## Creating join columns

We'll use `create_new_column()` to build a `join_var` column from `subject_id` and `record` in both datasets:

```{r ChangedDataJoin-InitialDataJoin}
ChangedDataJoin <- create_new_column(
  data = ChangedData,
  cols = c("subject_id", "record"),
  new_name = "join_var")
InitialDataJoin <- create_new_column(
  data = InitialData,
  cols = c("subject_id", "record"),
  new_name = "join_var")
```

`ChangedDataJoin` now starts with the new `join_var` column:

```{r ChangedDataJoin-show, echo=FALSE}
knitr::kable(ChangedDataJoin, caption = "ChangedDataJoin") |>
  kableExtra::kable_paper()
```

# create_changed_data()

## Arguments

`create_changed_data()` takes the same `compare`, `base`, `by`, `by_col`, and `cols` arguments as `create_new_data()`. The [reference table](create-new-data.html#arguments) in that vignette lists every combination.

The difference is what the function does with matched rows. `create_new_data()` and `create_deleted_data()` find rows with no match, but `create_changed_data()` compares the *values* within matched rows, column by column.

## Call structure

`create_changed_data()` is built on `diff_values_by_key()`, a base R engine that uses `compare_values()` to compare matched rows column by column. It also uses the same `rename_join_col()` and `create_new_column()` helpers as the other comparison functions. The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-create_changed_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "create_changed_data")
```

```
█─create_changed_data
├─select_cols
├─█─diff_values_by_key
│ ├─compare_values
│ └─class_diffs
├─rename_join_col
├─create_new_column
└─class_diffs
```

## Worked examples

### No `by` column: row-by-row comparison

When we supply only the two datasets, `create_changed_data()` compares them row by row. The `$num_diffs` and `$var_diffs` tables are shown side by side below:

```{r changed-no-by, eval=FALSE}
create_changed_data(
  compare = ChangedData,
  base = InitialData)
```

```{r changed-no-by-show, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs),
    knitr::kable(changed$var_diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```

### A single `by` column

Passing `join_var` as the `by` column matches rows on that key instead of on row position:

```{r changed-single-by, eval=FALSE}
create_changed_data(
  compare = ChangedDataJoin,
  base = InitialDataJoin,
  by = "join_var")
```

```{r changed-single-by-show, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedDataJoin,
  base = InitialDataJoin,
  by = "join_var")
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs),
    knitr::kable(changed$var_diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```

### Multiple `by` columns, a new `by_col`, and `cols`

This is the most complete case. The `by` columns are combined into a single key named with `by_col`, and only the columns in `cols` are compared:

```{r changed-multi-by, eval=FALSE}
create_changed_data(
  compare = ChangedData,
  base = InitialData,
  by = c("subject_id", "record"),
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
```

```{r changed-multi-by-show, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData,
  by = c("subject_id", "record"),
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs),
    knitr::kable(changed$var_diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```

***

# create_modified_data()

`create_modified_data()` answers the same question as `create_changed_data()`, but its output is shaped for the "Review Changes" report instead of `compare_data()`.

## `arsenal::comparedf()`, for comparison

Before version 2.1.0, `create_modified_data()` was built directly on `arsenal::comparedf()`. To see where the current output comes from, we'll run `arsenal::comparedf()` on the same data:

```{r modified_cdf}
modified_cdf <- arsenal::comparedf(
  x = ChangedDataJoin,
  y = InitialDataJoin,
  by = "join_var")
modified_cdf_sum <- summary(modified_cdf)
```

The summary contains several tables, but we're only interested in `diffs.byvar.table` and `diffs.table`:

```{r comparedf-diffs.byvar.table, echo=FALSE}
knitr::kable(modified_cdf_sum$diffs.byvar.table,
  caption = "modified_cdf_sum$diffs.byvar.table") |>
  kableExtra::kable_paper()
```

```{r comparedf-diffs.table, echo=FALSE}
knitr::kable(modified_cdf_sum$diffs.table,
  caption = "modified_cdf_sum$diffs.table") |>
  kableExtra::kable_paper()
```

`create_modified_data()` returns only a small fraction of this output, reshaped into `$diffs_byvar` and `$diffs`. The extra `var.y` column is dropped, `var.x` becomes `Variable name`, and `values.x` and `values.y` become `Current Value` and `Previous Value`. The row position columns (`row.x` and `row.y`) are also dropped in favor of the join key.

We can see the reshaped output by running `create_modified_data()` on the same joined datasets:

```{r create_modified_data}
modified <- create_modified_data(
  compare = ChangedDataJoin,
  base = InitialDataJoin,
  by = "join_var")
```

```{r modified-diffs_byvar-show, echo=FALSE}
knitr::kable(modified$diffs_byvar, caption = "modified$diffs_byvar") |>
  kableExtra::kable_paper()
```

```{r modified-diffs-show, echo=FALSE}
knitr::kable(modified$diffs, caption = "modified$diffs") |>
  kableExtra::kable_paper()
```

## Arguments

`create_modified_data()` takes the same `compare`, `base`, `by`, `by_col`, and `cols` arguments as `create_new_data()` and `create_changed_data()`. The [reference table](create-new-data.html#arguments) lists every combination.

## Call structure

`create_modified_data()` is built on the same `diff_values_by_key()` engine as `create_changed_data()`. The call tree below was generated with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-create_modified_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "create_modified_data")
```

```
█─create_modified_data
├─select_cols
├─█─diff_values_by_key
│ ├─compare_values
│ └─class_diffs
├─rename_join_col
├─create_new_column
└─class_diffs
```

## Worked examples

### No `by` column: row-by-row comparison

When we supply only the two datasets, `create_modified_data()` compares them row by row. The `$diffs_byvar` and `$diffs` tables are shown side by side below:

```{r modified-no-by, eval=FALSE}
create_modified_data(
  compare = ChangedData,
  base = InitialData)
```

```{r modified-no-by-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar),
    knitr::kable(compare_list$diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```

### A single `by` column

Passing `join_var` as the `by` column matches rows on that key instead of on row position:

```{r modified-single-by, eval=FALSE}
create_modified_data(
  compare = ChangedDataJoin,
  base = InitialDataJoin,
  by = "join_var")
```

```{r modified-single-by-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin,
  base = InitialDataJoin,
  by = "join_var")
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar),
    knitr::kable(compare_list$diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```

### Multiple `by` columns, a new `by_col`, and `cols`

As with `create_changed_data()`, the `by` columns are combined into a single key named with `by_col`, and only the columns in `cols` are compared:

```{r modified-multi-by, eval=FALSE}
create_modified_data(
  compare = ChangedData,
  base = InitialData,
  by = c("subject_id", "record"),
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
```

```{r modified-multi-by-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData,
  base = InitialData,
  by = c("subject_id", "record"),
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar),
    knitr::kable(compare_list$diffs)
    ), format = "html") |>
  kableExtra::kable_paper()
```
