---
title: "changed-modified-data"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
    df_print: kable
vignette: >
  %\VignetteIndexEntry{changed-modified-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

The goal of the `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
**3. What values have been changed?**   

This vignette takes us through two functions: 

1. `create_changed_data()`: which uses the `arsenal::comparedf()` function

2. `create_modified_data()`: which uses the `diffdf::diffdf()` function

Both functions answers the "*What values have been changed?*"

## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
```

## Changed data 

We have two datasets to test what's been modified:

### Initial Data

```{r InitialData}
InitialData <- dfdiffs::InitialData
```

```{r InitialData-show, echo=FALSE}
knitr::kable(InitialData, caption = "InitialData") |> 
  kableExtra::kable_paper()
```

### Changed Data

```{r ChangedData}
ChangedData <- dfdiffs::ChangedData
```

```{r ChangedData-show, echo=FALSE}
knitr::kable(ChangedData, caption = "ChangedData") |> 
  kableExtra::kable_paper()
```


## Creating join columns

We will use our `create_new_column()` function to create `join_var` from `subject_id` and `record`

```{r ChangedDataJoin, eval=FALSE}
ChangedDataJoin <- dfdiffs::create_new_column(
  data = ChangedData, 
  cols = c("subject_id", "record"), 
  new_name = "join_var")
ChangedDataJoin
```

```{r ChangedDataJoin-show, echo=FALSE}
ChangedDataJoin <- dfdiffs::create_new_column(
  data = ChangedData, 
  cols = c("subject_id", "record"), 
  new_name = "join_var")
knitr::kable(ChangedDataJoin, caption = "ChangedDataJoin") |> 
  kableExtra::kable_paper()
```

```{r InitialDataJoin, eval=FALSE}
InitialDataJoin <- dfdiffs::create_new_column(
  data = InitialData, 
  cols = c("subject_id", "record"), 
  new_name = "join_var")
InitialDataJoin
```

```{r InitialDataJoin-show, echo=FALSE}
InitialDataJoin <- dfdiffs::create_new_column(
  data = InitialData, 
  cols = c("subject_id", "record"), 
  new_name = "join_var")
knitr::kable(InitialDataJoin, caption = "InitialDataJoin") |> 
  kableExtra::kable_paper()
```


# create_changed_data()

Each function in the `dfdiffs` package assumes the following conditions: 

1) Two datasets

2) Multiple columns to compare (`cols`)

3) Single by column

4) Single `by` column, new column name (`by_col`)

5) Single `by` column, multiple compare columns (`cols`)

6) `Single` by column, new column name (`by_col`), multiple compare columns (`cols`)

7) Multiple `by` columns

8) Multiple `by` columns, new column name (`by_col`)

9) Multiple `by` columns, multiple compare columns (`cols`)

10) Multiple `by` columns, a new `by_col`, and `cols`


## Single by column conditions

### 1) Two datasets

compare all columns:

```{r cond_1, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData)
```

```{r , echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData)
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 2) Multiple columns to compare (`cols`)

```{r cond-2, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  cols = c("text_value_a", "text_value_b"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```



### 3) Single `by` column

No new column name

```{r cond-3, eval=FALSE}
create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var")
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var")
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 4) Single `by` column, new column name (`by_col`)

```{r cond-4, eval=FALSE}
create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  by_col = "join")
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  by_col = "join")
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 5) Single `by` column, multiple compare columns (`cols`)

```{r cond-5, eval=FALSE}
create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  cols = c("text_value_a", "text_value_b"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

```{r cond-6, eval=FALSE}
create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```



## Multiple by column conditions

### 7) Multiple `by` columns

```{r cond-7, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```



### 8) Multiple `by` columns, new column name (`by_col`)

```{r cond-8, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var")
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var")
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 9) Multiple `by` columns, multiple compare columns (`cols`)

```{r cond-9, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  cols = c("text_value_a", "text_value_b"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 10) Multiple `by` columns, a new `by_col`, and `cols`

```{r cond-10, eval=FALSE}
create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
```

```{r, echo=FALSE}
changed <- create_changed_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "join",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(changed$num_diffs), 
    knitr::kable(changed$var_diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


# create_modified_data()

Below we have the `create_modified_data()` function, which takes two data frames (`compare` and `base`), a `by` column, and the `cols` to compare. 

```{r args-create_modified_data, eval=FALSE}
create_modified_data(compare, base, by = NULL, by_col = NULL, cols = NULL)
```

## Single joining (`by`) column conditions 

### 1) Two datasets

+ No `by` columns (only two datasets)

```{r compare_list-01, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData)
```

```{r ompare_list-01-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData)
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```


### 2) Multiple columns to compare (`cols`)

+ No `by` columns (only two datasets) and multiple compare (`cols`)  

```{r compare_list-02, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  cols = c("text_value_a", "text_value_b"))
```

```{r ompare_list-02-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 3) Single `by` column

+ We can provide a single `by` column (using our `InitialDataJoin` and `ChangedDataJoin`) datasets we created above. 

```{r compare_list-03, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var")
```

```{r ompare_list-03-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var")
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 4) Single `by` column, new column name (`by_col`)

+ We can also provide a single `by` column (for unique identifiers) and a new name for the `by_col`


```{r compare_list-04, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  by_col = "new_join_var")
```

```{r compare_list-04-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  by_col = "new_join_var")
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 5) Single `by` column, multiple compare columns `cols`

+ Single `by` column and multiple compare columns (`cols`) 

```{r compare_list-05, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  cols = c("text_value_a", "text_value_b"))
```

```{r compare_list-05-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

- Single `by` column, a new name for the by column (`by_col`), and multiple compare columns (`cols`)

```{r compare_list-06, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var", 
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
```

```{r compare_list-06-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedDataJoin, 
  base = InitialDataJoin, 
  by = "join_var",
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

## Multiple `by` column conditions

The next conditions cover when two existing columns are used to create a unique identifier between the two datasets. 

### 7) Multiple `by` columns

+ Multiple `by` columns (assuming the columns create a unique identifier).

```{r multiple-by-07, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"))
```

```{r multiple-by-07-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

This creates a `join` column from the `by` columns. 

### 8) Multiple `by` columns, new column name (`by_col`)

We can provide multiple `by` columns, a new `by_col`, and  **no `cols`**

```{r multiple-by-08, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var")
```

```{r multiple-by-08-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var")
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

### 9) Multiple `by` columns, multiple compare columns (`cols`)  

+ Multiple `by` columns and multiple compare columns (`cols`), and **no new `by_col`**.

```{r multiple-by-09, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  cols = c("text_value_a", "text_value_b"))
```

```{r multiple-by-09-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

This creates a `join` column from the `by` columns. 

### 10) Multiple `by` columns, a new `by_col`, and `cols`

+ We can provide multiple `by` columns, a new `by_col`, and multiple `cols`.

```{r multiple-by-10, eval=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
```

```{r multiple-by-10-show, echo=FALSE}
compare_list <- create_modified_data(
  compare = ChangedData, 
  base = InitialData, 
  by = c("subject_id", "record"), 
  by_col = "new_join_var",
  cols = c("text_value_a", "text_value_b"))
knitr::kables(
  x = list(
    knitr::kable(compare_list$diffs_byvar), 
    knitr::kable(compare_list$diffs)
    ), format = "html") |> 
  kableExtra::kable_paper()
```

```{r app-data, include=FALSE, eval=FALSE}
prev_test_data <- tibble::tribble(
     ~record, ~id_number, ~integer, ~double, ~boolean,    ~type,
         "A1",         1L,       4L,     1.1,     TRUE, "stable",
         "B2",         2L,      16L,     4.1,    FALSE, "stable",
         "C3",         3L,      24L,     3.1,     TRUE,    "del",
         "E5",         5L,      32L,     5.1,    FALSE,    "mod")
curr_test_data <- tibble::tribble(
     ~record, ~id_number, ~integer, ~double, ~boolean,    ~type,
         "A1",         1L,       4L,     1.1,     TRUE, "stable",
         "B2",         2L,      16L,     4.1,    FALSE, "stable",
         "D4",         4L,      28L,     3.1,    FALSE,    "new",
         "E5",         5L,      32L,     5.2,    TRUE,     "mod")
# test new
create_new_data(compare = curr_test_data, base = prev_test_data, 
  by = "record", 
  cols = c("integer", "double", "boolean"))
# test deleted 
create_deleted_data(compare = curr_test_data, base = prev_test_data, 
  by = "record", 
  cols = c("integer", "double", "boolean"))
# test modified 
create_modified_data(compare = curr_test_data, base = prev_test_data, 
  by = "record",
  cols = c("integer", "double", "boolean"))
```




