---
title: "compare-data"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
    df_print: kable
vignette: >
  %\VignetteIndexEntry{compare-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(openxlsx)
library(janitor) 
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(stringr)
library(lubridate)
library(glue)
library(purrr)
library(reactable)
library(haven)
library(readxl)
```


# Motivation

The other vignettes answer each `dfdiffs` question with its own function. In this vignette I'll cover `compare_data()`, which answers all three questions in a single call:

1. What rows are here now that weren't here before?
2. What rows were here before that aren't here now?
3. What values have been changed?

## Test data

`Roster2021` and `Roster2022` are two yearly pulls of a synthetic (not real) clinical site roster. They're larger than the toy datasets in the other vignettes, so they show how `dfdiffs` works at a more realistic scale (see `?Roster2021` for details). The code below loads both rosters and prints their row counts:

```{r masters}
r21 <- dfdiffs::Roster2021
nrow(r21)
r22 <- dfdiffs::Roster2022
nrow(r22)
```

## The `compare_data()` function

`compare_data()` takes the following arguments:

```{r compare_data-args, eval=FALSE}
compare_data(compare = , base = , by = , by_col = , cols = , ignore_case = , trim_ws = )
```

```{r compare_data, echo=FALSE, eval=FALSE}
compare_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL,
                          ignore_case = FALSE, trim_ws = FALSE) {

  new_data <- create_new_data(
    compare = compare, base = base, by = by, by_col = by_col, cols = cols)

  deleted_data <- create_deleted_data(
    compare = compare, base = base, by = by, by_col = by_col, cols = cols)

  changed_data <- create_changed_data(
    compare = compare, base = base, by = by, by_col = by_col, cols = cols,
    ignore_case = ignore_case, trim_ws = trim_ws)

  changed_num_diffs <- changed_data$num_diffs
  changed_var_diffs <- changed_data$var_diffs

  return(list(
    'new_data' = new_data,
    'deleted_data' = deleted_data,
    'changed_num_diffs' = changed_num_diffs,
    'changed_var_diffs' = changed_var_diffs,
    'changed_class_diffs' = changed_data$class_diffs,
    'column_diffs' = compare_columns(base = base, compare = compare)
  ))
}
```

The code below compares the 2022 roster against the 2021 roster, matches rows on `subject_id`, and limits the comparison to four columns. `compare_data()` returns a named list:

```{r test-compare_data}
comparisons <- compare_data(
  compare = r22, base = r21,
  by = "subject_id",
  cols = c("first_name", "last_name", "full_name", "height_cm"))
names(comparisons)
```

## Call structure

`compare_data()` runs `create_new_data()`, `create_deleted_data()`, and `create_changed_data()`, adds the column-structure diff from `compare_columns()`, and returns everything in a single list. I generated the call tree below from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`), and it only shows functions defined in `dfdiffs`.

```{r call-tree-compare_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "compare_data")
```

```
█─compare_data
├─█─create_new_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─█─create_deleted_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─█─create_changed_data
│ ├─select_cols
│ ├─█─diff_values_by_key
│ │ ├─compare_values
│ │ └─class_diffs
│ ├─rename_join_col
│ ├─create_new_column
│ └─class_diffs
├─class_diffs
└─compare_columns
```

The Shiny app's compare module (`mod_compare_server()`) orchestrates three comparison functions: `create_new_data()`, `create_deleted_data()`, and `create_modified_data()` (`compare_data()` uses `create_changed_data()` instead). The module uses `compare_columns()` for its column-intersection reactive, and it uses the `%nin%` helper to drop `join_column`, `data_source`, and `join_source` from the list of compared columns. Its download handler calls `create_comparison_report()` directly, so the module has no report-building logic of its own:

```{r call-tree-mod_compare_server, eval=FALSE}
stackcallr::call_tree_dir("R", root = "mod_compare_server")
```

```
█─mod_compare_server
├─compare_columns
├─%nin%
├─info_react_theme
├─█─create_new_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─new_react_theme
├─█─create_deleted_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─deleted_react_theme
├─█─create_modified_data
│ ├─select_cols
│ ├─█─diff_values_by_key
│ │ ├─compare_values
│ │ └─class_diffs
│ ├─rename_join_col
│ ├─create_new_column
│ └─class_diffs
├─select_cols
├─changed_react_theme
├─left_join_base
└─█─create_comparison_report
  ├─█─create_new_data
  │ ├─anti_join_base
  │ ├─select_cols
  │ ├─rename_join_col
  │ └─create_new_column
  ├─█─create_deleted_data
  │ ├─anti_join_base
  │ ├─select_cols
  │ ├─rename_join_col
  │ └─create_new_column
  ├─█─create_modified_data
  │ ├─select_cols
  │ ├─█─diff_values_by_key
  │ │ ├─compare_values
  │ │ └─class_diffs
  │ ├─rename_join_col
  │ ├─create_new_column
  │ └─class_diffs
  ├─create_empty_tbl
  ├─compare_columns
  ├─class_diffs
  └─█─compare_summary
    └─█─compare_data
      ├─█─create_new_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_deleted_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_changed_data
      │ ├─select_cols
      │ ├─█─diff_values_by_key
      │ │ ├─compare_values
      │ │ └─class_diffs
      │ ├─rename_join_col
      │ ├─create_new_column
      │ └─class_diffs
      ├─class_diffs
      └─compare_columns
```

## `$new_data`

`$new_data` holds the rows in `r22` that aren't in `r21`. The table below shows the first six:

```{r new_data, eval=FALSE}
comparisons$new_data
```

```{r , echo=FALSE}
knitr::kable(head(comparisons$new_data), caption = "head(comparisons$new_data)") |> 
  kableExtra::kable_paper()
```


## `$deleted_data`

`$deleted_data` holds the rows in `r21` that are missing from `r22`. The table below shows the first six:

```{r deleted_data, eval=FALSE}
comparisons$deleted_data
```

```{r , echo=FALSE}
knitr::kable(head(comparisons$deleted_data), 
  caption = "head(comparisons$deleted_data)") |> 
  kableExtra::kable_paper()
```


## `$changed_num_diffs`

`$changed_num_diffs` stores the number of changed values for each compared column:

```{r changed_num_diffs, eval=FALSE}
comparisons$changed_num_diffs
```

```{r , echo=FALSE}
knitr::kable(comparisons$changed_num_diffs, 
  caption = "comparisons$changed_num_diffs") |> 
  kableExtra::kable_paper()
```


## `$changed_var_diffs`

`$changed_var_diffs` lists the changed values row by row:

```{r changed_var_diffs, eval=FALSE}
comparisons$changed_var_diffs
```

```{r , echo=FALSE}
knitr::kable(comparisons$changed_var_diffs,
  caption = "comparisons$changed_var_diffs") |>
  kableExtra::kable_paper()
```

## `$changed_class_diffs`

`$changed_class_diffs` lists class mismatches on shared columns, and it's
empty when every shared column has a matching class. `Date`/`POSIXct` pairs
and factor/character pairs are unified before the check, so they aren't
reported as mismatches:

```{r changed_class_diffs, eval=FALSE}
comparisons$changed_class_diffs
```

```{r , echo=FALSE}
knitr::kable(comparisons$changed_class_diffs,
  caption = "comparisons$changed_class_diffs") |>
  kableExtra::kable_paper()
```

## `$column_diffs`

`$column_diffs` is a `compare_columns()` object that lists the columns found
only in `base`, only in `compare`, and in both. It prints readably at the
console:

```{r column_diffs}
comparisons$column_diffs
```

