---
title: "create-join-column"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create-join-column}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  str = strOptions(strict.width = "wrap"),
  width = 75,
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

This vignette walks through the `create_join_column()` function, which creates new columns in the `dfdiffs` Shiny application.

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(shiny)
library(data.table)
library(stringr)
library(lubridate)
library(glue)
library(purrr)
library(reactable)
library(haven)
library(readxl)
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## create_join_column()

This function allows us to create a new 'join' column from any number of column inputs.

```{r create_join_column, eval=FALSE}
create_join_column(df, by_colums, new_by_column_name)
```

## Call structure

`create_join_column()` doesn't call any other `dfdiffs` functions. Inside the Shiny app, it's called by the select module (`mod_select_server()`) to build the `join_column` from the columns chosen in the *Select Join Columns* panel. The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`); only functions defined in `dfdiffs` are shown.

```{r call-tree-mod_select_server, eval=FALSE}
stackcallr::call_tree_dir("R", root = "mod_select_server")
```

```
█─mod_select_server
└─create_join_column
```

## Test data

We'll load some test data from the synthetic (not real) `dfdiffs` site roster (see `?Roster2021`) to demonstrate how this function works.

### Roster pull (2021)

Below is the 2021 roster pull.

```{r lahman2010_files}
roster2021_files <- "../inst/extdata/csv/site-roster/2021/"
fs::dir_tree(roster2021_files)
```


```{r master_2010}
roster2021_csv_paths <- list.files(
  path = roster2021_files,
  full.names = TRUE,
  pattern = ".csv$")
roster2021_path <- roster2021_csv_paths[grepl(pattern = "Roster", roster2021_csv_paths)]
roster_2021 <- data.table::fread(input = roster2021_path)
str(roster_2021)
```

### Roster pull (2022)

Below is the 2022 roster pull.

```{r lahman2015_files}
roster2022_files <- "../inst/extdata/csv/site-roster/2022/"
fs::dir_tree(roster2022_files)
```


```{r master_2015}
roster2022_csv_paths <- list.files(path = roster2022_files,
                                    full.names = TRUE,
                                    pattern = ".csv$")
roster2022_path <- roster2022_csv_paths[grepl(pattern = "Roster", roster2022_csv_paths)]
roster_2022 <- data.table::fread(input = roster2022_path)
str(roster_2022)
```

Assume we want to create a new join column that identifies unique rows by `first_name`, `last_name`, and `enroll_year`.

```{r name_year_id}
join_roster_2021 <- create_join_column(df = roster_2021,
  by_colums = c("first_name", "last_name", "enroll_year"),
  new_by_column_name = "name_year_id")

join_roster_2022 <- create_join_column(df = roster_2022,
  by_colums = c("first_name", "last_name", "enroll_year"),
  new_by_column_name = "name_year_id")

anti_join_base(
  # return all rows from join_roster_2021 WITHOUT a match in join_roster_2022
  x = join_roster_2021,
  y = join_roster_2022,
  by = "name_year_id") |> str()
```

These are the rows in `join_roster_2021` that aren't in `join_roster_2022`.
