---
title: "create-join-column"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create-join-column}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  str = strOptions(strict.width = "wrap"),
  width = 75,
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

This vignette walks through the `create_join_column()` function, which is used for creating new columns in the `dfdiffs` shiny application. 

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(shiny)
library(data.table)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(forcats)
library(glue)
library(purrr)
library(vroom)
library(reactable)
library(haven)
library(readxl)
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## create_join_column()

This function allows us to create a new 'join' column from any number of column inputs. 

```{r create_join_column}
create_join_column <- function(df, join_cols, new_col_name){
  # select by_vars
  tmp <- select(df, all_of(join_cols))
  # convert to character
  tmp <- mutate(tmp, across(.fns = as.character))
  # rename data 
  join_col_data <- df
  # assign new col
  join_col_data$new_col <- pmap_chr(.l = tmp, .f = paste, sep = "-")
  # rename 
  names(join_col_data)[names(join_col_data) == "new_col"] <- new_col_name
  # relocate
  join_col_data <- relocate(join_col_data, all_of(new_col_name))
  # return
  return(join_col_data)
}
```

## Test data 

We will load some test data from the `Lahman` package to demonstrate how this function works: 

### Hall of Fame (2010)

Below is the Hall of Fame data from 2010.

```{r lahman2010_files}
lahman2010_files <- "../inst/extdata/csv/2010-lahman/"
fs::dir_tree(lahman2010_files)
```


```{r master_2010}
lahman_2010_csv_paths <- list.files(
  path = lahman2010_files, 
  full.names = TRUE, 
  pattern = ".csv$")
master_2010_path <- lahman_2010_csv_paths[grepl(pattern = "Master", lahman_2010_csv_paths)]
master_2010 <- data.table::fread(input = master_2010_path)
glimpse(master_2010)
```

### Master data (2015)

And the 'Master' data from 2015.

```{r lahman2015_files}
lahman2015_files <- "../inst/extdata/csv/2015-baseballdatabank/"
fs::dir_tree(lahman2015_files)
```


```{r master_2015}
lahman_2015_csv_paths <- list.files(path = lahman2015_files, 
                                    full.names = TRUE, 
                                    pattern = ".csv$")
master_2015_path <- lahman_2015_csv_paths[grepl(pattern = "Master", lahman_2015_csv_paths)]
master_2015 <- data.table::fread(input = master_2015_path)
glimpse(master_2015)
```

Assume we want to create a new join column for unique rows of `nameFirst`, `nameLast`, and `birthYear`.

```{r name_year_id}
join_master_2010 <- create_join_column(df = master_2010, 
  join_cols = c("nameFirst", "nameLast", "birthYear"), 
  new_col_name = "name_birth_year_id")

join_master_2015 <- create_join_column(df = master_2015, 
  join_cols = c("nameFirst", "nameLast", "birthYear"), 
  new_col_name = "name_birth_year_id")

anti_join(
  # return all rows from join_master_2010 WITHOUT a match in join_master_2015
  x = join_master_2010, 
  y = join_master_2015, 
  by = "name_birth_year_id") %>% glimpse()
```

These are the rows in `join_master_2010` that aren't in `join_master_2015`. 
