---
title: "create-new-data"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{create-new-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

The goal of `dfdiffs` is to answer the following questions:

1. *What rows are here now that weren't here before?*
2. What rows were here before that aren't here now?   
3. What values have been changed?   

This vignette takes us through the `create_new_data()` function, which answers the question, "*What rows are here now that weren't here before?*"

## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(dplyr)
library(stringr)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
```

## What rows are here now that weren't here before? 

We're going to use two test datasets to demonstrate the `create_new_data()` function.

### Base Data 

The `T1Data` below contains six rows and eight variables.

```{r Original}
T1Data <- dfdiffs::T1Data
glimpse(T1Data)
```

The unique identifier in this dataset is the combination of subject and record (which we can see below):

```{r distinct-subject-record}
distinct(T1Data, subject)
distinct(T1Data, subject, record)
```


### Compare Data 

We'll compare `T2Data` to the original data. `T2Data` has the same six rows as `T1Data`, plus three additional rows.

```{r T2Data}
T2Data <- dfdiffs::T2Data
glimpse(T2Data)
```

The unique identifier in this dataset is the combination of subject and record (which we can see below):

```{r distinct-subject-record-T2Data}
distinct(T2Data, subject)
distinct(T2Data, subject, record)
```

### Creating a unique identifier

We also need a function that lets users specify the joining variables. It creates a new `join_var` from the supplied columns that together form a unique identifier in each dataset.

For example, we can create a unique identifier named `join_var` in `T1Data` and `T2Data`:

```{r T1DataJoin-T2DataJoin}
T1DataJoin <- mutate(T1Data, 
  join_var = as.character(row_number())) %>% 
  dplyr::relocate(join_var, everything())
T2DataJoin <- mutate(T2Data, 
  join_var = as.character(row_number())) %>% 
  dplyr::relocate(join_var, everything())
```

```{r T1DataJoin, eval=FALSE}
T1DataJoin
```

```{r show-T1DataJoin, echo=FALSE}
knitr::kable(T1DataJoin)
```

```{r T2DataJoin, eval=FALSE}
T2DataJoin
```

```{r show-T2DataJoin, echo=FALSE}
knitr::kable(T2DataJoin)
```

## Conditions 

Each function in the `dfdiffs` package assumes the following conditions: 

1) Two datasets

2) Multiple columns to compare (`cols`)

3) Single `by` column

4) Single `by` column, new column name (`by_col`)

5) Single `by` column, multiple compare columns (`cols`)

6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

7) Multiple `by` columns

8) Multiple `by` columns, new column name (`by_col`)

9) Multiple `by` columns, multiple compare columns (`cols`)

10) Multiple `by` columns, a new `by_col`, and `cols`

### Single by column conditions

1) Two datasets, compare all columns:

    ```{r cond-1, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data)
    ```

2) Multiple columns to compare (`cols`): 

    ```{r cond-2, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data,
      cols = c("text_var", "factor_var"))
    ```


3) Single `by` column, no new column name

    ```{r cond-3, eval=FALSE}
    create_new_data(
      compare = T2DataJoin, 
      base = T1DataJoin, 
      by = "join_var")
    ```

4) Single `by` column, new column name (`by_col`)

    ```{r cond-4, eval=FALSE}
    create_new_data(
      compare = T2DataJoin, 
      base = T1DataJoin, 
      by = "join_var", 
      by_col = 'new_join_var')
    ```

5) Single `by` column, multiple compare columns (`cols`)

    ```{r cond-5, eval=FALSE}
    create_new_data(
      compare = T2DataJoin, 
      base = T1DataJoin, 
      by = "join_var", 
      cols = c("text_var", "factor_var", "subject", "record"))
    ```

6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

    ```{r cond-6, eval=FALSE}
    create_new_data(
    compare = T2DataJoin, 
    base = T1DataJoin, 
    # unique id
    by = "join_var", 
    # new name for id
    by_col = 'new_join_var', 
    # cols to compare
    cols = c("subject", "record", "text_var", "factor_var"))
    ```


### Multiple by column conditions

7) Multiple `by` columns

    ```{r cond-7, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data, 
      by = c('subject', 'record'))
    ```


8) Multiple `by` columns, new column name (`by_col`)

    ```{r cond-8, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data, 
      by = c('subject', 'record'),
      by_col = "new_join_col")
    ```

9) Multiple `by` columns, multiple compare columns (`cols`)

    ```{r cond-9, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data, 
      by = c('subject', 'record'),
      cols = c("subject",  "record", "factor_var", "text_var"))
    ```

10) Multiple `by` columns, a new `by_col`, and `cols`

    ```{r cond-10, eval=FALSE}
    create_new_data(
      compare = T2Data, 
      base = T1Data, 
      by = c('subject', 'record'),
      by_col = "new_join_col",
      cols = c("subject", "record", "text_var", "factor_var"))
    ```


# create_new_data()

Below is the `create_new_data()` function:

```{r args-create_new_data, eval=FALSE}
create_new_data(compare, base, by = NULL, by_col = NULL, cols = NULL)
```

```{r dfdiffs-create_new_data, eval=FALSE, echo=TRUE}
create_new_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL) {
  # convert all columns to character
  compare[] <- lapply(compare, as.character)
  base[] <- lapply(base, as.character)

  if (is.null(by) & is.null(by_col) & is.null(cols)) {
    # 1) no 'by', no 'by_col', no 'cols' -----
    new_data_join <- anti_join_base(x = compare, y = base,
                                     by = intersect(names(compare), names(base)))
    new_data <- unique(new_data_join)
  } else if (is.null(by) & is.null(by_col) & !is.null(cols)) {
    # 2) no 'by', no 'by_col', multiple compare 'cols' -----
    compare_join_cols <- select_cols(compare, cols)
    base_join_cols <- select_cols(base, cols)
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & is.null(by_col) & is.null(cols)) {
    # 3) single 'by' column ----
    new_data_join <- anti_join_base(x = compare, y = base, by = by)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & length(by_col) == 1 & is.null(cols)) {
    # 4) single 'by' column, new 'by_col' ----
    compare <- rename_join_col(compare, by = by, by_col = by_col)
    base <- rename_join_col(base, by = by, by_col = by_col)
    new_data_join <- anti_join_base(x = compare, y = base, by = by_col)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & is.null(by_col) & !is.null(cols)) {
    # 5) single 'by' column, multiple compare 'cols' ----
    compare_cols <- select_cols(compare, c(by, cols))
    base_cols <- select_cols(base, c(by, cols))
    new_data_join <- anti_join_base(x = compare_cols, y = base_cols, by = by)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & !is.null(by_col) & !is.null(cols)) {
    # 6) single 'by' column, new 'by_col', multiple compare 'cols' ----
    compare_cols <- rename_join_col(compare, by = by, by_col = by_col)
    base_cols <- rename_join_col(base, by = by, by_col = by_col)
    compare_join <- select_cols(compare_cols, c(by_col, cols))
    base_join <- select_cols(base_cols, c(by_col, cols))
    new_data_join <- anti_join_base(x = compare_join, y = base_join, by = by_col)
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & is.null(by_col) & is.null(cols)) {
    # 7) multiple 'by' ----
    # no 'by_col', no multiple compare 'cols'
    compare_join <- create_new_column(data = compare, cols = by, new_name = "join")
    base_join <- create_new_column(data = base, cols = by, new_name = "join")
    new_data_join <- anti_join_base(x = compare_join, y = base_join,
                            by = intersect(names(compare_join), names(base_join)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & !is.null(by_col) & is.null(cols)) {
    # 8) multiple 'by' and 'by_col' ----
    # no multiple compare 'cols'
    compare_join <- create_new_column(data = compare, cols = by, new_name = by_col)
    base_join <- create_new_column(data = base, cols = by, new_name = by_col)
    new_data_join <- anti_join_base(x = compare_join, y = base_join,
                            by = intersect(names(compare_join), names(base_join)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & is.null(by_col) & !is.null(cols)) {
    # 9) multiple 'by' & multiple compare 'cols' ----
    # no 'by_col'
    compare_join <- create_new_column(data = compare, cols = by, new_name = "join")
    base_join <- create_new_column(data = base, cols = by, new_name = "join")
    compare_join_cols <- select_cols(compare_join, c("join", cols))
    base_join_cols <- select_cols(base_join, c("join", cols))
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & !is.null(by_col) & !is.null(cols)) {
    # 10) multiple 'by', new 'by_col' & compare multiple 'cols' ----
    compare_join <- create_new_column(data = compare, cols = by, new_name = by_col)
    base_join <- create_new_column(data = base, cols = by, new_name = by_col)
    compare_join_cols <- select_cols(compare_join, c(by_col, cols))
    base_join_cols <- select_cols(base_join, c(by_col, cols))
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  }

  return(new_data)
}
```

The `cond-1` through `cond-10` examples above call the real, exported `dfdiffs::create_new_data()`. The source shown here is display-only (`eval=FALSE`) and no longer redefines the function, so it can't shadow the package's own implementation.

```{r create_new_data_cond_test, eval=FALSE, echo=FALSE}
create_new_data_cond_test <- function(compare, base, by = NULL, by_col = NULL, cols = NULL) {
  # convert all columns to character
  compare <- mutate(compare, across(.cols = everything(), .fns = as.character))
  base <- mutate(base, across(.cols = everything(), .fns = as.character))
  compare <- rename_join_col(compare, by = by, by_col = by_col)
  base <- rename_join_col(base, by = by, by_col = by_col)
  compare_cols <- select(compare, all_of(by_col), all_of(cols))
  base_cols <- select(base, all_of(by_col), all_of(cols))
  new_data_join <- dplyr::anti_join(x = compare_cols, y = base_cols,
                                    by = {{by_col}})
  new_data <- dplyr::distinct(new_data_join)
    return(new_data)
}
create_new_data_cond_test(
  # data 
  compare = T2DataJoin, 
  base = T1DataJoin, 
  # unique id
  by = "join_var", 
  # new name for id
  by_col = 'new_join_var', 
  # cols to compare
  cols = c("text_var", "factor_var", "subject", "record")
)
```



`compare` = The current or new dataset in the comparison 

`base` = The previous or old dataset in the comparison 

`by` = the unique identifier for joining the two tables 

`by_col` = a new name for the joining column 

`cols` = the columns to compare (if none are provided, all columns are compared)

## Call structure

`create_new_data()` relies on four helpers from the package: `anti_join_base()` and `select_cols()` (base-R replacements for `dplyr::anti_join()`/`dplyr::select()`), `rename_join_col()` (renames the join column), and `create_new_column()` (builds the join column from the `by` columns). The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-create_new_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "create_new_data")
```

```
█─create_new_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

## Single `by` column conditions

The function also needs to handle multiple conditions. Below we cover the conditions for a single `by` column (assuming each dataset already has a unique identifier). First, we'll cover a few uncommon conditions, like a missing `by` column, or a missing `by` column with specific columns selected for comparison.

### 1) Two datasets

+ No `by` columns (only two datasets)

```{r only-datasets, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data)
```

```{r only-datasets-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data)
)
```


```{r only-datasets-join-data, eval=FALSE}
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin)
```

```{r only-datasets-join-data-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin)
)
```

We can see this performs a row-by-row comparison. 

### 2) Multiple columns to compare (`cols`)

+ No `by` columns (only two datasets) and multiple compare columns (`cols`)

```{r create_new_data-no-by-no-by_col-multiple-cols, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data,
  cols = c("text_var", "factor_var"))
```

```{r create_new_data-no-by-no-by_col-multiple-cols-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data,
  cols = c("text_var", "factor_var"))
)
```

```{r create_new_data-no-by-no-by_col-multiple-cols-join-data, eval=FALSE}
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin,
  cols = c("text_var", "factor_var"))
```

```{r create_new_data-no-by-no-by_col-multiple-cols-join-data-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin,
  cols = c("text_var", "factor_var")))
```

### 3) Single `by` column

+ We can provide a single `by` column using the `T1DataJoin` and `T2DataJoin` datasets we created above.

```{r single-by-column, eval=FALSE}
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var")
```

```{r single-by-column-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var")
)
```

### 4) Single `by` column, new column name (`by_col`)

+ We can also provide a single `by` column (the unique identifier) and a new name for it with `by_col`.

```{r single-by-column-new-by_col, eval=FALSE}
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var", 
  by_col = 'new_join_var')
```


```{r single-by-column-new-by_col-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var", 
  by_col = 'new_join_var')
)
```


### 5) Single `by` column, multiple compare columns (`cols`)

+ Single `by` column and multiple compare columns (`cols`)   

```{r  create_new_data-single-by-multiple-cols, eval=FALSE}
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var", 
  cols = c("text_var", "factor_var", "subject", "record"))
```

```{r  kable-create_new_data-single-by-multiple-cols, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin, 
  base = T1DataJoin, 
  by = "join_var", 
  cols = c("text_var", "factor_var", "subject", "record"))
)
```


### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

- Single `by` column, a new name for the by column (`by_col`), and multiple compare columns (`cols`)

```{r create_new_data-multiple-by-by_col-multiple-cols, eval=FALSE}
create_new_data(
  # data 
  compare = T2DataJoin, 
  base = T1DataJoin, 
  # unique id
  by = "join_var", 
  # new name for id
  by_col = 'new_join_var', 
  # cols to compare
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r create_new_data-multiple-by-by_col-multiple-cols-show, echo=FALSE}
knitr::kable(
create_new_data(
  # data 
  compare = T2DataJoin, 
  base = T1DataJoin, 
  # unique id
  by = "join_var", 
  # new name for id
  by_col = 'new_join_var', 
  # cols to compare
  cols =  c("subject", "record", "text_var", "factor_var"))
)
```

## Multiple `by` column conditions

Next, we'll test conditions in which multiple columns are used to create a unique identifier.

### 7) Multiple `by` columns

+ Multiple `by` columns (assuming the columns create a unique identifier) 

```{r multiple-by, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'))
```

```{r multiple-by-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'))
)
```

This creates a new `join` column that combines `subject` and `record`.

### 8) Multiple `by` columns, new column name (`by_col`)

We can provide multiple `by` columns, a new `by_col`, and **no `cols`**.

```{r multiple-by-new-by_col, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  by_col = "new_join_col")
```


```{r multiple-by-new-by_col-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  by_col = "new_join_col")
)
```

### 9) Multiple `by` columns, multiple compare columns (`cols`)  

+ Multiple `by` columns and multiple compare columns (`cols`), and **no new `by_col`**.

```{r create_new_data-multiple-by-cols, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  cols = c("subject",  "record", "factor_var", "text_var"))
```


```{r create_new_datamultiple-by-cols-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  cols = c("subject",  "record", "factor_var", "text_var"))
)
```

### 10) Multiple `by` columns, a new `by_col`, and `cols`

We can provide multiple `by` columns, a new `by_col`, and multiple `cols`.

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r multiple-by-new-by_col-cols-eval, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data, 
  base = T1Data, 
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
)
```
