---
title: "create-deleted-data"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{create-deleted-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
library(vroom)
library(haven)
library(readxl)
library(shiny)
library(data.table)
library(flextable)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

The goal of the `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. *What rows were here before that aren't here now?*
3. What values have been changed?   
4. What columns have been moved?    
5. What rows have been moved?    

This vignette takes us through the `create_deleted_data()` function, which answers the "*What rows were here before that aren't here now?*"  


## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
```

## What rows were here before that aren't here now? 

We will need three datasets to test for deleted data: `CompleteData`,  `IncompleteData`, and `DeletedData`

### CompleteData 

The `CompleteData` has 9 rows and 7 column. Unique rows are identified by a combination of `subject` and `record`:

```{r CompleteData}
CompleteData <- dfdiffs::CompleteData
flextable::qflextable(CompleteData)
```

### IncompleteData 

`IncompeleteData` has 5 rows (4 have been removed)

```{r IncompleteData}
IncompleteData <- dfdiffs::IncompleteData
flextable::qflextable(IncompleteData) |> 
  flextable::set_table_properties(layout = "autofit")
```

### DeletedData 

`DeletedData` contains the 4 rows of data removed from `CompleteData` to create `IncompleteData`

```{r DeletedData}
DeletedData <- dfdiffs::DeletedData
flextable::qflextable(DeletedData, cwidth = ) |>
  flextable::set_table_properties(layout = "autofit")
```

If we check, the combination of `IncompleteData` and `DeletedData` create `CompleteData`. 

```{r all_equal-CompleteData}
dplyr::all_equal(target = bind_rows(IncompleteData, DeletedData), 
                 current = CompleteData)
```

## Conditions 

Each function in the `dfdiffs` package assumes the following conditions: 

### Single by column conditions

1) Two datasets, compare all columns:

    ```{r cond-21, eval=FALSE}
    create_deleted_data(
      compare = IncompleteData, 
      base = CompleteData)
    ```

2) Multiple columns to compare (`cols`): 

    ```{r cond-2, eval=FALSE}
    create_deleted_data(
      compare = IncompleteDataJoin, 
      base = CompleteDataJoin,
      cols = c("text_var", "factor_var"))
    ```


3) Single `by` column, no new column name

    ```{r cond-3, eval=FALSE}
    create_deleted_data(
      compare = IncompleteDataJoin, 
      base = CompleteDataJoin, 
      by = "join_var")
    ```

4) Single by column, new column name (`by_col`)

    ```{r cond-4, eval=FALSE}
    create_deleted_data(
      compare = IncompleteDataJoin, 
      base = CompleteDataJoin, 
      by = "join_var", 
      by_col = 'new_join_var')
    ```

5) Single `by` column, multiple compare columns `cols`

    ```{r cond-5, eval=FALSE}
    create_deleted_data(
      compare = IncompleteDataJoin, 
      base = CompleteDataJoin, 
      by = "join_var", 
      cols = c("subject", "record", "factor_var", "text_var"))
    ```

6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

    ```{r cond-6, eval=FALSE}
    create_deleted_data(
      # data 
      compare = IncompleteDataJoin, 
      base = CompleteDataJoin, 
      # unique id
      by = "join_var", 
      # new name for id
      by_col = 'new_join_var', 
      # cols to compare
      cols = c("subject", "record", "text_var", "factor_var"))
    ```


### Multiple by column conditions

7) Multiple `by` columns

    ```{r cond-7, eval=FALSE}
    create_deleted_data(
      compare = IncompleteData, 
      base = CompleteData, 
      by = c('subject', 'record'))
    ```


8) Multiple `by` columns, new column name (`by_col`)

    ```{r cond-8, eval=FALSE}
    create_deleted_data(
      compare = IncompleteData, 
      base = CompleteData,
      by = c('subject', 'record'),
      by_col = "new_join_col")
    ```

9) Multiple `by` columns, multiple compare columns (`cols`)

    ```{r cond-9, eval=FALSE}
    create_deleted_data(
      compare = IncompleteData, 
      base = CompleteData, 
      by = c('subject', 'record'),
      cols = c("subject",  "record", "factor_var", "text_var"))
    ```

10) Multiple `by` columns, a new `by_col`, and `cols`

    ```{r cond-10, eval=FALSE}
    create_deleted_data(
      compare = IncompleteData, 
      base = CompleteData, 
      by = c('subject', 'record'),
      by_col = "new_join_col",
      cols = c("subject", "record", "text_var", "factor_var"))
    ```


## create_new_column()

We have a small helper function to create the join variables, `create_new_column()`:

```{r create_new_column, eval=FALSE}
create_new_column(data = , cols = , new_name = )
```

We can use `create_new_column()` with `CompleteData` and `IncompleteData` to create a joining variable with `subject` and `record`:

```{r CompleteDataJoin, eval=FALSE}
CompleteDataJoin <- create_new_column(data = CompleteData, 
  cols = c("subject", "record"), 
  new_name = "join_var")
CompleteDataJoin
```

```{r CompleteDataJoin-show, echo=FALSE}
CompleteDataJoin <- create_new_column(data = CompleteData, 
  cols = c("subject", "record"), 
  new_name = "join_var")
flextable::qflextable(CompleteDataJoin) |>
  flextable::set_table_properties(layout = "autofit")
```

```{r IncompleteDataJoin, eval=FALSE}
IncompleteDataJoin <- create_new_column(data = IncompleteData, 
  cols = c("subject", "record"), 
  new_name = "join_var")
IncompleteDataJoin
```

```{r IncompleteDataJoin-show, echo=FALSE}
IncompleteDataJoin <- create_new_column(
  data = IncompleteData, 
  cols = c("subject", "record"), 
  new_name = "join_var")
flextable::qflextable(IncompleteDataJoin) |>
  flextable::set_table_properties(layout = "autofit")
```

# create_deleted_data() 

Below is our `create_deleted_data()` function, which returns a tibble of the deleted rows. 

```{r args-create_deleted_data, eval=FALSE}
create_deleted_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL)
```

## Single `by` column conditions 

The function should also be able to handle multiple conditions. Below we cover the conditions for a single `by` columns (assuming there is an existing unique identifier in each dataset). But first, we'll cover a few uncommon conditions, like a missing `by` column, or a missing `by` column and specific columns selected for comparison. 

### 1) Two datasets

+ No `by` columns (only two datasets)

```{r only-datasets, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData)
```

```{r only-datasets-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData)
) |>
  flextable::set_table_properties(layout = "autofit")
```
 

When we compare this `DeletedData`, we can see this performs a row-by-row comparison.

```{r create_deleted_data-only-datasets-show-DeletedData, echo=FALSE}
flextable::qflextable(
  arrange(DeletedData, subject)
) |>
  flextable::set_table_properties(layout = "autofit")
```


### 2) Multiple columns to compare (`cols`)

+ No `by` columns (only two datasets) and multiple compare (`cols`)  

```{r create_deleted_data-no-by-no-by_col-multiple-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin,
  cols = c("text_var", "factor_var"))
```

```{r create_deleted_data-no-by-no-by_col-multiple-cols-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin,
  cols = c("text_var", "factor_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```

When we compare this `DeletedData`, we can see the `text_var` and `factor_var` are identical.  

```{r create_deleted_data-multiple-cols-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 3) Single `by` column

+ If the tables have a joining column, like `CompleteDataJoin` and `IncompleteDataJoin`, we can supply the (`by`) joining column 

```{r create_deleted_data-single-by-column, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var")
```

```{r create_deleted_data-single-by-column-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var")
) |>
  flextable::set_table_properties(layout = "autofit")
```

When we compare this to `DeletedData`, we can see the rows are identical.

```{r create_deleted_data-single-by-column-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 4) Single `by` column, new column name (`by_col`)

+ We can also provide a single `by` column (for unique identifiers) and a new name for the `by_col`

```{r single-by-column-new-by_col, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  by_col = 'new_join_var')
```


```{r single-by-column-new-by_col-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  by_col = 'new_join_var')
) |>
  flextable::set_table_properties(layout = "autofit")
```

When we compare this to `DeletedData`, we can see the rows are identical.

```{r single-by-column-new-by_col-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 5) Single `by` column, multiple compare columns `cols`

+ Single `by` column and multiple compare columns (`cols`)  

```{r  create_deleted_data-single-by-multiple-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  cols = c("subject", "record", "factor_var", "text_var"))
```

```{r  kable-create_deleted_data-single-by-multiple-cols, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  cols = c("subject", "record", "factor_var", "text_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```

When we compare this to `DeletedData`, we can see the rows are identical.

```{r kable-create_deleted_data-single-by-multiple-cols-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

- Single `by` column, a new name for the by column (`by_col`), and multiple compare columns (`cols`)

```{r create_deleted_data-multiple-by-by_col-multiple-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  by_col = 'new_join_var', 
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r create_deleted_data-multiple-by-by_col-multiple-cols-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin, 
  base = CompleteDataJoin, 
  by = "join_var", 
  by_col = 'new_join_var', 
  cols =  c("subject", "record", "text_var", "factor_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```

When we compare this to `DeletedData`, we can see the rows are identical.

```{r create_deleted_data-multiple-by-by_col-multiple-cols-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

## Multiple `by` column conditions

Now we're going to test conditions in which there are multiple columns used to create a unique identifier. 

### 7) Multiple `by` columns

+ Multiple `by` columns (assuming the columns create a unique identifier) 

```{r multiple-by, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'))
```

```{r multiple-by-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'))
) |>
  flextable::set_table_properties(layout = "autofit")
```

This creates a new `join` column and it's a combination of `subject` and `record`, and when we compare this to `DeletedData`, we can see the rows are identical.

```{r multiple-by-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 8) Multiple `by` columns, new column name (`by_col`)

We can provide multiple `by` columns, a new `by_col`, and  **no `cols`**

```{r multiple-by-new-by_col, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData,
  by = c('subject', 'record'),
  by_col = "new_join_col")
```


```{r multiple-by-new-by_col-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'),
  by_col = "new_join_col")
) |>
  flextable::set_table_properties(layout = "autofit")
```

This creates a new `new_join_col` column and it's a combination of `subject` and `record`, and when we compare this to `DeletedData`, we can see the rows are identical.

```{r multiple-by-new-by_col-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 9) Multiple `by` columns, multiple compare columns (`cols`)  

+ Multiple `by` columns and multiple compare columns (`cols`), and **no new `by_col`**.

```{r create_deleted_data-multiple-by-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'),
  cols = c("subject",  "record", "factor_var", "text_var"))
```


```{r create_deleted_datamultiple-by-cols-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'),
  cols = c("subject",  "record", "factor_var", "text_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```

This creates a new `join` column, and it's a combination of `subject` and `record`, and when we compare this to `DeletedData`, we can see the rows are identical.

```{r create_deleted_datamultiple-by-cols-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

### 10) Multiple `by` columns, a new `by_col`, and `cols`

We can provide multiple `by` columns, new `by_col`, and multiple `cols`

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r multiple-by-new-by_col-cols-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData, 
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```

This creates a new `join` column, and it's a combination of `subject` and `record`, and when we compare this to `DeletedData`, we can see the rows are identical.

```{r multiple-by-new-by_col-cols-show-DeletedData, echo=FALSE}
flextable::qflextable(
DeletedData
) |>
  flextable::set_table_properties(layout = "autofit")
```

## App testing 

We'll test this function on the Lahman master dataset--it's a little larger data, and more similar to data that will be used in the app: 

```{r import-master}
master15 <- readxl::read_excel(
  path = "../inst/extdata/xlsx/lahman_compare.xlsx", 
  sheet = "master-2015")
master20 <- readxl::read_excel(
  path = "../inst/extdata/xlsx/lahman_compare.xlsx", 
  sheet = "master-2020")
deleted_master_15_20 <- create_deleted_data(
  compare = master20, 
  base = master15, 
  by = "playerID",
  cols = c("playerID", "nameGiven", "height", "bats", "throws"))
```

```{r deleted_master_15_20}
glimpse(deleted_master_15_20)
```


<!--
```{r old-create_deleted_data, include=FALSE, eval=FALSE, echo=FALSE}
create_deleted_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL) {
  # convert all columns to character
  compare <- mutate(compare, across(.cols = everything(), .fns = as.character))
  base <- mutate(base, across(.cols = everything(), .fns = as.character))
  if (is.null(by) & is.null(by_col) & is.null(cols)) {
     # 1) NOTHING ----
     # (no 'by', no 'by_col', and no 'cols') 
      deleted_join <- dplyr::anti_join(x = base, y = compare,
                                    by = dplyr::intersect(
                                      x = names(base),
                                      y = names(compare)))
     deleted_data <- dplyr::distinct(deleted_join)
  } else if (is.null(by) & is.null(by_col) & !is.null(cols)) {
     # 2) multiple compare columns ----
     # (no 'by', no 'by_col', and multiple compare 'cols') 
    compare_cols <- select(compare, all_of(cols))
    base_cols <- select(base, all_of(cols))
    deleted_join <- dplyr::anti_join(x = base_cols, y = compare_cols,
                                    by = dplyr::intersect(
                                      x = names(base_cols),
                                      y = names(compare_cols)))
    deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) == 1 & is.null(by_col) & is.null(cols)) {
     # 3) single 'by' column ----
     deleted_join <- dplyr::anti_join(x = base, y = compare,
                                        by = {{by}})
     deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) == 1 & !is.null(by_col) & is.null(cols)) {
    # 4) single `by` column, new `by_col` ----
    compare_cols <- rename_join_col(data = compare, by = by, by_col = by_col)
    base_cols <- rename_join_col(data = base, by = by, by_col = by_col)
    deleted_join <- dplyr::anti_join(x = base_cols, y = compare_cols,
                                        by = {{by_col}})
    deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) == 1 & is.null(by_col) & !is.null(cols)) {
    # 5) single 'by' and multiple compare 'cols' ----
    # no 'by_col'
    compare_cols <- dplyr::select(compare, all_of(by), all_of(cols))
    base_cols <- dplyr::select(base, all_of(by), all_of(cols))
    deleted_join <- dplyr::anti_join(x = base_cols, y = compare_cols,
                                    by = dplyr::intersect(
                                    x = names(base_cols),
                                    y = names(compare_cols)))
      deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) == 1 & !is.null(by_col) & !is.null(cols)) {
    # 6) multiple 'by', new 'by_col', multiple compare 'cols' ----
    base_join <- create_new_column(data = base,
                      cols = all_of(by), new_name = {{by_col}})
    compare_join <- create_new_column(data = compare,
                      cols = all_of(by), new_name = {{by_col}})
    compare_cols <- dplyr::select(compare_join, {{by_col}}, all_of(cols))
    base_cols <- dplyr::select(base_join, {{by_col}}, all_of(cols))
    deleted_join <- dplyr::anti_join(x = base_cols, y = compare_cols,
                                      by = dplyr::intersect(
                                      x = names(base_cols),
                                      y = names(compare_cols)))
    deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) > 1 & is.null(by_col) & is.null(cols)) {
      # 7) multiple 'by', no 'by_col', no 'cols' -----
      base_join <- create_new_column(data = base,
                        cols = all_of(by), new_name = "join")
      compare_join <- create_new_column(data = compare,
                        cols = all_of(by), new_name = "join")
      deleted_join <- dplyr::anti_join(x = base_join, y = compare_join,
                                    by = dplyr::intersect(
                                      x = names(base_join),
                                      y = names(compare_join)))
      deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) > 1 & !is.null(by_col) & is.null(cols)) {
      # 8) multiple 'by', new column ('by_col') -----
      # no compare 'cols' 
      base_join <- create_new_column(data = base,
                        cols = all_of(by), new_name = {{by_col}})
      compare_join <- create_new_column(data = compare,
                        cols = all_of(by), new_name = {{by_col}})
      deleted_join <- dplyr::anti_join(x = base_join, y = compare_join,
                                    by = dplyr::intersect(
                                      x = names(base_join),
                                      y = names(compare_join)))
      deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) > 1 & is.null(by_col) & !is.null(cols)) {
      # 9) multiple 'by', multiple compare columns ('cols') -----
      # no 'by_col' 
      base_cols <- create_new_column(data = base,
                        cols = all_of(by), new_name = "join")
      compare_cols <- create_new_column(data = compare,
                        cols = all_of(by), new_name = "join")
      base_join <- dplyr::select(base_cols, matches("join"), all_of(cols))
      compare_join <- dplyr::select(compare_cols, matches("join"), all_of(cols))
      deleted_join <- dplyr::anti_join(x = base_join, y = compare_join,
                                    by = dplyr::intersect(
                                      x = names(base_join),
                                      y = names(compare_join)))
      deleted_data <- dplyr::distinct(deleted_join)
  } else if (length(by) > 1 & !is.null(by_col) & !is.null(cols)) {
    # 10) multiple 'by', new column ('by_col'), multiple compare 'cols' -----
      compare_cols <- create_new_column(data = compare,
                        cols = all_of(by), new_name = {{by_col}})
      base_cols <- create_new_column(data = base,
                        cols = all_of(by), new_name = {{by_col}})
      compare_join <- dplyr::select(compare_cols, {{by_col}}, all_of(cols))
      base_join <- dplyr::select(base_cols, {{by_col}}, all_of(cols))
      deleted_join <- dplyr::anti_join(x = base_join, y = compare_join,
                                        by = dplyr::intersect(
                                          x = names(base_join),
                                          y = names(compare_join)))
      deleted_data <- dplyr::distinct(deleted_join)
  }
  
  return(deleted_data)
}
```
-->
