---
title: "create-deleted-data"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{create-deleted-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

# Motivation

When we receive an updated version of a dataset, `dfdiffs` helps us answer three questions:

1. What rows are here now that weren't here before?
2. *What rows were here before that aren't here now?*
3. What values have been changed?

This vignette covers `create_deleted_data()`, which answers the second question: "*What rows were here before that aren't here now?*"


## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(flextable)
```

## What rows were here before that aren't here now?

We'll use three datasets from the package to test for deleted data: `CompleteData`, `IncompleteData`, and `DeletedData`.

### CompleteData

`CompleteData` has 9 rows and 7 columns. Each row is uniquely identified by the combination of `subject` and `record`:

```{r CompleteData}
CompleteData <- dfdiffs::CompleteData
flextable::qflextable(CompleteData)
```

### IncompleteData

`IncompleteData` is `CompleteData` with 4 rows removed, leaving 5 rows:

```{r IncompleteData}
IncompleteData <- dfdiffs::IncompleteData
flextable::qflextable(IncompleteData) |>
  flextable::set_table_properties(layout = "autofit")
```

### DeletedData

`DeletedData` contains the 4 rows removed from `CompleteData` to create `IncompleteData`, so it's the result we expect `create_deleted_data()` to return:

```{r DeletedData}
DeletedData <- dfdiffs::DeletedData
flextable::qflextable(DeletedData) |>
  flextable::set_table_properties(layout = "autofit")
```

We can confirm that combining `IncompleteData` and `DeletedData` recreates `CompleteData`:

```{r all_equal-CompleteData}
combined <- rbind(IncompleteData, DeletedData)
combined <- combined[do.call(order, combined), ]
complete <- CompleteData[do.call(order, CompleteData), ]
all.equal(combined, complete, check.attributes = FALSE)
```

## Arguments

`create_deleted_data()` takes the same `compare`, `base`, `by`, `by_col`, and `cols` arguments as `create_new_data()`. The [reference table](create-new-data.html#arguments) in that vignette lists every combination.

The only difference between the two functions is the direction of the anti join. `create_new_data()` looks for `compare` rows missing from `base`, and `create_deleted_data()` does the mirror image: it looks for `base` rows missing from `compare`.

# create_deleted_data()

The signature for `create_deleted_data()` is below. It returns a data frame of the deleted rows.

```{r args-create_deleted_data, eval=FALSE}
create_deleted_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL)
```

## Call structure

`create_deleted_data()` relies on the same four helpers as `create_new_data()`: `anti_join_base()`, `select_cols()`, `rename_join_col()` (which renames the join column), and `create_new_column()` (which builds the join column from the `by` columns, as shown in the [create-new-data](create-new-data.html#create_new_column) vignette). The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-create_deleted_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "create_deleted_data")
```

```
█─create_deleted_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

## Worked examples

### No `by` column: matching whole rows

When we supply only the two datasets, `create_deleted_data()` treats a row in `base` as deleted if no row in `compare` has identical values across every shared column. The result below matches `DeletedData`:

```{r only-datasets, eval=FALSE}
create_deleted_data(
  compare = IncompleteData,
  base = CompleteData)
```

```{r only-datasets-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData,
  base = CompleteData)
) |>
  flextable::set_table_properties(layout = "autofit")
```

### A single `by` column

First, we'll use `create_new_column()` to build a `join_var` column from `subject` and `record` in both datasets:

```{r CompleteDataJoin-IncompleteDataJoin}
CompleteDataJoin <- create_new_column(data = CompleteData,
  cols = c("subject", "record"),
  new_name = "join_var")
IncompleteDataJoin <- create_new_column(data = IncompleteData,
  cols = c("subject", "record"),
  new_name = "join_var")
```

Passing `join_var` as the `by` column matches rows on that key instead of on row position:

```{r single-by-column, eval=FALSE}
create_deleted_data(
  compare = IncompleteDataJoin,
  base = CompleteDataJoin,
  by = "join_var")
```

```{r single-by-column-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteDataJoin,
  base = CompleteDataJoin,
  by = "join_var")
) |>
  flextable::set_table_properties(layout = "autofit")
```

### Multiple `by` columns, a new `by_col`, and `cols`

This is the most complete case. The `by` columns are combined into a single key named with `by_col`, and only the columns in `cols` are compared:

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_deleted_data(
  compare = IncompleteData,
  base = CompleteData,
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r multiple-by-new-by_col-cols-show, echo=FALSE}
flextable::qflextable(
create_deleted_data(
  compare = IncompleteData,
  base = CompleteData,
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
) |>
  flextable::set_table_properties(layout = "autofit")
```
