---
title: "create-new-data"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{create-new-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

When we receive an updated version of a dataset, `dfdiffs` helps us answer three questions:

1. *What rows are here now that weren't here before?*
2. What rows were here before that aren't here now?
3. What values have been changed?

This vignette covers `create_new_data()`, which answers the first question: "*What rows are here now that weren't here before?*"

## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(dfdiffs)
library(dplyr)
```

## What rows are here now that weren't here before? 

We'll use two test datasets from the package to demonstrate `create_new_data()`.

### Base Data

`T1Data` is our base (original) dataset. It contains six rows and eight variables.

```{r Original}
T1Data <- dfdiffs::T1Data
glimpse(T1Data)
```

Each row in `T1Data` is uniquely identified by the combination of `subject` and `record`, which we can confirm below:

```{r distinct-subject-record}
distinct(T1Data, subject)
distinct(T1Data, subject, record)
```


### Compare Data 

`T2Data` is the dataset we'll compare against `T1Data`. It has the same six rows as `T1Data`, plus three additional rows.

```{r T2Data}
T2Data <- dfdiffs::T2Data
glimpse(T2Data)
```

The same `subject` and `record` combination uniquely identifies each row in `T2Data`:

```{r distinct-subject-record-T2Data}
distinct(T2Data, subject)
distinct(T2Data, subject, record)
```

### Creating a unique identifier

Matching rows between two datasets requires a joining variable that uniquely identifies each row. The simplest option is a row number, so we'll start by adding a `join_var` column to both `T1Data` and `T2Data` and moving it to the front:

```{r T1DataJoin-T2DataJoin}
T1DataJoin <- mutate(T1Data, 
  join_var = as.character(row_number())) %>% 
  dplyr::relocate(join_var, everything())
T2DataJoin <- mutate(T2Data, 
  join_var = as.character(row_number())) %>% 
  dplyr::relocate(join_var, everything())
```

Both datasets now start with a `join_var` column:

```{r T1DataJoin, eval=FALSE}
T1DataJoin
```

```{r show-T1DataJoin, echo=FALSE}
knitr::kable(T1DataJoin)
```

```{r T2DataJoin, eval=FALSE}
T2DataJoin
```

```{r show-T2DataJoin, echo=FALSE}
knitr::kable(T2DataJoin)
```

## `create_new_column()`

`T1DataJoin` and `T2DataJoin` use a plain row number as the join key. When the unique identifier is a *combination* of existing columns (like `subject` and `record`), `create_new_column()` pastes those columns together into a single new column. The function takes the data, the columns to combine, and a name for the new column:

```{r create_new_column, eval=FALSE}
create_new_column(data = , cols = , new_name = )
```

Below we combine `subject` and `record` into `join_var` for both datasets:

```{r T1DataSubjRec-T2DataSubjRec}
T1DataSubjRec <- create_new_column(data = T1Data,
  cols = c("subject", "record"),
  new_name = "join_var")
T2DataSubjRec <- create_new_column(data = T2Data,
  cols = c("subject", "record"),
  new_name = "join_var")
```

```{r show-T1DataSubjRec, echo=FALSE}
knitr::kable(T1DataSubjRec)
```

`create_new_data()`, `create_deleted_data()`, `create_changed_data()`, and `create_modified_data()` all call this same helper internally whenever `by` has more than one column (see the call trees below and in [create-deleted-data](create-deleted-data.html#call-structure)).

## Arguments

`create_new_data()` takes five arguments, and every comparison function in the package shares them: `compare`, `base`, `by`, `by_col`, and `cols`. The last three are optional, and the combination we supply determines how rows are matched and which columns are compared:

| `by` | `by_col` | `cols` | What happens |
|:---|:---|:---|:---|
| (none) | (none) | (none) | rows are matched on their values across every shared column (`create_changed_data()` and `create_modified_data()` match by row position instead) |
| (none) | (none) | given | the same matching, restricted to `cols` |
| single column | (none) | (none) | join on `by`, compare every shared column |
| single column | given | (none) | join, renaming the key column to `by_col` |
| single column | (none) | given | join on `by`, compare only `cols` |
| single column | given | given | join renamed to `by_col`, compare only `cols` |
| multiple columns | (none) | (none) | `by` columns combined into one `join` key |
| multiple columns | given | (none) | `by` columns combined into a key named `by_col` |
| multiple columns | (none) | given | combined key, compare only `cols` |
| multiple columns | given | given | combined key named `by_col`, compare only `cols` |

The worked examples below cover three of these cases: no `by` column, a single `by` column, and multiple `by` columns with `by_col` and `cols`. Every other row in the table is a combination of the same pieces.

# create_new_data()

The signature and source for `create_new_data()` are below:

```{r args-create_new_data, eval=FALSE}
create_new_data(compare, base, by = NULL, by_col = NULL, cols = NULL)
```

```{r dfdiffs-create_new_data, eval=FALSE, echo=TRUE}
create_new_data <- function(compare, base, by = NULL, by_col = NULL, cols = NULL) {
  # convert all columns to character
  compare[] <- lapply(compare, as.character)
  base[] <- lapply(base, as.character)

  if (is.null(by) & is.null(by_col) & is.null(cols)) {
    # 1) no 'by', no 'by_col', no 'cols' -----
    new_data_join <- anti_join_base(x = compare, y = base,
                                     by = intersect(names(compare), names(base)))
    new_data <- unique(new_data_join)
  } else if (is.null(by) & is.null(by_col) & !is.null(cols)) {
    # 2) no 'by', no 'by_col', multiple compare 'cols' -----
    compare_join_cols <- select_cols(compare, cols)
    base_join_cols <- select_cols(base, cols)
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & is.null(by_col) & is.null(cols)) {
    # 3) single 'by' column ----
    new_data_join <- anti_join_base(x = compare, y = base, by = by)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & length(by_col) == 1 & is.null(cols)) {
    # 4) single 'by' column, new 'by_col' ----
    compare <- rename_join_col(compare, by = by, by_col = by_col)
    base <- rename_join_col(base, by = by, by_col = by_col)
    new_data_join <- anti_join_base(x = compare, y = base, by = by_col)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & is.null(by_col) & !is.null(cols)) {
    # 5) single 'by' column, multiple compare 'cols' ----
    compare_cols <- select_cols(compare, c(by, cols))
    base_cols <- select_cols(base, c(by, cols))
    new_data_join <- anti_join_base(x = compare_cols, y = base_cols, by = by)
    new_data <- unique(new_data_join)
  } else if (length(by) == 1 & !is.null(by_col) & !is.null(cols)) {
    # 6) single 'by' column, new 'by_col', multiple compare 'cols' ----
    compare_cols <- rename_join_col(compare, by = by, by_col = by_col)
    base_cols <- rename_join_col(base, by = by, by_col = by_col)
    compare_join <- select_cols(compare_cols, c(by_col, cols))
    base_join <- select_cols(base_cols, c(by_col, cols))
    new_data_join <- anti_join_base(x = compare_join, y = base_join, by = by_col)
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & is.null(by_col) & is.null(cols)) {
    # 7) multiple 'by' ----
    # no 'by_col', no multiple compare 'cols'
    compare_join <- create_new_column(data = compare, cols = by, new_name = "join")
    base_join <- create_new_column(data = base, cols = by, new_name = "join")
    new_data_join <- anti_join_base(x = compare_join, y = base_join,
                            by = intersect(names(compare_join), names(base_join)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & !is.null(by_col) & is.null(cols)) {
    # 8) multiple 'by' and 'by_col' ----
    # no multiple compare 'cols'
    compare_join <- create_new_column(data = compare, cols = by, new_name = by_col)
    base_join <- create_new_column(data = base, cols = by, new_name = by_col)
    new_data_join <- anti_join_base(x = compare_join, y = base_join,
                            by = intersect(names(compare_join), names(base_join)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & is.null(by_col) & !is.null(cols)) {
    # 9) multiple 'by' & multiple compare 'cols' ----
    # no 'by_col'
    compare_join <- create_new_column(data = compare, cols = by, new_name = "join")
    base_join <- create_new_column(data = base, cols = by, new_name = "join")
    compare_join_cols <- select_cols(compare_join, c("join", cols))
    base_join_cols <- select_cols(base_join, c("join", cols))
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  } else if (length(by) > 1 & !is.null(by_col) & !is.null(cols)) {
    # 10) multiple 'by', new 'by_col' & compare multiple 'cols' ----
    compare_join <- create_new_column(data = compare, cols = by, new_name = by_col)
    base_join <- create_new_column(data = base, cols = by, new_name = by_col)
    compare_join_cols <- select_cols(compare_join, c(by_col, cols))
    base_join_cols <- select_cols(base_join, c(by_col, cols))
    new_data_join <- anti_join_base(x = compare_join_cols, y = base_join_cols,
                            by = intersect(names(compare_join_cols), names(base_join_cols)))
    new_data <- unique(new_data_join)
  }

  return(new_data)
}
```

The source above is for display only (`eval=FALSE`). The examples below call the exported `dfdiffs::create_new_data()`.

`compare` = the current (new) dataset in the comparison

`base` = the previous (old) dataset in the comparison

`by` = the unique identifier for joining the two tables

`by_col` = a new name for the joining column

`cols` = the columns to compare (if none are provided, all columns are compared)

## Call structure

`create_new_data()` relies on four helpers from the package. `anti_join_base()` and `select_cols()` are base R replacements for `dplyr::anti_join()` and `dplyr::select()`, `rename_join_col()` renames the join column, and `create_new_column()` builds the join column from the `by` columns.

The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-create_new_data, eval=FALSE}
stackcallr::call_tree_dir("R", root = "create_new_data")
```

```
█─create_new_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

## Worked examples

### No `by` column: matching whole rows

When we supply only the two datasets, `create_new_data()` treats a row in `compare` as new if no row in `base` has identical values across every shared column:

```{r only-datasets, eval=FALSE}
create_new_data(
  compare = T2Data,
  base = T1Data)
```

```{r only-datasets-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data,
  base = T1Data)
)
```

### A single `by` column

`T1DataJoin` and `T2DataJoin` both contain the `join_var` identifier we created above. Passing it as a single `by` column matches rows on that key instead of on row position:

```{r single-by-column, eval=FALSE}
create_new_data(
  compare = T2DataJoin,
  base = T1DataJoin,
  by = "join_var")
```

```{r single-by-column-show, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2DataJoin,
  base = T1DataJoin,
  by = "join_var")
)
```

### Multiple `by` columns, a new `by_col`, and `cols`

This is the most complete case. The `by` columns are combined into a single key named with `by_col`, and only the columns in `cols` are compared:

```{r multiple-by-new-by_col-cols, eval=FALSE}
create_new_data(
  compare = T2Data,
  base = T1Data,
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
```

```{r multiple-by-new-by_col-cols-eval, echo=FALSE}
knitr::kable(
create_new_data(
  compare = T2Data,
  base = T1Data,
  by = c('subject', 'record'),
  by_col = "new_join_col",
  cols = c("subject", "record", "text_var", "factor_var"))
)
```
