---
title: "create-join-column"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create-join-column}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  str = strOptions(strict.width = "wrap"),
  width = 75,
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

# Motivation

Comparing two datasets requires a way to match each row in one to its counterpart in the other. When no single column identifies a row, we need to combine several columns into one. This vignette walks through `create_join_column()`, the function the `dfdiffs` Shiny app uses to build that combined 'join' column.

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(data.table)
```

## create_join_column()

`create_join_column()` combines any number of columns in `df` into a single join column. The columns to combine go in `by_colums`, and `new_by_column_name` sets the name of the new column.

```{r create_join_column, eval=FALSE}
create_join_column(df, by_colums, new_by_column_name)
```

## Call structure

`create_join_column()` calls one other `dfdiffs` function, `select_cols()`, to pull out the columns being combined before pasting them together. Inside the Shiny app, the select module (`mod_select_server()`) calls `create_join_column()` to build the `join_column` from the columns chosen in the *Select Join Columns* panel. The call tree below was generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-mod_select_server, eval=FALSE}
stackcallr::call_tree_dir("R", root = "mod_select_server")
```

```
█─mod_select_server
├─select_cols
├─base_react_theme
├─comp_react_theme
├─info_react_theme
└─█─create_join_column
  └─select_cols
```

## Test data

To show how `create_join_column()` works, we'll load two pulls of the synthetic (not real) `dfdiffs` site roster (see `?Roster2021`).

### Roster pull (2021)

The 2021 roster pull is stored in `inst/extdata/csv/site-roster/2021/`. The code below lists the files in that folder.

```{r lahman2010_files}
roster2021_files <- "../inst/extdata/csv/site-roster/2021/"
fs::dir_tree(roster2021_files)
```

Next, we'll locate the `Roster` file and read it with `data.table::fread()`.

```{r master_2010}
roster2021_csv_paths <- list.files(
  path = roster2021_files,
  full.names = TRUE,
  pattern = ".csv$")
roster2021_path <- roster2021_csv_paths[grepl(pattern = "Roster", roster2021_csv_paths)]
roster_2021 <- data.table::fread(input = roster2021_path)
str(roster_2021)
```

### Roster pull (2022)

The 2022 roster pull follows the same layout in `inst/extdata/csv/site-roster/2022/`.

```{r lahman2015_files}
roster2022_files <- "../inst/extdata/csv/site-roster/2022/"
fs::dir_tree(roster2022_files)
```

We'll read the 2022 `Roster` file the same way.

```{r master_2015}
roster2022_csv_paths <- list.files(path = roster2022_files,
                                    full.names = TRUE,
                                    pattern = ".csv$")
roster2022_path <- roster2022_csv_paths[grepl(pattern = "Roster", roster2022_csv_paths)]
roster_2022 <- data.table::fread(input = roster2022_path)
str(roster_2022)
```

Suppose we want a join column that identifies unique rows by `first_name`, `last_name`, and `enroll_year`. The code below creates `name_year_id` in both roster pulls, then passes them to `anti_join_base()` to find the 2021 rows with no match in 2022.

```{r name_year_id}
join_roster_2021 <- create_join_column(df = roster_2021,
  by_colums = c("first_name", "last_name", "enroll_year"),
  new_by_column_name = "name_year_id")

join_roster_2022 <- create_join_column(df = roster_2022,
  by_colums = c("first_name", "last_name", "enroll_year"),
  new_by_column_name = "name_year_id")

anti_join_base(
  # return all rows from join_roster_2021 WITHOUT a match in join_roster_2022
  x = join_roster_2021,
  y = join_roster_2022,
  by = "name_year_id") |> str()
```

The output above shows the rows in `join_roster_2021` that aren't in `join_roster_2022`. Because `name_year_id` combines three columns, a row only matches when the first name, last name, and enrollment year all agree.
