---
title: "function-checks"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{function-checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
# id <- as.integer(c(c(7:12), c(7:9)))
# group <- factor(rep(c("D", "E", "F"), times  = 3), levels = c("D", "E", "F"))
# var_x <- round(rnorm(n = 9, mean = 50, sd = 15), 1)
# var_y <- as.integer(round(rnorm(n = 9, mean = 300, sd = 120), 0))
# var_z <- sample(x = str_to_upper(words), size = 9, replace = FALSE)
```

```{r setup, eval=TRUE}
library(dfdiffs)
```


# Motivation

This vignette takes us through the functions used to compare datasets in the `dfdiffs` package. The goal  of `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   


## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide', eval=TRUE}
library(shiny)
library(data.table)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
library(gt)
```

```{r data, eval=TRUE, echo=FALSE}
# new data
compare_new_df <- tibble::tribble(
  ~id, ~group, ~var_x, ~var_y,     ~var_z,
    1,    "A",    5.5,   111L,   "notice",
    2,    "B",      9,   106L,    "admit",
    3,    "C",    5.2,   109L,  "america",
    4,    "A",    4.4,    96L,      "try",
    5,    "B",    9.1,   116L,    "serve",
    6,    "C",    6.7,    98L,  "because",
    1,    "A",    2.8,    97L,   "unless",
    2,    "B",      8,   113L,  "believe",
    3,    "C",    6.8,   111L,    "union"
  )
base_new_df <- tibble::tribble(
   ~id, ~group, ~var_x, ~var_y,    ~var_z,
     1,    "A",    5.5,   111L,  "notice",
     2,    "B",      9,   106L,   "admit",
     3,    "C",    5.2,   109L, "america",
     4,    "A",    4.4,    96L,     "try",
     5,    "B",    9.1,   116L,   "serve",
     6,    "C",    6.7,    98L, "because"
   )
# deleted data
base_deleted_df <- tibble::tribble(
  ~id, ~group, ~var_x, ~var_y,     ~var_z,
   7L,    "D",   38.3,   207L,   "APPEAR",
   8L,    "E",   72.8,   383L,     "SAVE",
   9L,    "F",     57,   274L,     "EAST",
  10L,    "D",   43.1,   307L,  "SIMILAR",
  11L,    "E",     40,   305L,     "STEP",
  12L,    "F",   38.7,   168L, "CONSIDER",
   7L,    "D",   43.3,   338L,     "EVER",
   8L,    "E",   49.6,   347L,   "EFFECT",
   9L,    "F",   58.7,   478L,      "BIT"
  )
compare_deleted_df <- tibble::tribble(
  ~id, ~group, ~var_x, ~var_y,     ~var_z,
   7L,    "D",   38.3,   207L,   "APPEAR",
   8L,    "E",   72.8,   383L,     "SAVE",
   9L,    "F",     57,   274L,     "EAST",
  10L,    "D",   43.1,   307L,  "SIMILAR",
  11L,    "E",     40,   305L,     "STEP",
  12L,    "F",   38.7,   168L, "CONSIDER"
  )
# changed data 
base_change_df <- tibble::tribble(
  ~id, ~group, ~var_x, ~var_y,     ~var_z,
    1,    "A",    5.5,   111L,   "notice",
    2,    "B",      9,   106L,    "admit",
    3,    "C",    5.2,   109L,  "america",
    4,    "A",    4.4,    96L,      "try",
    5,    "B",    9.1,   116L,    "serve",
    6,    "C",    6.7,    98L,  "because",
  )
compare_change_df <- tibble::tribble(
  ~id, ~group, ~var_x, ~var_y,     ~var_z,
    1,    "A",    5.5,   111L,   "notice",
    2,    "B",    3.3,   106L,    "admit",
    3,    "C",    5.2,   109L,  "America",
    4,    "A",    4.4,    73L,      NA_character_,
    5,    "B",    9.1,   116L,    "serve",
    6,    "C",    6.7,    98L,  "HiStorY",
  )
```

## New Data

Below are the test `base` and `compare` new datasets: `base_new_df` and `compare_new_df`

### Base 

```{r base_new_df}
base_new_df |> gt::gt(data = )
```

### Compare 

```{r compare_new_df}
compare_new_df |> gt::gt(data = )
```


## Deleted Data 

Below are the test `base` and `compare` deleted datasets: `base_deleted_df` and `compare_deleted_df`

### Base 

```{r base_deleted_df}
base_deleted_df |> gt::gt(data = )
```

### Compare 

```{r compare_deleted_df}
compare_deleted_df |> gt::gt(data = )
```

## Changed Data 

Below are the test `base` and `compare` changed datasets: `base_change_df` and `compare_change_df`

### Base

```{r base_change_df}
base_change_df |> gt::gt(data = )
```

### Compare 

```{r compare_change_df}
compare_change_df |> gt::gt(data = )
```


## Single joining (`by`) column conditions 

### 1) Two datasets

+ No `by` columns (only two datasets)

```{r test-create_new_data-1}
create_new_data(
   compare = compare_new_df,
   base = base_new_df) |> gt::gt(data = )
```

```{r test-create_deleted_data-1}
create_deleted_data(
   compare = compare_deleted_df,
   base = base_deleted_df) |> gt::gt(data = )
```

```{r test-create_changed_data-1}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df)
```

```{r num_diffs-1}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-1}
changes$var_diffs |> gt::gt(data = )
```

### 2) Multiple columns to compare (`cols`)

+ No `by` columns (only two datasets) and multiple compare (`cols`)  

```{r test-create_new_data-2}
create_new_data(
   compare = compare_new_df,
   base = base_new_df, 
   cols = c('var_x', 'var_z')) |> 
  gt::gt(data = )
```


```{r test-create_deleted_data-2}
create_deleted_data(
   compare = compare_deleted_df,
   base = base_deleted_df, 
   cols = c('var_x', 'var_y', 'var_z')) |> 
  gt::gt(data = )
```

```{r test-create_changed_data-2}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df, 
  cols = c('var_y', 'var_z'))
```

```{r num_diffs-2}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-2}
changes$var_diffs |> gt::gt(data = )
```

### 3) Single `by` column

+ Provide a single `by` column (in this case, we will create the join column using `row_number()`)

```{r test-create_new_data-3}
compare <- mutate(compare_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_new_data(
   compare = compare,
   base =  base,
   by = "join_var") |> gt::gt(data = )
```

```{r test-create_deleted_data-3}
compare <- mutate(compare_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_deleted_data(
   compare = compare,
   base = base, 
   by = 'join_var') |> 
  gt::gt(data = )
```


```{r test-create_changed_data-3}
compare <- mutate(compare_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
changes <- create_changed_data(
  compare = compare, 
  base = base, 
  by = 'join_var')
```

```{r num_diffs-3}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-3}
changes$var_diffs |> gt::gt(data = )
```


### 4) Single `by` column, new column name (`by_col`)

+ We can also provide a single `by` column (for unique identifiers) and a new name for the `by_col` 

+ We will create the join column (`join_var`) using `row_number()`, then name it `join_column`

```{r test-create_new_data-4}
compare <- mutate(compare_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_new_data(
   compare = compare,
   base =  base,
   by = "join_var", 
   by_col = 'join_column') |> gt::gt(data = )
```


```{r test-create_deleted_data-4}
compare <- mutate(compare_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_deleted_data(
   compare = compare,
   base = base, 
   by = 'join_var', 
   by_col = 'join_column') |> 
  gt::gt(data = )
```

```{r test-create_changed_data-4}
compare <- mutate(compare_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
changes <- create_changed_data(
  compare = compare, 
  base = base, 
  by = 'join_var', 
  by_col = 'join_column')
```

```{r num_diffs-4}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-4}
changes$var_diffs |> gt::gt(data = )
```

### 5) Single `by` column, multiple compare columns `cols`

+ Single `by` column and multiple compare columns (`cols`) 

```{r test-create_new_data-5}
compare <- mutate(compare_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_new_data(
   compare = compare,
   base =  base,
   by = "join_var", 
   cols = c('var_x', 'var_z')) |> gt::gt(data = )
```


```{r test-create_deleted_data-5}
compare <- mutate(compare_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_deleted_data(
   compare = compare,
   base = base, 
   by = 'join_var', 
   cols = c('var_x', 'var_z')) |> 
  gt::gt(data = )
```


```{r test-create_changed_data-5}
compare <- mutate(compare_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
changes <- create_changed_data(
  compare = compare, 
  base = base, 
  by = 'join_var', 
  cols = c('var_x', 'var_z'))
```

```{r num_diffs-5}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-5}
changes$var_diffs |> gt::gt(data = )
```


### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

- Single `by` column, a new name for the by column (`by_col`), and multiple compare columns (`cols`)

```{r test-create_new_data-6}
compare <- mutate(compare_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_new_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_new_data(
   compare = compare,
   base =  base,
   by = "join_var", 
   by_col = "join_column",
   cols = c('var_x', 'var_z')) |> gt::gt(data = )
```

```{r test-create_deleted_data-6}
compare <- mutate(compare_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_deleted_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
create_deleted_data(
   compare = compare,
   base = base, 
   by = 'join_var', 
   by_col = "join_column",
   cols = c('var_x', 'var_z')) |> 
  gt::gt(data = )
```

```{r test-create_changed_data-6}
compare <- mutate(compare_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
base <- mutate(base_change_df, 
  join_var = as.character(row_number())) |>  
  dplyr::relocate(join_var, everything())
changes <- create_changed_data(
  compare = compare, 
  base = base, 
  by = 'join_var', 
  by_col = "join_column",
  cols = c('var_x', 'var_z'))
```

```{r num_diffs-6}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-6}
changes$var_diffs |> gt::gt(data = )
```

## Multiple `by` column conditions

### 7) Multiple `by` columns

+ Multiple `by` columns (assuming the columns create a unique identifier).

```{r test-create_new_data-7}
create_new_data(
   compare = compare_new_df,
   base = base_new_df,
   by = c("id", "group")) |> gt::gt(data = )
```

```{r test-create_deleted_data-7}
create_deleted_data(
  compare = compare_deleted_df,
  base = base_deleted_df,
  by = c("id", "group")) |> gt::gt(data = )
```

```{r test-create_changed_data-7}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df,
  by = c("id", "group"))
```

```{r num_diffs-7}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-7}
changes$var_diffs |> gt::gt(data = )
```


### 8) Multiple `by` columns, new column name (`by_col`)

We can provide multiple `by` columns, a new `by_col`, and  **no `cols`**

```{r test-create_new_data-8}
create_new_data(
   compare = compare_new_df,
   base = base_new_df,
   by = c("id", "group"),
   by_col = "join_column") |> 
   gt::gt(data = )
```

```{r test-create_deleted_data-8}
create_deleted_data(
  compare = compare_deleted_df,
  base = base_deleted_df,
  by = c("id", "group"),
  by_col = "join_column") |> gt::gt(data = )
```

```{r test-create_changed_data-8}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df,
  by = c("id", "group"),
  by_col = "join_column")
```

```{r num_diffs-8}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-8}
changes$var_diffs |> gt::gt(data = )
```

### 9) Multiple `by` columns, multiple compare columns (`cols`)  

+ Multiple `by` columns and multiple compare columns (`cols`), and **no new `by_col`**.

```{r test-create_new_data-9}
create_new_data(
   compare = compare_new_df,
   base = base_new_df,
   by = c("id", "group"),
   cols = c('var_x', 'var_z')) |> 
   gt::gt(data = )
```

```{r test-create_deleted_data-9}
create_deleted_data(
  compare = compare_deleted_df,
  base = base_deleted_df,
  by = c("id", "group"),
  cols = c('var_x', 'var_z')) |> 
  gt::gt(data = )
```


```{r test-create_changed_data-9}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df,
  by = c("id", "group"),
  cols = c('var_x', 'var_z'))
```

```{r num_diffs-9}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-9}
changes$var_diffs |> gt::gt(data = )
```


### 10) Multiple `by` columns, a new `by_col`, and `cols`

We can provide multiple `by` columns, new `by_col`, and multiple `cols`.

```{r test-create_new_data-10}
create_new_data(
   compare = compare_new_df,
   base = base_new_df,
   by = c("id", "group"),
   by_col = "join_column",
   cols = c('var_x', 'var_z')) |> 
   gt::gt(data = )
```

```{r test-create_deleted_data-10}
create_deleted_data(
  compare = compare_deleted_df,
  base = base_deleted_df,
  by = c("id", "group"),
  by_col = "join_column",
  cols = c('var_x', 'var_z')) |> 
  gt::gt(data = )
```


```{r test-create_changed_data-10}
changes <- create_changed_data(
  compare = compare_change_df, 
  base = base_change_df,
  by = c("id", "group"),
  by_col = "join_column",
  cols = c('var_x', 'var_z'))
```

```{r num_diffs-10}
changes$num_diffs |> gt::gt(data = )
```

```{r var_diffs-10}
changes$var_diffs |> gt::gt(data = )
```
