---
title: "function-checks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{function-checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dfdiffs)
```


# Motivation

This vignette takes us through some of the functions used to compare datasets in the `dfdiffs` package. The goal  of `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   
4. What columns have been moved?    
5. What rows have been moved?    


## Packages

```{r pkgs, message=FALSE, warning=FALSE, results='hide'}
library(shiny)
library(data.table)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## Single joining (`by`) column conditions 

### 1) Two datasets

+ No `by` columns (only two datasets)

### 2) Multiple columns to compare (`cols`)

+ No `by` columns (only two datasets) and multiple compare (`cols`)  

### 3) Single `by` column

+ We can provide a single `by` column (using our `T1DataJoin` and `T2DataJoin`) datasets we created above. 

### 4) Single `by` column, new column name (`by_col`)

+ We can also provide a single `by` column (for unique identifiers) and a new name for the `by_col`

### 5) Single `by` column, multiple compare columns `cols`

+ Single `by` column and multiple compare columns (`cols`) 

### 6) Single `by` column, new column name (`by_col`), multiple compare columns (`cols`)

- Single `by` column, a new name for the by column (`by_col`), and multiple compare columns (`cols`)


## Multiple `by` column conditions

### 7) Multiple `by` columns

+ Multiple `by` columns (assuming the columns create a unique identifier).

### 8) Multiple `by` columns, new column name (`by_col`)

We can provide multiple `by` columns, a new `by_col`, and  **no `cols`**

### 9) Multiple `by` columns, multiple compare columns (`cols`)  

+ Multiple `by` columns and multiple compare columns (`cols`), and **no new `by_col`**.

### 10) Multiple `by` columns, a new `by_col`, and `cols`

We can provide multiple `by` columns, new `by_col`, and multiple `cols`.

***

# Loading data 

Below we load three datasets from the [`Lahman` package](https://cran.r-project.org/web/packages/Lahman/index.html)

## 2010 Lahamn 'Master'

```{r lahman_2010_master_path}
lahman_2010_master_path <- list.files(
  path = "../inst/extdata/csv/2010-lahman", 
  full.names = TRUE, 
  pattern = "^Master")
master_2010_raw <- load_flat_file(path = lahman_2010_master_path)
master_2010 <- mutate(master_2010_raw, 
  debut = lubridate::mdy_hms(debut),
  finalGame = lubridate::mdy_hms(finalGame)
  )
glimpse(master_2010)
```

## 2015 Lahamn 'Master'

```{r master_2015}
lahman_2015_master_path <- list.files(
  path = "../inst/extdata/csv/2015-baseballdatabank", 
  full.names = TRUE, pattern = "^Master")
master_2015 <- load_flat_file(path = lahman_2015_master_path)
glimpse(master_2015)
```

## 2020 Lahamn 'People'

The name for this dataset changed from `Master` to `People`.

```{r people_2020}
lahman_2020_people_path <- list.files(
  path = "../inst/extdata/csv/2020-baseballdatabank/", 
  full.names = TRUE, pattern = "^People")
people_2020 <- load_flat_file(path = lahman_2020_people_path)
glimpse(people_2020)
```

# New data: "*What rows are here now that weren't here before?*"

```{r create_new_data-args, eval=FALSE}
dfdiffs::create_new_data(newdf = , olddf = , by = , by_col = , cols = )
```


We can test this function by looking at the new data between 2010, 2015, and 2020 datasets. 

## New data between 2010 - 2020

The new data between 2010 and 2020 is created below. 

```{r new_2010thru2020}
new_2010thru2020 <- create_new_data(newdf = people_2020, 
                                    olddf = master_2010, 
                                    by = "playerID")
```

The most straightforward way to check if this function is working is to get the `playerID`s that are in `people_2020` but not in `master_2010`. 

```{r new_2010thru2020_playerIDs}
new_2010thru2020_playerIDs <- unique(new_2010thru2020$playerID)
length(new_2010thru2020_playerIDs)
```

Now we can check to see if these names are the identical number of rows filtered to the `playerID`s in the 2020 data:

```{r identical-new_2010thru2020-new_2010thru2020_playerIDs}
identical(x = nrow(new_2010thru2020), 
  y = nrow(filter(people_2020, playerID %in% new_2010thru2020_playerIDs)))
```


We can also check the `debut` variable in the `new_2010thru2020` data. The earliest and latest dates for `debut` in `new_2010thru2020` are below:

```{r min-max-new_2010thru2020-debut}
min(new_2010thru2020$debut, na.rm = TRUE)
max(new_2010thru2020$debut, na.rm = TRUE)
```


If we filter each dataset to for the earliest date, we find the player who was added to the 2020 data:

```{r 1873-player}
people_2020_1873 <- filter(people_2020, debut == lubridate::as_date("1873-04-14"))
master_2010_1873 <- filter(master_2010, debut == lubridate::as_date("1873-04-14"))
player_debut_1873 <- setdiff(x = people_2020_1873$playerID, 
  y = master_2010_1873$playerID)
player_debut_1873
```

Our missing player is `r player_debut_1873`. If we use this player to check the 2010 and 2020 data, we can see this player is in the 2020 dataset...

```{r check-2020-player_debut_1873}
knitr::kable(filter(people_2020, playerID == player_debut_1873) %>% 
  select(playerID, debut))
```

...and not the 2010 data:

```{r check-2010-player_debut_1873}
knitr::kable(filter(master_2010, playerID == player_debut_1873) %>% 
  select(playerID, debut))
```

## New data between 2010 - 2015

Below we perform similar checks for 2010 vs. 2015 data

```{r new_2010thru2015}
new_2010thru2015 <- create_new_data(newdf = master_2015, 
                                    olddf = master_2010, 
                                    by = "playerID")
```

```{r new_2010thru2015_playerIDs}
new_2010thru2015_playerIDs <- unique(new_2010thru2015$playerID)
# check length
length(new_2010thru2015_playerIDs)
```

Check to see if these names are the identical rows in the 2015 data

```{r identical-new_2010thru2015-new_2010thru2015_playerIDs}
identical(x = nrow(new_2010thru2015), 
  y = nrow(filter(master_2015, playerID %in% new_2010thru2015_playerIDs)))
```

## New data between 2015 - 2020

```{r new_2015thru2020}
new_2015thru2020 <- create_new_data(newdf = people_2020, 
                                    olddf = master_2015, 
                                    by = "playerID")
```

```{r new_2015thru2020_playerIDs}
new_2015thru2020_playerIDs <- unique(new_2015thru2020$playerID)
length(new_2015thru2020_playerIDs)
```

Check to see if these names are the identical rows in the 2020 data

```{r identical-new_2015thru2020-new_2015thru2020_playerIDs}
identical(x = nrow(new_2015thru2020), 
  y = nrow(filter(people_2020, playerID %in% new_2015thru2020_playerIDs)))
```

# Deleted Data 

Now that we've covered the creation of 'new data', we can move onto the 'deleted data'. 

## Deleted data: "*What rows were here before that aren't here now?*"

For the deleted datasets, we'll be using the `create_deleted_data()` function.  

```{r create_deleted_data-args, eval=FALSE}
dfdiffs::create_deleted_data(newdf = , olddf = , by = , by_col = , cols = )
```

## Deleted data between 2010 - 2020

We can test to see if any data have been removed between the 2020 and 2010 'Master/People' datasets.

```{r deleted_2010thru2020}
deleted_2010thru2020 <- create_deleted_data(
  newdf = people_2020, 
  olddf = master_2010, 
  by = "playerID")
```

Now if we check the `playerID`s from the `deleted_2010thru2020` data, they shouldn't be in the `people_2020` data. 

```{r identical-people_2020-deleted_2010thru2020_playerIDs}
deleted_2010thru2020_playerIDs <- unique(deleted_2010thru2020$playerID)
deleted_2010thru2020_playerIDs
```

And the number of IDs in the `master_2010` data should be identical to the number of rows in the `deleted_2010thru2020` dataset. 

```{r identical-deleted_2010thru2020-deleted_2010thru2020_playerIDs}
identical(
  x = nrow(filter(master_2010, playerID %in% deleted_2010thru2020_playerIDs)),
  y = nrow(deleted_2010thru2020))
```

## Deleted data between 2010 - 2015

We'll also test to see if any data have been removed between the 2015 and 2010 'Master/People' datasets.

```{r deleted_2010thru2015}
deleted_2010thru2015 <- create_deleted_data(
  newdf = master_2015, 
  olddf = master_2010, 
  by = "playerID")
# get unique ids
deleted_2010thru2015_playerIDs <- unique(deleted_2010thru2015$playerID)
deleted_2010thru2015_playerIDs
```
```{r identical-deleted_2010thru2015-deleted_2010thru2015_playerIDs}
# check identical
identical(
  x = nrow(filter(master_2010, playerID %in% deleted_2010thru2015_playerIDs)),
  y = nrow(deleted_2010thru2015))
```

## Deleted data between 2015 - 2020

```{r deleted_2015thru2020}
deleted_2015thru2020 <- create_deleted_data(
  newdf = people_2020, 
  olddf = master_2015, 
  by = "playerID")
# get unique ids
deleted_2015thru2020_playerIDs <- unique(deleted_2015thru2020$playerID)
deleted_2015thru2020_playerIDs
```


```{r identical-deleted_2015thru2020-deleted_2015thru2020_playerIDs}
# check identical
identical(
  x = nrow(filter(master_2015, playerID %in% deleted_2015thru2020_playerIDs)),
  y = nrow(deleted_2015thru2020))
```

## Modified data: "*What values have been changed?*"

We now have functions for returning new records and deleted records, but what about records that have been modified? Both `create_new_data()` and `create_deleted_data()` are built using `dplyr` joins, but to find records where specific values have changed row-by-row, we need to enlist the help of a few functions from other packages. 


Other functions for comparing datasets to each other are used in the `dfdiffs` package, and these have all been rewritten to return a `tibble`. 


