---
title: "getting-started"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

When a new version of a dataset arrives, the first thing we usually want to know is what changed since the last one. `dfdiffs` answers that with three questions:

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   

## Packages 

We'll start by loading `dfdiffs`:

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

The code below loads the other packages used in this vignette:

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(shiny)
library(data.table)
library(stringr)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(haven)
library(readxl)
library(janitor)
library(reactable)
library(labelled)
library(gtsummary)
```

## Package functions 

`dfdiffs` has a function for each question above, and each function comes with a pair of example datasets that show how it works. We'll walk through each pair below.

### Call structure

The four comparison functions are built from a few small helpers, and the same functions power the package's Shiny app. I generated the call trees in this section from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`), and they only show functions defined in `dfdiffs`.

```{r call-tree-getting-started, eval=FALSE}
library(stackcallr)
call_tree_dir("R", root = "create_new_data")
call_tree_dir("R", root = "create_deleted_data")
call_tree_dir("R", root = "create_changed_data")
call_tree_dir("R", root = "create_modified_data")
```

`create_new_data()` and `create_deleted_data()` answer the first two questions (new rows and deleted rows), and both rely on the same four helpers: `anti_join_base()`, `select_cols()`, `rename_join_col()`, and `create_new_column()`. These helpers are written in base R, so no external comparison package is involved:

```
█─create_new_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

```
█─create_deleted_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

`create_changed_data()` and `create_modified_data()` answer the third question (changed values), and both build on `diff_values_by_key()`. `diff_values_by_key()` compares matched rows column by column with `compare_values()`, which is tolerance-aware for numeric and date-time values and label-aware for factors. Any class mismatches are recorded with `class_diffs()`. The call tree for `create_changed_data()` is shown below:

```
█─create_changed_data
├─select_cols
├─█─diff_values_by_key
│ ├─compare_values
│ └─class_diffs
├─rename_join_col
├─create_new_column
└─class_diffs
```

The Shiny app (`launch_app()`) wires these functions together with three modules (upload, select, and compare), and each module has a UI function and a server function. `app_ui()` and `app_server()` assemble the modules, and `dfdiffs_fresh_theme()` supplies the app's theme. The code below generates the full call tree for the app:

```{r call-tree-launch-app, eval=FALSE}
call_tree_dir("R", root = "launch_app")
```

```
█─launch_app
├─█─app_ui
│ ├─dfdiffs_fresh_theme
│ ├─mod_upload_ui
│ ├─mod_select_ui
│ └─mod_compare_ui
└─█─app_server
  ├─█─mod_upload_server
  │ ├─█─upload_data
  │ │ └─load_flat_file
  │ ├─base_react_theme
  │ └─comp_react_theme
  ├─█─mod_select_server
  │ ├─select_cols
  │ ├─base_react_theme
  │ ├─comp_react_theme
  │ ├─info_react_theme
  │ └─█─create_join_column
  │   └─select_cols
  └─█─mod_compare_server
    ├─compare_columns
    ├─%nin%
    ├─info_react_theme
    ├─█─create_new_data
    │ ├─anti_join_base
    │ ├─select_cols
    │ ├─rename_join_col
    │ └─create_new_column
    ├─new_react_theme
    ├─█─create_deleted_data
    │ ├─anti_join_base
    │ ├─select_cols
    │ ├─rename_join_col
    │ └─create_new_column
    ├─deleted_react_theme
    ├─█─create_modified_data
    │ ├─select_cols
    │ ├─█─diff_values_by_key
    │ │ ├─compare_values
    │ │ └─class_diffs
    │ ├─rename_join_col
    │ ├─create_new_column
    │ └─class_diffs
    ├─select_cols
    ├─changed_react_theme
    ├─left_join_base
    └─█─create_comparison_report
      ├─█─create_new_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_deleted_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_modified_data
      │ ├─select_cols
      │ ├─█─diff_values_by_key
      │ │ ├─compare_values
      │ │ └─class_diffs
      │ ├─rename_join_col
      │ ├─create_new_column
      │ └─class_diffs
      ├─create_empty_tbl
      ├─compare_columns
      ├─class_diffs
      └─█─compare_summary
        └─█─compare_data
          ├─█─create_new_data
          │ ├─anti_join_base
          │ ├─select_cols
          │ ├─rename_join_col
          │ └─create_new_column
          ├─█─create_deleted_data
          │ ├─anti_join_base
          │ ├─select_cols
          │ ├─rename_join_col
          │ └─create_new_column
          ├─█─create_changed_data
          │ ├─select_cols
          │ ├─█─diff_values_by_key
          │ │ ├─compare_values
          │ │ └─class_diffs
          │ ├─rename_join_col
          │ ├─create_new_column
          │ └─class_diffs
          ├─class_diffs
          └─compare_columns
```

## *What rows are here now that weren't here before?*

To find new rows, we'll compare `T1Data` and `T2Data`, then check the result against `NewData`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
NewData <- dfdiffs::NewData
```

### Timepoint 1 data (original)

`T1Data` represents data collected at the first timepoint (T1).

**`T1Data`**: Simulated 'time-point 1' data

```{r T1Data, echo=FALSE}
reactable::reactable(T1Data, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### Timepoint 2 data (new)

`T2Data` is the 'new' dataset, representing data collected at the second timepoint (T2).

**`T2Data`**: Simulated 'time-point 2' data

```{r T2Data, echo=FALSE}
reactable::reactable(T2Data, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_new_data()`

`create_new_data()` returns the 'new data' (i.e., the rows that are here now but weren't here before). We pass the newer dataset to `compare` and the original dataset to `base`:

```{r , eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data)
```


Output from **`create_new_data()`**: Differences between 'time-point 1' and 'time-point 2'

```{r create_new_data, echo=FALSE}
create_new_data(
  compare = T2Data,
  base = T1Data) %>%
  reactable::reactable(resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

We can confirm this output is correct by checking it against the stored `NewData` dataset. The two tables should match.

**`NewData`**: stored differences from `create_new_data()`

```{r NewData, echo=FALSE}
reactable::reactable(NewData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

## *What rows were here before that aren't here now?*

To test for deleted data, we'll compare `CompleteData` and `IncompleteData`, then check the result against `DeletedData`.

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
DeletedData <- dfdiffs::DeletedData
```

### A complete dataset

`CompleteData` is the full dataset, before any rows were removed.

**`CompleteData`**: simulated data for checking 'deleted data'

```{r CompleteData, echo=FALSE}
reactable::reactable(CompleteData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


### An incomplete dataset

`IncompleteData` is a copy of `CompleteData` with some rows removed.

**`IncompleteData`**: simulated data for checking 'deleted data'

```{r IncompleteData, echo=FALSE}
reactable::reactable(IncompleteData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_deleted_data()`

`create_deleted_data()` returns the rows in `base` that are missing from `compare`. Here, that means the rows in `CompleteData` that were dropped from `IncompleteData`:

```{r create_deleted_data, echo=TRUE, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) 
```

Output from **`create_deleted_data()`**: Differences between `CompleteData` and `IncompleteData`

```{r , echo=FALSE}
deleted <- create_deleted_data(
  compare = IncompleteData,
  base = CompleteData)
deleted <- deleted[order(deleted$subject), ]
reactable::reactable(deleted, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### The deleted data

The output above matches the data stored in `DeletedData`, which confirms the deleted rows were identified correctly.

**`DeletedData`**: Output from **`create_deleted_data()`**

```{r DeletedData, echo=FALSE}
reactable::reactable(DeletedData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


## *What values have been changed?*

`dfdiffs` has two functions for this question: `create_changed_data()` and `create_modified_data()`. Both build on the same base-R engine (`diff_values_by_key()` and `compare_values()`), and they differ only in the shape of their output. `create_changed_data()` returns `$num_diffs` and `$var_diffs`, which `compare_data()` uses. `create_modified_data()` returns `$diffs_byvar` and `$diffs`, which the app's compare module (`mod_compare`) and `create_comparison_report()` use.

To check for changed values, we'll compare `InitialData` and `ChangedData`:

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
```

### Initial data

`InitialData` is the original version of the data.

**`InitialData`**: simulated data for checking 'changed/modified data'

```{r InitialData, echo=FALSE}
reactable::reactable(InitialData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### Changed data

`ChangedData` is the version we'll compare against `InitialData`.

**`ChangedData`**: simulated data for checking 'modified data'

```{r ChangedData, echo=FALSE}
reactable::reactable(ChangedData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_changed_data()` 

`create_changed_data()` returns a list of tables. The code below stores the result and prints the names of its elements:

```{r changed}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
names(changed)
```

#### Counts of changes (`num_diffs`)

`num_diffs` stores the number of changes for each variable.

**`num_diffs`**: counts of changes for 'changed data'

```{r num_diffs, echo=FALSE}
reactable::reactable(changed$num_diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

#### Changes by row (`var_diffs`)

`var_diffs` lists the changes row by row.

**`var_diffs`**: Row-by-row of changes for 'changed data'

```{r var_diffs, echo=FALSE}
reactable::reactable(changed$var_diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


### `create_modified_data()` 

`create_modified_data()` also returns a list of tables, with different element names:

```{r modified}
modified <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
names(modified)
```

#### Counts of changes (`diffs_byvar`)

`diffs_byvar` stores the number of changes for each variable.

**`diffs_byvar`**: counts of changes for 'modified data'

```{r diffs_byvar, echo=FALSE}
reactable::reactable(modified$diffs_byvar, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

#### Changes by row (`diffs`)

`diffs` lists the changes row by row.

**`diffs`**: Row-by-row changes of 'modified data'

```{r diffs, echo=FALSE}
reactable::reactable(modified$diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

