---
title: "getting-started"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

This vignette is an intro to the `dfdiffs` package. The goal  of `dfdiffs` is to answer the following questions: 

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   

## Packages 

Our package

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

Dependencies

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(shiny)
library(data.table)
library(dplyr)
library(stringr)
library(forcats)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(vroom)
library(haven)
library(readxl)
library(janitor) 
library(arsenal) 
library(gt)
library(labelled)
library(gtsummary)
```

## Package functions 

We have functions for answering each of the questions posed above. Each function has a pair of datasets to demonstrate how they work (which we'll cover below).

### Call structure

The four comparison functions below are built from a few small helpers, and the same functions power the package's Shiny app. The call trees in this section were generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown: calls to other packages (like `diffdf::diffdf()` or `arsenal::comparedf()`) aren't followed.

```{r call-tree-getting-started, eval=FALSE}
library(stackcallr)
call_tree_dir("R", root = "create_new_data")
call_tree_dir("R", root = "create_deleted_data")
call_tree_dir("R", root = "create_changed_data")
call_tree_dir("R", root = "create_modified_data")
```

`create_new_data()`, `create_deleted_data()` and `create_modified_data()` each call `rename_join_col()` and `create_new_column()` to build (and name) the join column:

```
█─create_new_data
├─rename_join_col
└─create_new_column
```

```
█─create_deleted_data
├─rename_join_col
└─create_new_column
```

```
█─create_modified_data
├─rename_join_col
└─create_new_column
```

`create_changed_data()` uses the same helpers, plus `extract_df_tables()` to turn the output of `diffdf::diffdf()` into tables:

```
█─create_changed_data
├─extract_df_tables
├─rename_join_col
└─create_new_column
```

The Shiny app (`launch_app()`) wires these functions together with three modules (upload, select, and compare), each with a UI function and a server function. `app_ui()` and `app_server()` assemble the modules, and `dfdiffs_fresh_theme()` supplies the app's theme:

```{r call-tree-launch-app, eval=FALSE}
call_tree_dir("R", root = "launch_app")
```

```
█─launch_app
├─█─app_ui
│ ├─dfdiffs_fresh_theme
│ ├─mod_upload_ui
│ ├─mod_select_ui
│ └─mod_compare_ui
└─█─app_server
  ├─█─mod_upload_server
  │ └─█─upload_data
  │   └─load_flat_file
  ├─█─mod_select_server
  │ └─create_join_column
  └─█─mod_compare_server
    ├─%nin%
    ├─█─create_new_data
    │ ├─rename_join_col
    │ └─create_new_column
    ├─█─create_deleted_data
    │ ├─rename_join_col
    │ └─create_new_column
    ├─█─create_modified_data
    │ ├─rename_join_col
    │ └─create_new_column
    └─█─create_changed_data
      ├─extract_df_tables
      ├─rename_join_col
      └─create_new_column
```

## *What rows are here now that weren't here before?*

To check new data, we're going to use `T1Data` and `T2Data`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
NewData <- dfdiffs::NewData
```

### Timepoint 1 data (original)

These data represent data taken at T1. 

```{r T1Data, echo=FALSE}
gt::gt(T1Data) %>% 
  gt::tab_header(data = ., 
    title = md("**`T1Data`**: Simulated 'time-point 1' data"))
```

### Timepoint 2 data (new)

This is a 'new' dataset representing T2.

```{r T2Data, echo=FALSE}
gt::gt(T2Data) %>% 
  gt::tab_header(data = ., 
    title = md("**`T2Data`**: Simulated 'time-point 2' data"))
```

### `create_new_data()`

The `create_new_data()` function shows us the 'new data' (i.e. what is here now that wasn't here before?)

```{r , eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data)
```


```{r create_new_data, echo=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data) %>% 
  gt::gt(data = .) %>% 
  gt::tab_header(data = ., 
  title = md("Output from **`create_new_data()`**"), 
    subtitle = md("Difference between 'time-point 1' & 'time-point 2'"))
```

We can check this against the `NewData` dataset (which should match the output from `create_new_data()`)

```{r NewData, echo=FALSE}
gt::gt(NewData) %>% 
    gt::tab_header(data = ., 
      title = md("**`NewData`**"), 
      subtitle = md("stored differences from `create_new_data()`"))
```

## *What rows were here before that aren't here now?*

To test for the deleted data, we use the `CompleteData`, `IncompleteData`, and check these with `DeletedData`. 

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
DeletedData <- dfdiffs::DeletedData
```

### A complete dataset

`CompleteData` represents a 'complete' set of data. 

```{r CompleteData, echo=FALSE}
gt::gt(CompleteData) %>% 
  gt::tab_header(data = ., 
  title = md("**`CompleteData`**"), 
    subtitle = md("simulated data for checking 'deleted data'"))
```


### An incomplete dataset

This is a dataset with rows removed from `CompleteData`.

```{r IncompleteData, echo=FALSE}
gt::gt(IncompleteData) %>% 
  gt::tab_header(data = ., 
  title = md("**`IncompleteData`**"), 
    subtitle = md("simulated data for checking 'deleted data'"))
```

### `create_deleted_data()`

When we run the `create_deleted_data()`, we check for the deleted rows between `IncompleteData` and `CompleteData`.

```{r create_deleted_data, echo=TRUE, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) 
```

```{r , echo=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) %>% 
  dplyr::arrange(subject) %>% 
  gt::gt(data = .) %>% 
  gt::tab_header(data = ., 
  title = md("Output from **`create_deleted_data()`**"), 
    subtitle = md("Differences between `CompleteData` and `IncompleteData`"))
```

### The deleted data 

This is identical to the data stored in `DeletedData`

```{r DeletedData, echo=FALSE}
gt::gt(DeletedData) %>% 
  gt::tab_header(data = ., 
  title = md("**`DeletedData`**"), 
    subtitle = md("Output from **`create_deleted_data()`**"))
```


## *What values have been changed?*

To answer this question, we have two options: `create_changed_data()` and `create_modified_data()`.

+ `create_changed_data()` relies on the `diffdf()` function from the [`diffdf` package](https://gowerc.github.io/diffdf/reference/diffdf.html) package. 

+ `create_modified_data()` relies on the `comparedf()` function from the [`arsenal` package](https://mayoverse.github.io/arsenal/reference/comparedf.html) package.

To check for changes between two datasets, we use the `InitialData` and `ChangedData`. 

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
```

### Initial data 

```{r InitialData, echo=FALSE}
gt::gt(InitialData) %>% 
  gt::tab_header(data = ., 
  title = md("**`InitialData`**"), 
    subtitle = md("simulated data for checking 'changed/modified data'"))
```

### Changed data

```{r ChangedData, echo=FALSE}
gt::gt(ChangedData) %>% 
  gt::tab_header(data = ., 
  title = md("**`ChangedData`**"), 
    subtitle = md("simulated data for checking 'modified data'"))
```

### `create_changed_data()` 

`create_changed_data()` creates a list of tables. 

```{r changed}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
names(changed)
```

#### Counts of changes (`num_diffs`)

The counts of changes by variable are stored in `num_diffs`.

```{r num_diffs, echo=FALSE}
gt::gt(data = changed$num_diffs) %>% 
    gt::tab_header(data = ., 
  title = md("**`num_diffs`**"), 
    subtitle = md("counts of changes for 'changed data'"))
```

#### Changes by row (`var_diffs`)

The changes by row are stored in `var_diffs`.

```{r var_diffs, echo=FALSE}
gt::gt(data = changed$var_diffs) %>% 
    gt::tab_header(data = ., 
  title = md("**`var_diffs`**"), 
    subtitle = md("Row-by-row of changes for 'changed data'"))
```


### `create_modified_data()` 

The `create_modified_data()` function also creates a list of tables. 

```{r modified}
modified <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
names(modified)
```

#### Counts of changes (`diffs_byvar`)

The counts of changes by variable are stored in `diffs_byvar`.

```{r diffs_byvar, echo=FALSE}
gt::gt(data = modified$diffs_byvar) %>% 
    gt::tab_header(data = ., 
  title = md("**`diffs_byvar`**"), 
    subtitle = md("counts of changes for 'modified data'"))
```

#### Changes by row 

The changes by row are stored in `diffs`.

```{r diffs, echo=FALSE}
gt::gt(data = modified$diffs) %>% 
    gt::tab_header(data = ., 
  title = md("**`diffs`**"), 
    subtitle = md("Row-by-row changes of 'modified data'"))
```

