---
title: "getting-started"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation

This vignette introduces the `dfdiffs` package. The goal of `dfdiffs` is to answer the following questions:

1. What rows are here now that weren't here before?  
2. What rows were here before that aren't here now?   
3. What values have been changed?   

## Packages 

Load `dfdiffs`:

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

Load the other packages used in this vignette:

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(shiny)
library(data.table)
library(stringr)
library(lubridate)
library(fs)
library(vctrs)
library(glue)
library(purrr)
library(haven)
library(readxl)
library(janitor)
library(reactable)
library(labelled)
library(gtsummary)
```

## Package functions 

`dfdiffs` has a function for each of the questions posed above, and each function comes with a pair of example datasets that demonstrate how it works (which we'll cover below).

### Call structure

The four comparison functions below are built from a few small helpers, and the same functions power the package's Shiny app. The call trees in this section were generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-getting-started, eval=FALSE}
library(stackcallr)
call_tree_dir("R", root = "create_new_data")
call_tree_dir("R", root = "create_deleted_data")
call_tree_dir("R", root = "create_changed_data")
call_tree_dir("R", root = "create_modified_data")
```

`create_new_data()` and `create_deleted_data()` (the "new rows"/"deleted rows" answers) share `anti_join_base()`, `select_cols()`, `rename_join_col()`, and `create_new_column()`. These are all base-R helpers, so no external comparison package is involved:

```
█─create_new_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

```
█─create_deleted_data
├─anti_join_base
├─select_cols
├─rename_join_col
└─create_new_column
```

`create_changed_data()` and `create_modified_data()` (the "changed values" answer) both build on `diff_values_by_key()`, which compares matched rows column-by-column with `compare_values()` (tolerance-aware for numeric/date-time values, label-aware for factors) and records any class mismatches via `class_diffs()`:

```
█─create_changed_data
├─select_cols
├─█─diff_values_by_key
│ ├─compare_values
│ └─class_diffs
├─rename_join_col
└─create_new_column
```

The Shiny app (`launch_app()`) wires these functions together with three modules (upload, select, and compare), each with a UI function and a server function. `app_ui()` and `app_server()` assemble the modules, and `dfdiffs_fresh_theme()` supplies the app's theme:

```{r call-tree-launch-app, eval=FALSE}
call_tree_dir("R", root = "launch_app")
```

```
█─launch_app
├─█─app_ui
│ ├─dfdiffs_fresh_theme
│ ├─mod_upload_ui
│ ├─mod_select_ui
│ └─mod_compare_ui
└─█─app_server
  ├─█─mod_upload_server
  │ ├─█─upload_data
  │ │ └─load_flat_file
  │ ├─base_react_theme
  │ └─comp_react_theme
  ├─█─mod_select_server
  │ ├─select_cols
  │ ├─base_react_theme
  │ ├─comp_react_theme
  │ ├─info_react_theme
  │ └─█─create_join_column
  │   └─select_cols
  └─█─mod_compare_server
    ├─%nin%
    ├─info_react_theme
    ├─█─create_new_data
    │ ├─anti_join_base
    │ ├─select_cols
    │ ├─rename_join_col
    │ └─create_new_column
    ├─new_react_theme
    ├─█─create_deleted_data
    │ ├─anti_join_base
    │ ├─select_cols
    │ ├─rename_join_col
    │ └─create_new_column
    ├─deleted_react_theme
    ├─█─create_modified_data
    │ ├─select_cols
    │ ├─█─diff_values_by_key
    │ │ ├─compare_values
    │ │ └─class_diffs
    │ ├─rename_join_col
    │ └─create_new_column
    ├─select_cols
    ├─changed_react_theme
    └─left_join_base
```

## *What rows are here now that weren't here before?*

To check for new data, we'll use `T1Data` and `T2Data`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
NewData <- dfdiffs::NewData
```

### Timepoint 1 data (original)

`T1Data` represents data collected at the first timepoint (T1).

**`T1Data`**: Simulated 'time-point 1' data

```{r T1Data, echo=FALSE}
reactable::reactable(T1Data, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### Timepoint 2 data (new)

`T2Data` is the 'new' dataset, representing data collected at the second timepoint (T2).

**`T2Data`**: Simulated 'time-point 2' data

```{r T2Data, echo=FALSE}
reactable::reactable(T2Data, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_new_data()`

The `create_new_data()` function returns the 'new data' (i.e., the rows that are here now but weren't here before).

```{r , eval=FALSE}
create_new_data(
  compare = T2Data, 
  base = T1Data)
```


Output from **`create_new_data()`**: Difference between 'time-point 1' & 'time-point 2'

```{r create_new_data, echo=FALSE}
create_new_data(
  compare = T2Data,
  base = T1Data) %>%
  reactable::reactable(resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

We can check this against the `NewData` dataset, which should match the output from `create_new_data()`.

**`NewData`**: stored differences from `create_new_data()`

```{r NewData, echo=FALSE}
reactable::reactable(NewData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

## *What rows were here before that aren't here now?*

To test for deleted data, we'll compare `CompleteData` and `IncompleteData`, then check the result against `DeletedData`.

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
DeletedData <- dfdiffs::DeletedData
```

### A complete dataset

`CompleteData` represents a 'complete' set of data. 

**`CompleteData`**: simulated data for checking 'deleted data'

```{r CompleteData, echo=FALSE}
reactable::reactable(CompleteData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


### An incomplete dataset

This is a dataset with rows removed from `CompleteData`.

**`IncompleteData`**: simulated data for checking 'deleted data'

```{r IncompleteData, echo=FALSE}
reactable::reactable(IncompleteData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_deleted_data()`

Running `create_deleted_data()` checks for rows that were deleted between `CompleteData` and `IncompleteData`.

```{r create_deleted_data, echo=TRUE, eval=FALSE}
create_deleted_data(
  compare = IncompleteData, 
  base = CompleteData) 
```

Output from **`create_deleted_data()`**: Differences between `CompleteData` and `IncompleteData`

```{r , echo=FALSE}
deleted <- create_deleted_data(
  compare = IncompleteData,
  base = CompleteData)
deleted <- deleted[order(deleted$subject), ]
reactable::reactable(deleted, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### The deleted data

The output above is identical to the data stored in `DeletedData`.

**`DeletedData`**: Output from **`create_deleted_data()`**

```{r DeletedData, echo=FALSE}
reactable::reactable(DeletedData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


## *What values have been changed?*

To answer this question, we have two options: `create_changed_data()` and `create_modified_data()`. Both build on the same base-R `diff_values_by_key()`/`compare_values()` engine and differ only in the shape of their output: `create_changed_data()` returns `$num_diffs`/`$var_diffs` (used by `compare_data()`), and `create_modified_data()` returns `$diffs_byvar`/`$diffs` (used by `mod_compare` and `create_comparison_report()`).

To check for changes between two datasets, we'll use `InitialData` and `ChangedData`.

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
```

### Initial data 

**`InitialData`**: simulated data for checking 'changed/modified data'

```{r InitialData, echo=FALSE}
reactable::reactable(InitialData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### Changed data

**`ChangedData`**: simulated data for checking 'modified data'

```{r ChangedData, echo=FALSE}
reactable::reactable(ChangedData, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

### `create_changed_data()` 

`create_changed_data()` creates a list of tables. 

```{r changed}
changed <- create_changed_data(
  compare = ChangedData,
  base = InitialData)
names(changed)
```

#### Counts of changes (`num_diffs`)

The counts of changes by variable are stored in `num_diffs`.

**`num_diffs`**: counts of changes for 'changed data'

```{r num_diffs, echo=FALSE}
reactable::reactable(changed$num_diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

#### Changes by row (`var_diffs`)

The changes by row are stored in `var_diffs`.

**`var_diffs`**: Row-by-row of changes for 'changed data'

```{r var_diffs, echo=FALSE}
reactable::reactable(changed$var_diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```


### `create_modified_data()` 

The `create_modified_data()` function also creates a list of tables. 

```{r modified}
modified <- create_modified_data(
  compare = ChangedData,
  base = InitialData)
names(modified)
```

#### Counts of changes (`diffs_byvar`)

The counts of changes by variable are stored in `diffs_byvar`.

**`diffs_byvar`**: counts of changes for 'modified data'

```{r diffs_byvar, echo=FALSE}
reactable::reactable(modified$diffs_byvar, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

#### Changes by row 

The changes by row are stored in `diffs`.

**`diffs`**: Row-by-row changes of 'modified data'

```{r diffs, echo=FALSE}
reactable::reactable(modified$diffs, resizable = TRUE, highlight = TRUE, compact = TRUE, wrap = FALSE, bordered = TRUE, filterable = TRUE)
```

