---
title: "multiple-comparisons"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{multiple-comparisons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(labelled)
library(gt)
library(gtsummary)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation 

This vignette covers how to apply the functions in `dfdiffs` to multiple datasets (separated by folder)

## Packages 

Our package

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```


Similar packages 

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## Folder structure 

Below is a folder structure of datasets separated by year. 

```{r eval=FALSE, echo=FALSE}
People20 <- readr::read_csv("../inst/extdata/csv/by-year/People2020.csv")
PlayerBirth20 <- People20 |> select(playerID, starts_with("birth"))
write_csv(x = PlayerBirth20, "../inst/extdata/csv/by-year/20/PlayerBirth.csv")
PlayerDeath20 <- People20 |> select(playerID, starts_with("death"))
write_csv(x = PlayerDeath20, "../inst/extdata/csv/by-year/20/PlayerDeath.csv")
PlayerName20 <- People20 |> select(playerID, starts_with("name"))
write_csv(x = PlayerName20, "../inst/extdata/csv/by-year/20/PlayerName.csv")

People21 <- readr::read_csv("../inst/extdata/csv/by-year/People2021.csv")
PlayerBirth21 <- People21 |> select(playerID, starts_with("birth"))
write_csv(x = PlayerBirth21, "../inst/extdata/csv/by-year/21/PlayerBirth.csv")
PlayerDeath21 <- People21 |> select(playerID, starts_with("death"))
write_csv(x = PlayerDeath21, "../inst/extdata/csv/by-year/21/PlayerDeath.csv")
PlayerName21 <- People21 |> select(playerID, starts_with("name"))
write_csv(x = PlayerName21, "../inst/extdata/csv/by-year/21/PlayerName.csv")
```

```{r lahman-tree}
fs::dir_tree("../inst/extdata/csv/by-year")
```

We start by storing these files in a vector with `list.files()`, then we clean up the names a bit, and pass the list of paths to `readr::read_csv` with `purrr::map()`:

```{r base_files}
base_files <- list.files(path = "../inst/extdata/csv/by-year/20", 
  pattern = ".csv", recursive = TRUE, full.names = TRUE)
base_files_nms <- map_chr(base_files, base::basename)
base_dfs <- base_files |> 
  purrr::set_names(base_files_nms) |> 
  purrr::map(readr::read_csv)
```

```{r compare_files}
compare_files <- list.files(path = "../inst/extdata/csv/by-year/21", 
  pattern = ".csv", recursive = TRUE, full.names = TRUE)
compare_files_nms <- map_chr(compare_files, base::basename)
compare_dfs <- compare_files |> 
  purrr::set_names(compare_files_nms) |> 
  purrr::map(readr::read_csv)
```

# Iteration

We're going to follow the iteration steps from [Charlotte Wickham's Happy R Users Purrr Tutorial](https://www.rstudio.com/resources/rstudioconf-2017/happy-r-users-purrr-tutorial-/)

1. DO IT FOR ONE   
2. TURN IT INTO A RECIPE  
3. DO IT FOR ALL!  

# Do it for one (or two)

We're going to use the `arsenal::comparedf()` function first. Remember that this function requires an additional `summary()` call! 

## `arsenal::comparedf`

```{r cdf_people}
by_var <- "playerID"
cdf_people <- arsenal::comparedf(
  x = base_dfs$PlayerName.csv, 
  y = compare_dfs$PlayerName.csv, 
  by = by_var)
sumcdf_people <- summary(cdf_people)
names(sumcdf_people)
```


## `diffdf::diffdf`

Next we'll use the `diffdf::diffdf()` function.

```{r diffdf_people}
diffdf_people <- diffdf::diffdf(
  compare = compare_dfs$PlayerName.csv, 
  base = base_dfs$PlayerName.csv, 
  keys = "playerID")
diffdf_people |> names()
```


Just to ensure we're getting the same display in the output from both functions, we're going to use the `diffs.byvar.table`/`NumDiff` tables, and the `diffs.table`/ `VarDiff_` tables. 

### `diffs.byvar.table`

```{r diffs.byvar.table}
sumcdf_people[["diffs.byvar.table"]] |> 
  select(Variable = var.x, 
         `No of Differences` = n) |> 
  knitr::kable(caption = "diffs.byvar.table") |> 
  kableExtra::kable_paper()
```

### `NumDiff`

```{r NumDiff}
diffdf_people[["NumDiff"]] |> 
  knitr::kable(caption = "NumDiff") |> 
  kableExtra::kable_paper()
```


### `diffs.table`

```{r diffs.table}
sumcdf_people[["diffs.table"]] |> 
  unnest(values.x) |> 
  unnest(values.y) |> 
  select(
    VARIABLE = var.x,
    all_of(by_var),
    BASE = values.x,
    COMPARE = values.y
  ) |> 
  knitr::kable(caption = "diffs.table") |> 
  kableExtra::kable_paper()
```


### `VarDiff_` tables

```{r VarDiff_tables}
tbl_names <- names(diffdf_people)
vardiff_tbls <- tbl_names[stringr::str_detect(tbl_names, "VarDiff_")]
bind_rows(diffdf_people[vardiff_tbls]) |> 
  knitr::kable(caption = "'VarDiff_' tables") |> 
  kableExtra::kable_paper()
```

## Turn it into a recipe

We have two vectors with file paths (`.x` and `.y`)

```{r x-y, eval=FALSE}
map2(.x = base_dfs, .y = compare_files, )
```

And the `.f` is our `arsenal::comparedf` or `diffdf::diffdf` functions. 

```{r function-map2, eval=FALSE}
map2(.x = base_dfs, .y = compare_files, .f = arsenal::comparedf)
map2(.x = compare_files, .y = base_dfs, .f = diffdf::diffdf)
```

We can pass the `by` and `keys` arguments to `...`.

```{r eval=FALSE}
purrr::map2(.x =  base_dfs, .y = compare_dfs, 
            .f = arsenal::comparedf, by = "playerID")
```

```{r eval=FALSE}
purrr::map2(.x =  compare_dfs,  .y = base_dfs, 
            .f = diffdf::diffdf, keys = "playerID")
```


## Do it for all

### `map2()` + `arsenal::comparedf`

Now we use the `purrr::map2()` function to compare all of the items in `base_dfs` to `compare_dfs`.

```{r all_cdfs}
all_cdfs <- purrr::map2(
  .x =  base_dfs, 
  .y = compare_dfs, 
  .f = arsenal::comparedf, 
  by = "playerID")
all_scdfs <- purrr::map(all_cdfs, .f = summary)
all_scdfs[["PlayerName.csv"]] |> names()
```

### `map2()` + `diffdf::diffdf`

```{r all_diffdfs}
all_diffdfs <- purrr::map2(
  .x =  base_dfs, 
  .y = compare_dfs, 
  .f = diffdf::diffdf, 
  keys = "playerID")
all_diffdfs[["PlayerName.csv"]] |> names()
```

## Extracting tables 

Now that we have two lists with multiple comparisons, we need to be able to extract the two table 
