---
title: "multiple-comparisons"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{multiple-comparisons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation 

The goal of `dfdiffs` is to answer the following questions:

1.  What rows are here now that weren't here before?
2.  What rows were here before that aren't here now?
3.  **What values have been changed?**

This vignette covers how to apply the functions in `dfdiffs` to multiple datasets (separated by folder).

## Packages

Load `dfdiffs`:

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

We'll also need packages for import/export, iteration, and wrangling:

```{r more-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(glue)
```

These packages are for building tables:

```{r tbl-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(labelled)
library(gtsummary)
library(kableExtra)
```

These packages offer similar comparison functions:

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## Folder structure

Below is a folder structure of datasets separated by year. It holds two yearly pulls of the synthetic (not real) `dfdiffs` site roster (see `?Roster2021`), which `data-raw/Roster.R` has already split into `Enroll`/`Visit`/`Name` files.

```{r lahman-tree}
fs::dir_tree("../inst/extdata/csv/site-roster")
```

We start by storing these file paths in a vector with `list.files()`. Then we clean up the names a bit and pass the paths to `readr::read_csv` with `purrr::map()`.

### Import base dfs

Below we import the base data tables (from the 2021 pull).

```{r base_files, message=FALSE, warning=FALSE}
base_files <- list.files(path = "../inst/extdata/csv/site-roster/2021",
  pattern = "Enroll|Visit|Name", recursive = TRUE, full.names = TRUE)
base_files_nms <- purrr::map_chr(base_files, base::basename)
base_dfs <- base_files |>
  purrr::set_names(base_files_nms) |>
  purrr::map(readr::read_csv)
base_dfs |> str()
```

### Import compare dfs

Below we import the compare data tables (from the 2022 pull).

```{r compare_files, message=FALSE, warning=FALSE}
compare_files <- list.files(path = "../inst/extdata/csv/site-roster/2022",
  pattern = "Enroll|Visit|Name", recursive = TRUE, full.names = TRUE)
compare_files_nms <- purrr::map_chr(compare_files, base::basename)
compare_dfs <- compare_files |>
  purrr::set_names(compare_files_nms) |>
  purrr::map(readr::read_csv)
compare_dfs |> str()
```

# Iteration

We're going to follow the iteration steps from [Charlotte Wickham's Happy R Users Purrr Tutorial](https://www.rstudio.com/resources/rstudioconf-2017/happy-r-users-purrr-tutorial-/):

1. DO IT FOR ONE   
2. TURN IT INTO A RECIPE  
3. DO IT FOR ALL!  

## 1) Do it for one (or two)

We'll start with the `arsenal::comparedf()` function. Remember that this function requires an additional `summary()` call.

### `arsenal::comparedf`

```{r cdf_people}
by_var <- "subject_id"
cdf_people <- arsenal::comparedf(
  x = base_dfs$Enroll.csv, 
  y = compare_dfs$Enroll.csv, 
  by = by_var)
sumcdf_people <- summary(cdf_people)
names(sumcdf_people)
```


### `diffdf::diffdf`

Next we'll use the `diffdf::diffdf()` function.

```{r diffdf_people}
diffdf_people <- diffdf::diffdf(
  compare = compare_dfs$Enroll.csv, 
  base = base_dfs$Enroll.csv, 
  keys = "subject_id")
diffdf_people |> names()
```


To make sure the output from both functions is displayed the same way, we'll use the `diffs.byvar.table`/`NumDiff` tables and the `diffs.table`/`VarDiff_` tables.

#### `diffs.byvar.table`

```{r diffs.byvar.table}
sumcdf_people[["diffs.byvar.table"]] |> 
  select(Variable = var.x, 
         `No of Differences` = n) |> 
  knitr::kable(caption = "diffs.byvar.table") |> 
  kableExtra::kable_paper()
```

#### `NumDiff`

```{r NumDiff}
diffdf_people[["NumDiff"]] |> 
  knitr::kable(caption = "NumDiff") |> 
  kableExtra::kable_paper()
```


#### `diffs.table`

```{r diffs.table}
sumcdf_people[["diffs.table"]] |> 
  unnest(values.x) |> 
  unnest(values.y) |> 
  select(
    VARIABLE = var.x,
    all_of(by_var),
    BASE = values.x,
    COMPARE = values.y
  ) |> 
  knitr::kable(caption = "diffs.table") |> 
  kableExtra::kable_paper()
```


#### `VarDiff_` tables

```{r VarDiff_tables}
tbl_names <- names(diffdf_people)
vardiff_tbls <- tbl_names[stringr::str_detect(tbl_names, "VarDiff_")]
bind_rows(diffdf_people[vardiff_tbls]) |> 
  knitr::kable(caption = "'VarDiff_' tables") |> 
  kableExtra::kable_paper()
```

## 2) Turn it into a recipe

We have two vectors with file paths (`.x` and `.y`).

```{r x-y, eval=FALSE}
map2(.x = base_dfs, .y = compare_files, )
```

The `.f` argument is either the `arsenal::comparedf` or the `diffdf::diffdf` function.

```{r function-map2, eval=FALSE}
map2(.x = base_dfs, .y = compare_files, .f = arsenal::comparedf)
map2(.x = compare_files, .y = base_dfs, .f = diffdf::diffdf)
```

We can pass the `by` and `keys` arguments through `...`.

```{r eval=FALSE}
purrr::map2(.x =  base_dfs, .y = compare_dfs, 
            .f = arsenal::comparedf, by = "subject_id")
```

```{r eval=FALSE}
purrr::map2(.x =  compare_dfs,  .y = base_dfs, 
            .f = diffdf::diffdf, keys = "subject_id")
```


## 3) Do it for all

### `map2()` + `arsenal::comparedf`

Now we use the `purrr::map2()` function to compare all of the items in `base_dfs` to `compare_dfs`.

```{r all_cdfs}
all_cdfs <- purrr::map2(
  .x =  base_dfs, 
  .y = compare_dfs, 
  .f = arsenal::comparedf, 
  by = "subject_id")
all_scdfs <- purrr::map(all_cdfs, .f = summary)
all_scdfs[["Name.csv"]] |> names()
```

### `map2()` + `diffdf::diffdf`

We can do the same with `diffdf::diffdf()`.

```{r all_diffdfs}
all_diffdfs <- purrr::map2(
  .x = base_dfs, 
  .y = compare_dfs, 
  .f = diffdf::diffdf, 
  keys = "subject_id")
all_diffdfs[["Name.csv"]] |> names()
```


