---
title: "multiple-comparisons"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{multiple-comparisons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation 

The goal of `dfdiffs` is to answer the following questions:

1.  What rows are here now that weren't here before?
2.  What rows were here before that aren't here now?
3.  **What values have been changed?**

In this vignette I'll cover how to apply the functions in `dfdiffs` to multiple pairs of datasets stored in separate folders.

## Packages

First, we'll load `dfdiffs`.

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

We'll also need packages for importing data, iteration, and building tables.

```{r more-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(readr)
library(purrr)
```

## Folder structure

The folder structure below stores datasets separated by year. It holds two yearly pulls of the synthetic (not real) `dfdiffs` site roster (see `?Roster2021`), which `data-raw/Roster.R` has already split into `Enroll`, `Visit`, and `Name` files.

```{r site-roster-tree}
fs::dir_tree("../inst/extdata/csv/site-roster")
```

We'll store the file paths in a vector with `list.files()` and name each path with its file name. Then we'll pass the paths to `readr::read_csv()` with `purrr::map()`.

### Import base dfs

The code below imports the base data tables from the 2021 pull.

```{r base_files, message=FALSE, warning=FALSE}
base_files <- list.files(path = "../inst/extdata/csv/site-roster/2021",
  pattern = "Enroll|Visit|Name", recursive = TRUE, full.names = TRUE)
base_files_nms <- purrr::map_chr(base_files, base::basename)
base_dfs <- base_files |>
  purrr::set_names(base_files_nms) |>
  purrr::map(readr::read_csv)
base_dfs |> str()
```

### Import compare dfs

The code below imports the compare data tables from the 2022 pull.

```{r compare_files, message=FALSE, warning=FALSE}
compare_files <- list.files(path = "../inst/extdata/csv/site-roster/2022",
  pattern = "Enroll|Visit|Name", recursive = TRUE, full.names = TRUE)
compare_files_nms <- purrr::map_chr(compare_files, base::basename)
compare_dfs <- compare_files |>
  purrr::set_names(compare_files_nms) |>
  purrr::map(readr::read_csv)
compare_dfs |> str()
```

# Iteration

We'll follow the three iteration steps from [Charlotte Wickham's Happy R Users Purrr Tutorial](https://www.rstudio.com/resources/rstudioconf-2017/happy-r-users-purrr-tutorial-/):

1. DO IT FOR ONE
2. TURN IT INTO A RECIPE
3. DO IT FOR ALL!

## 1) Do it for one (or two)

`compare_data()` runs all three comparisons (new, deleted, and changed) in a single call. We'll start with a single pair of tables, `Enroll.csv`.

```{r one_comparison}
by_var <- "subject_id"
enroll_comparison <- compare_data(
  compare = compare_dfs$Enroll.csv,
  base = base_dfs$Enroll.csv,
  by = by_var)
names(enroll_comparison)
```

`compare_summary()` adds a headline summary for the same pair of tables, in the style of SAS's `PROC COMPARE`.

```{r one_summary}
compare_summary(
  compare = compare_dfs$Enroll.csv,
  base = base_dfs$Enroll.csv,
  by = by_var)
```

## 2) Turn it into a recipe

`compare_data()` takes `compare` and `base` as its first two arguments, so `purrr::map2()` can pass the matching elements of `compare_dfs` and `base_dfs` straight through. The `by` argument is passed along via `...`. Note that `map2()` pairs the two lists by position, not by name, so this relies on `list.files()` returning the files in the same order in both folders (which it does here, because both folders hold the same three file names).

```{r function-map2, eval=FALSE}
purrr::map2(.x = compare_dfs, .y = base_dfs,
            .f = compare_data, by = "subject_id")
```

## 3) Do it for all

Now we'll apply the recipe to every pair of files in `compare_dfs` and `base_dfs`.

```{r all_comparisons}
all_comparisons <- purrr::map2(
  .x = compare_dfs,
  .y = base_dfs,
  .f = compare_data,
  by = "subject_id")
names(all_comparisons)
names(all_comparisons$Name.csv)
```

The same recipe works with `compare_summary()` and gives us one headline summary per file. Below we print the summary for `Visit.csv`.

```{r all_summaries}
all_summaries <- purrr::map2(
  .x = compare_dfs,
  .y = base_dfs,
  .f = compare_summary,
  by = "subject_id")
all_summaries$Visit.csv
```


