---
title: "proc-compare-parity"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{proc-compare-parity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

# Motivation

SAS's `PROC COMPARE` is the reference point for `dfdiffs`. Both tools answer
the same questions about two versions of a dataset: which rows are new,
which rows are gone, and which values changed? This vignette maps
`PROC COMPARE`'s options and output to the matching `dfdiffs` functions. It
also shows that each function produces readable output at the console on its
own, independent of the Shiny app. The app (`launch_app()`) is a thin layer
on top of these functions, and it doesn't compute anything they don't.

The only package we need is `dfdiffs`:

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
```

# Mapping `PROC COMPARE` to `dfdiffs`

The table below pairs each `PROC COMPARE` concept with the `dfdiffs`
function or argument that handles it.

| `PROC COMPARE` concept | `dfdiffs` function |
|:---|:---|
| `BASE=`/`COMPARE=` datasets | `base`/`compare` arguments, every comparison function |
| `ID`/`BY` statement (row matching key) | `by` argument |
| Variable Summary (vars only in one dataset) | `compare_columns()` |
| `METHOD=ABSOLUTE`/`CRITERION=` (numeric tolerance) | `tolerance`/`scale` arguments to `compare_values()` |
| case-sensitive text compare (`PROC COMPARE`'s default) | `ignore_case`/`trim_ws` (a `dfdiffs` addition, off by default) |
| duplicate `ID` note | a `warning()` from `diff_values_by_key()` |
| Observation Summary / Values Comparison Summary | `compare_summary()` |
| the full printed comparison listing | `create_comparison_report()`'s xlsx output |

# Roster data

We'll compare two snapshots of the synthetic clinical site roster, taken a
year apart:

```{r data}
Roster2021 <- dfdiffs::Roster2021
Roster2022 <- dfdiffs::Roster2022
nrow(Roster2021)
nrow(Roster2022)
```

# Variable Summary: `compare_columns()`

Before comparing any values, `PROC COMPARE` lists the variables that exist
in only one dataset. `compare_columns()` answers the same question and
prints readable output at the console:

```{r compare_columns}
compare_columns(base = Roster2021, compare = Roster2022)
```

# Observation Summary / Values Comparison Summary: `compare_summary()`

`PROC COMPARE` prints a headline summary of row, column, and value
differences before its detailed diff tables. `compare_summary()` gives the
same headline by building on `compare_data()` and `compare_columns()`, and
it adds no comparison logic of its own:

```{r compare_summary}
compare_summary(compare = Roster2022, base = Roster2021, by = "subject_id")
```

A small example makes the value-level counts easier to follow. In the two
data frames below, `note` differs in two of the three rows:

```{r compare_summary-diffs}
base <- data.frame(
  id = 1:3, note = c("ok", "ok", "ok"), val = c(1, 2, 3), stringsAsFactors = FALSE
)
compare <- data.frame(
  id = 1:3, note = c("OK", "ok", "changed"), val = c(1, 2, 3), stringsAsFactors = FALSE
)
compare_summary(compare = compare, base = base, by = "id")
```

`PROC COMPARE` compares character values case-sensitively with no option to
change that. Setting `ignore_case = TRUE` treats `"ok"` and `"OK"` as equal,
an option `dfdiffs` offers beyond `PROC COMPARE` parity:

```{r compare_summary-ignore_case}
compare_summary(compare = compare, base = base, by = "id", ignore_case = TRUE)
```

# Duplicate keys

When the `BY`/`ID` variables don't uniquely identify rows, `PROC COMPARE`
prints a note instead of stopping with an error. `diff_values_by_key()` does
the same by issuing a warning:

```{r duplicate-keys}
base_dupe <- data.frame(id = c(1, 1, 2), val = c("a", "z", "b"), stringsAsFactors = FALSE)
compare_dupe <- data.frame(id = c(1, 2), val = c("A", "b"), stringsAsFactors = FALSE)
diff_values_by_key(base = base_dupe, compare = compare_dupe, by = "id")
```

# Numeric tolerance

`compare_values()` covers `PROC COMPARE`'s `CRITERION=` with its `tolerance`
argument. The first call below relies on the default tolerance
(`sqrt(.Machine$double.eps)`), and the second sets a wider one:

```{r tolerance}
compare_values(5.0, 5.00000001)
compare_values(5.0, 5.1, tolerance = 0.2)
```

# The full report: `create_comparison_report()`

`PROC COMPARE` ends with a full printed listing of the comparison.
`create_comparison_report()` plays that role by writing every table shown
above, plus the row-level and value-level diff tables, to a single xlsx file.
The workbook has nine sheets: New Data, Deleted Data, Changed Data, Review
Changes, Base Data, Compare Data, Column Diffs, Class Diffs, and Summary.

```{r create_comparison_report, eval=FALSE}
create_comparison_report(
  compare = Roster2022,
  base = Roster2021,
  by = "subject_id",
  file = "roster-comparison.xlsx"
)
```

The download button in the Shiny app calls this same function, so the app
supplies no additional comparison logic of its own.
