---
title: "similar-work"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{similar-work}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(tidyverse)
library(labelled)
library(gt)
library(gtsummary)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  error = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation 

This vignette covers similar functions/packages that compare two datasets to each other. 

## Packages 

Our package

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```


Similar packages 

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(janitor) # compare_df_cols
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(testthat) # expect_equal
library(vetr) # alike
```

## Data 

We'll be using the data in this package to cover similar `packages::functions()`. 

### New data

To check new data, we're going to use `T1Data` and `T2Data`. The `create_new_data()` function shows us the 'new data' (i.e. what is here now that wasn't here before?)

We can check this against the `NewData` dataset (which should match the output from `create_new_data()`)

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
create_new_data(newdf = T2Data, olddf = T1Data)
```

```{r NewData, echo=TRUE}
NewData <- dfdiffs::NewData
NewData
```


###  Deleted data 

To test for the deleted data, we use the `CompleteData`, `IncompleteData`, and `DeletedData`. `CompleteData` represents a 'complete' set of data, and `IncompleteData` is a dataset with rows removed from `CompleteData`. 

When we run the `create_deleted_data()`, we check for the deleted rows between `IncompleteData` and `CompleteData`.

```{r CompleteData-IncompleteData, echo = TRUE}
CompleteData <- dfdiffs::CompleteData
IncompleteData <- dfdiffs::IncompleteData
create_deleted_data(newdf = IncompleteData, olddf = CompleteData)
```

This is identical to the data stored in `DeletedData`

```{r DeletedData}
DeletedData <- dfdiffs::DeletedData
DeletedData
```


### Modified Data 

To check for changes between two datasets, we uses the `InitialData` and `ChangedData`. 

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
mods <- create_modified_data(
  newdf = ChangedData, 
  olddf = InitialData)
```

The changes by variable are stored in `diffs_byvar`.

```{r diffs_byvar}
mods$diffs_byvar
```

The changes by row are stored in `diffs`.

```{r diffs}
mods$diffs
```

## Lahman datasets 

We will also import two master datasets from the `Lahman` baseball package. 

```{r , echo=TRUE}
master15 <- readxl::read_excel(
  path = "../inst/extdata/xlsx/lahman_compare.xlsx", 
  sheet = "master-2015")
master20 <- readxl::read_excel(
  path = "../inst/extdata/xlsx/lahman_compare.xlsx", 
  sheet = "master-2020")
```

### Master (2015)

```{r master15, echo=TRUE}
glimpse(master15)
```

### Master (2020)

```{r master20, echo=TRUE}
glimpse(master20)
```


## `janitor::compare_df_cols_same()`

The `compare_df_cols_same()` function from `janitor` compares two datasets and "*indicates if they will successfully bind together by rows.*"

```{r compare_df_cols_same, echo=TRUE}
compare_df_cols_same(T1Data, T2Data, strict_description = FALSE)
compare_df_cols_same(CompleteData, IncompleteData, strict_description = FALSE)
compare_df_cols_same(InitialData, ChangedData, strict_description = FALSE)
```

All of our test datasets meet this condition, but this could be used as a step in one our `create_` functions (to see if they can be successfully bound together).

## `diffdf::diffdf()`

The `diffdf()` function is very similar to `arsenal::comparedf()`

If we run the function without assigning the output to an object, we see the following: 

```{r diffdf, echo=TRUE}
diffdf::diffdf(base = InitialData , compare = ChangedData)
```

If we assign the output to an object (`diffdf_output`), we can check it's structure:

```{r diffdf_output, echo=TRUE}
# assign to an output
diffdf_output <- diffdf::diffdf(base = InitialData , compare = ChangedData)
str(diffdf_output)
```

This list contains objects named according to `NumDiff` and `VarDiff`. We will test this with the larger master datasets. 


## `testthat::expect_equal()`

This works, but returns the result as an error. 

```{r , error=TRUE}
testthat::expect_equal(object = InitialData, expected = ChangedData)
```

```{r , error=TRUE}
testthat::expect_equal(object = T1Data, expected = T2Data)
```



## vetr::alike()

```{r}
vetr::alike(target = InitialData, current = ChangedData)
```

