---
title: "similar-work"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{similar-work}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
extdata_folder <- "../inst/extdata/"
```

# Motivation 

`dfdiffs` isn't the only way to compare two datasets in R. In this vignette I'll cover other functions and packages that solve similar problems, and how their output differs from what `dfdiffs` returns.

## Packages

First, we'll load `dfdiffs`.

```{r dfdiffs, echo = TRUE}
library(dfdiffs)
```

We'll also need packages for importing and exporting data, iteration, and string wrangling.

```{r more-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(readr)
library(stringr)
library(purrr)
library(glue)
```

The packages below are used for building tables.

```{r tbl-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(labelled)
library(gtsummary)
library(kableExtra)
```


The packages below offer comparison functions similar to those in `dfdiffs`.

```{r similar-pkgs, message=FALSE, warning=FALSE, echo = TRUE}
library(arsenal) # comparedf
library(diffdf)  # diffdf
library(janitor) # compare_df_cols
library(testthat) # expect_equal
library(vetr) # alike
```

## Data 

Before looking at the alternatives, we'll run the `dfdiffs` functions on the example datasets included in the package. These outputs give us a baseline for comparing the other `package::function()` options below.

### New data

To check for new data, we'll use `T1Data` and `T2Data`. The `create_new_data()` function returns the new data (i.e., the rows that are here now but weren't here before).

The `NewData` dataset stores the expected result, so we can check it against the output from `create_new_data()`.

```{r T1Data-T2Data-NewData, echo=TRUE}
T1Data <- dfdiffs::T1Data
T2Data <- dfdiffs::T2Data
T1T2New <- create_new_data(compare = T2Data, base = T1Data)
T1T2New |> 
  knitr::kable(caption = "New data T1 > T2") |> 
  kableExtra::kable_paper()
```

```{r NewData, echo=TRUE}
NewData <- dfdiffs::NewData
NewData |> 
  knitr::kable(caption = "New data (Comparison)") |> 
  kableExtra::kable_paper()
```


Comparing the two tables with `waldo::compare()` shows the only differences are the formats of the date columns.

```{r waldo-compare-T1T2New-NewData}
waldo::compare(x = T1T2New, y = NewData)
```


###  Deleted data 

To test for deleted data, we'll use `CompleteData`, `IncompleteData`, and `DeletedData`.

`CompleteData` is the complete (baseline) dataset.

```{r CompleteData}
CompleteData <- dfdiffs::CompleteData
CompleteData |> 
  knitr::kable(caption = "CompleteData") |> 
  kableExtra::kable_paper()
```


`IncompleteData` is `CompleteData` with some of its rows removed.

```{r IncompleteData}
IncompleteData <- dfdiffs::IncompleteData
IncompleteData |> 
  knitr::kable(caption = "IncompleteData") |> 
  kableExtra::kable_paper()
```


`create_deleted_data()` returns the rows in `CompleteData` that are missing from `IncompleteData`. We'll sort the result by `subject` before printing it.

```{r CompleteData-IncompleteData, echo = TRUE}
IncompCompDiff <- create_deleted_data(
  compare = IncompleteData,
  base = CompleteData)
IncompCompDiff <- IncompCompDiff[order(IncompCompDiff$subject), ]
IncompCompDiff |>
  knitr::kable(caption = "IncompCompDiff") |>
  kableExtra::kable_paper()
```

The output above should match the data stored in `DeletedData`.

```{r DeletedData}
DeletedData <- dfdiffs::DeletedData
DeletedData |> 
  knitr::kable(caption = "DeletedData") |> 
  kableExtra::kable_paper()
```

We can confirm the two tables are identical with `waldo::compare()`.

```{r waldo-compare-IncompCompDiff-DeletedData}
waldo::compare(x = IncompCompDiff, y = DeletedData)
```

### Changed data

To check for changed values between two datasets, we'll pass `InitialData` and `ChangedData` to `create_modified_data()`.

```{r InitialData-ChangedData, echo=TRUE}
InitialData <- dfdiffs::InitialData
ChangedData <- dfdiffs::ChangedData
mods <- create_modified_data(
  compare = ChangedData, base = InitialData)
```

`create_modified_data()` returns a list. The changes by variable are stored in `diffs_byvar`.

```{r diffs_byvar}
mods$diffs_byvar |> 
  knitr::kable(caption = "diffs_byvar") |> 
  kableExtra::kable_paper()
```

The changes by row are stored in `diffs`.

```{r diffs}
mods$diffs |> 
  knitr::kable(caption = "diffs") |> 
  kableExtra::kable_paper()
```


## `diffdf::diffdf()`

Before version 2.1.0, `create_changed_data()` was built directly on `diffdf::diffdf()`, which makes `diffdf` the closest thing to a drop-in alternative. It takes the same two data frames with no extra setup.

```{r diffdf_diffdf, echo=TRUE}
diffdf::diffdf(base = InitialData, compare = ChangedData)
```

`diffdf` is a solid, well-tested package (`dfdiffs` even inherited its numeric tolerance formula; see `?compare_values`). However, its output is a list of small `NumDiff` and `VarDiff_*` tables, one per differing variable, rather than the two tidy tables (`num_diffs` and `var_diffs`) returned by `create_changed_data()`.

## `arsenal::comparedf()`

`create_modified_data()` was originally built the same way, but on `arsenal::comparedf()`. The code below prints the by-variable table from the `comparedf()` summary.

```{r arsenal_comparedf, echo=TRUE}
arsenal_cdf <- arsenal::comparedf(x = InitialData, y = ChangedData)
summary(arsenal_cdf)$diffs.byvar.table
```

Both packages report differences well, but neither provides the row and column headline summary produced by SAS's `PROC COMPARE`. The `compare_summary()` and `compare_columns()` functions in `dfdiffs` fill that gap (see the [proc-compare-parity](proc-compare-parity.html) vignette).

## `janitor::compare_df_cols_same()`

The `compare_df_cols_same()` function from `janitor` compares the columns in two datasets and "*indicates if they will successfully bind together by rows.*"

```{r compare_df_cols_same, echo=TRUE}
compare_df_cols_same(T1Data, T2Data, strict_description = FALSE)
compare_df_cols_same(CompleteData, IncompleteData, strict_description = FALSE)
compare_df_cols_same(InitialData, ChangedData, strict_description = FALSE)
```

All of our test datasets pass this check. It could also serve as a step in one of the `create_` functions to confirm the datasets can be bound together before they're compared.

## `testthat::expect_equal()`

`testthat::expect_equal()` detects the differences, but it reports them by throwing an error. That's the right behavior inside a unit test, but it makes the result harder to work with in an interactive comparison.

```{r expect_equal-initial-changed, error=TRUE}
testthat::expect_equal(object = InitialData, expected = ChangedData)
```

```{r expect_equal-t1data-t2data, error=TRUE}
testthat::expect_equal(object = T1Data, expected = T2Data)
```


## `vetr::alike()`

`vetr::alike()` compares the structure of two objects rather than their values. It returns `TRUE` when `current` is structurally alike `target`, or a character string describing the first mismatch.

```{r}
vetr::alike(target = InitialData, current = ChangedData)
```

