---
title: "tests"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Motivation

`dfdiffs` has two kinds of code to test: the Shiny modules (upload, select, and compare) and the base-R comparison engine. The modules keep their reactive logic in server functions, which are tested with `shiny::testServer()`. `testServer()` runs a module's server function without starting an app or a browser. Each test sets inputs with `session$setInputs()`, then checks the reactive values (or rendered outputs) the module returns.

The comparison engine (`compare_values()`, `diff_values_by_key()`, and the join helpers) is tested directly with plain `test_that()` blocks, so no Shiny session is needed.

The tests live in `tests/testthat/`, and we can run the full suite with `devtools::test()`:

```{r run-tests, eval=FALSE}
devtools::test()
```

There are currently 10 test files containing 48 tests (`test_that()` blocks) and 118 expectations.

# Test inventory

Each row is one `test_that()` block, except the last few rows, which each summarize several tests from one file (the row says how many). The first column gives the file and the test name (the string passed to `test_that()`). The second column names the app feature or function the test protects, and the third describes what the test does and what it expects.

| Test name / file | Feature | Test |
|:-----------------|:--------|:-----|
| `test-mod_upload.R`<br>*mod_upload_server() reads an uploaded base file into base_data()* | Upload module: base file | Uploads `InitialData.csv` as `base_file` (with `base_new_name = "InitialData"`) and expects `base_data()` to be a data frame with 5 rows and a `subject_id` column. |
| `test-mod_upload.R`<br>*mod_upload_server() reads an uploaded compare file into comp_data()* | Upload module: compare file | Uploads `ChangedData.csv` as `comp_file` (with `comp_new_name = "ChangedData"`) and expects `comp_data()` to be a data frame with 5 rows and a `subject_id` column. |
| `test-mod_upload.R`<br>*mod_upload_server() reads an uploaded xlsx sheet into base_data()* | Upload module: xlsx file | Writes `InitialData` to a temp `.xlsx` file with `openxlsx::write.xlsx()`, uploads it as `base_file` with a sheet name, and expects `base_data()` to match `InitialData`'s row count and columns. |
| `test-mod_upload.R`<br>*mod_upload_server() uses base_new_name for base_name() when supplied* | Upload module: dataset naming | Uploads `InitialData.csv` with `base_new_name = "InitialData"` and expects `base_name()` to return `"InitialData"`. |
| `test-mod_select.R`<br>*mod_select_server() builds joined base/compare data from selected columns* | Select module: join column | Passes `InitialData` and `ChangedData` in as the uploaded data, selects every column, and sets `by = c("subject_id", "record")`. Expects `base_join_col_data()` and `comp_join_col_data()` to be data frames with the same number of rows as their source data, containing `join_column`, `join_source`, and `data_source` columns (with `data_source` equal to the dataset names). |
| `test-mod_select.R`<br>*mod_select_server() falls back to a row-by-row comparison with no 'by' columns* | Select module: row-by-row comparison | Selects every column but leaves `by` empty. Expects `base_join_col_data()` to have no `join_column`, but to keep the `data_source` column. |
| `test-mod_compare.R`<br>*mod_compare_server() detects changed values via the 'join_column' path* | Compare module: changed data | Builds joined base/compare data with `create_join_column()` (on `subject_id` and `record`) and passes it in as the selected data. Expects `compare_cols()` to include `join_column`, `changed_data()` to contain `diffs` and `diffs_byvar` (with at least one row in `diffs_byvar`), and `new_data()` and `deleted_data()` to have 0 rows. |
| `test-mod_compare.R`<br>*mod_compare_server() download report works via the 'join_column' path* | Compare module: report download | With joined base/compare data as the selected data, reads the downloaded `.xlsx` file and expects it to contain the "New Data", "Deleted Data", "Changed Data", and "Review Changes" sheets, with the "Review Changes" sheet including a `join_column`. |
| `test-mod_compare.R`<br>*mod_compare_server() download report works via the row-by-row path* | Compare module: report download (no join) | With no join column in the selected data, expects `compare_cols()` to exclude `join_column`, and the downloaded report's "Review Changes" sheet to include a `rownumber` column instead. |
| `test-compare_data.R`<br>*compare_data() returns new, deleted, and changed comparison tables* | `compare_data()` | Runs `compare_data()` on `Roster2021`/`Roster2022` with `by = "subject_id"` and expects a named list of `new_data`, `deleted_data`, `changed_num_diffs`, `changed_var_diffs`, `changed_class_diffs`, `column_diffs`, with `new_data`/`deleted_data` non-empty data frames and `column_diffs` a `dfdiffs_column_diff` object. |
| `test-create_comparison_report.R`<br>*create_comparison_report() writes all nine sheets to the xlsx file* | `create_comparison_report()` | Runs `create_comparison_report()` on `Roster2021`/`Roster2022` and expects the output file to exist with all nine expected sheets, the "New Data" sheet to be non-empty, the "Column Diffs" sheet to mark every shared column "common", and the "Summary" sheet to include `n_base`. |
| `test-launch_app.R`<br>*(four tests, one per launcher)* | App launchers | Calls `launch_app()`, `launch_app_dev()`, `launch_upload_demo()`, and `launch_select_demo()` and expects each to return a `shiny.appobj` without error. |
| `test-join_helpers.R`<br>*(seven tests)* | `select_cols()`, `anti_join_base()`, `left_join_base()` | Covers column subsetting (including a missing-column error), multi-column-key matching, `NA`-in-key handling, and left-join column attachment/`NA` fill for unmatched rows. |
| `test-compare_values.R`<br>*(nineteen tests)* | `compare_values()`, `diff_values_by_key()` | Covers `NA`-vs-`NA` and `NA`-vs-value handling, exact string comparison, numeric tolerance (default and custom `tolerance`/`scale`), integer/double comparison, `Date` vs. same-instant `POSIXct` (equal) and different `POSIXct` (not equal), factor/character label comparison (default and `strict_factor = TRUE`), `ignore_case`/`trim_ws` (including on factor labels), vectorization, and `diff_values_by_key()`'s row-matching, zero-diff, class-mismatch, duplicate-key-warning, and `ignore_case`/`trim_ws` pass-through behavior. |
| `test-compare_columns.R`<br>*(three tests)* | `compare_columns()` | Covers base-only/compare-only/common column reporting, the all-common case, and a `print.dfdiffs_column_diff()` smoke test. |
| `test-compare_summary.R`<br>*(four tests)* | `compare_summary()` | Covers row/column counts against the `Roster2021`/`Roster2022` fixture, base-only/compare-only column counts, value-diff/class-mismatch counts, and a `print.dfdiffs_summary()` smoke test. |

## Test data

The tests use the package's own datasets, so they don't depend on any files outside of `dfdiffs`:

- `InitialData` and `ChangedData` (from `data/`) are the base and compare data for the select and compare tests.
- `Roster2021`/`Roster2022` are the base and compare data for the `compare_data()`, `compare_summary()`, and `create_comparison_report()` tests.
- `inst/extdata/csv/InitialData.csv` and `inst/extdata/csv/ChangedData.csv` are the files uploaded in the CSV upload tests.

# Call structure

Each Shiny module test targets a module's server function, but those server functions call other `dfdiffs` functions along the way. The call trees below show the functions each module test exercises indirectly. For example, the compare module tests also run `create_new_data()`, `create_deleted_data()`, and `create_modified_data()`.

The trees were generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`). Only functions defined in `dfdiffs` are shown.

```{r call-tree-tests, eval=FALSE}
library(stackcallr)
call_tree_dir("R", root = "mod_upload_server")
call_tree_dir("R", root = "mod_select_server")
call_tree_dir("R", root = "mod_compare_server")
```

```
█─mod_upload_server
├─█─upload_data
│ └─load_flat_file
├─base_react_theme
└─comp_react_theme
```

```
█─mod_select_server
├─select_cols
├─base_react_theme
├─comp_react_theme
├─info_react_theme
└─█─create_join_column
  └─select_cols
```

```
█─mod_compare_server
├─compare_columns
├─%nin%
├─info_react_theme
├─█─create_new_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─new_react_theme
├─█─create_deleted_data
│ ├─anti_join_base
│ ├─select_cols
│ ├─rename_join_col
│ └─create_new_column
├─deleted_react_theme
├─█─create_modified_data
│ ├─select_cols
│ ├─█─diff_values_by_key
│ │ ├─compare_values
│ │ └─class_diffs
│ ├─rename_join_col
│ ├─create_new_column
│ └─class_diffs
├─select_cols
├─changed_react_theme
├─left_join_base
└─█─create_comparison_report
  ├─█─create_new_data
  │ ├─anti_join_base
  │ ├─select_cols
  │ ├─rename_join_col
  │ └─create_new_column
  ├─█─create_deleted_data
  │ ├─anti_join_base
  │ ├─select_cols
  │ ├─rename_join_col
  │ └─create_new_column
  ├─█─create_modified_data
  │ ├─select_cols
  │ ├─█─diff_values_by_key
  │ │ ├─compare_values
  │ │ └─class_diffs
  │ ├─rename_join_col
  │ ├─create_new_column
  │ └─class_diffs
  ├─create_empty_tbl
  ├─compare_columns
  ├─class_diffs
  └─█─compare_summary
    └─█─compare_data
      ├─█─create_new_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_deleted_data
      │ ├─anti_join_base
      │ ├─select_cols
      │ ├─rename_join_col
      │ └─create_new_column
      ├─█─create_changed_data
      │ ├─select_cols
      │ ├─█─diff_values_by_key
      │ │ ├─compare_values
      │ │ └─class_diffs
      │ ├─rename_join_col
      │ ├─create_new_column
      │ └─class_diffs
      ├─class_diffs
      └─compare_columns
```

The download-handler tests for `mod_compare_server()` also exercise `create_comparison_report()` (shown as a subtree above), because the module calls it instead of building the report itself.

The remaining test files (`test-compare_data.R`, `test-create_comparison_report.R`, `test-join_helpers.R`, `test-compare_values.R`, `test-compare_columns.R`, and `test-compare_summary.R`) call their target functions directly instead of going through a Shiny module. Their call trees are already covered in the [getting-started](getting-started.html) and [compare-data](compare-data.html) vignettes.

# What isn't tested

The current tests cover the reactive logic of the three modules, the base-R comparison engine, `compare_data()`, `compare_columns()`, `compare_summary()`, and `create_comparison_report()`. They also include a build smoke test for every `launch_*()` entry point. The following areas don't have tests yet:

- The UI functions (`mod_upload_ui()`, `mod_select_ui()`, `mod_compare_ui()`, and `app_ui()`).
- `cross_tabyl()`, and a standalone (non-module) test of `create_join_column()`. It's only exercised indirectly, through the select and compare module tests.
- `class_diffs()` beyond the one genuine-class-mismatch case covered in `test-compare_values.R`.
