---
title: "tests"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Motivation

The Shiny modules in `dfdiffs` (upload, select, compare, and display) keep their reactive logic in server functions. These are tested with `shiny::testServer()`, which runs a module's server function without starting an app or a browser: the test sets inputs with `session$setInputs()`, and then checks the reactive values (or rendered outputs) the module returns.

The tests live in `tests/testthat/`. Run them with:

```{r run-tests, eval=FALSE}
devtools::test()
```

There are currently 4 test files containing 7 tests (`test_that()` blocks) and 26 expectations.

# Test inventory

Each row is one `test_that()` block. The first column gives the file and the test name (the string passed to `test_that()`), the second column is the app feature the test protects, and the third describes what the test does and expects.

| Test name / file | Feature | Test |
|:-----------------|:--------|:-----|
| `test-mod_upload.R`<br>*mod_upload_server() reads an uploaded base file into base_data()* | Upload module: base file | Uploads `InitialData.csv` as `base_file` (with `base_new_name = "InitialData"`) and expects `base_data()` to be a data frame with 5 rows and a `subject_id` column. |
| `test-mod_upload.R`<br>*mod_upload_server() reads an uploaded compare file into comp_data()* | Upload module: compare file | Uploads `ChangedData.csv` as `comp_file` (with `comp_new_name = "ChangedData"`) and expects `comp_data()` to be a data frame with 5 rows and a `subject_id` column. |
| `test-mod_upload.R`<br>*mod_upload_server() uses base_new_name for base_name() when supplied* | Upload module: dataset naming | Uploads `InitialData.csv` with `base_new_name = "InitialData"` and expects `base_name()` to return `"InitialData"`. |
| `test-mod_select.R`<br>*mod_select_server() builds joined base/compare data from selected columns* | Select module: join column | Passes `InitialData` and `ChangedData` in as the uploaded data, selects every column, and sets `by = c("subject_id", "record")`. Expects `base_join_col_data()` and `comp_join_col_data()` to be data frames with the same number of rows as their source data, containing `join_column`, `join_source`, and `data_source` columns (with `data_source` equal to the dataset names). |
| `test-mod_select.R`<br>*mod_select_server() falls back to a row-by-row comparison with no 'by' columns* | Select module: row-by-row comparison | Selects every column but leaves `by` empty. Expects `base_join_col_data()` to have no `join_column`, but to keep the `data_source` column. |
| `test-mod_compare.R`<br>*mod_compare_server() detects changed values via the 'join_column' path* | Compare module: changed data | Builds joined base/compare data with `create_join_column()` (on `subject_id` and `record`) and passes it in as the selected data. Expects `compare_cols()` to include `join_column`, `changed_data()` to contain `diffs` and `diffs_byvar` (with at least one row in `diffs_byvar`), and `new_data()` and `deleted_data()` to have 0 rows. |
| `test-mod_display.R`<br>*mod_display_server() previews base and compare data from mod_upload_server()'s return shape* | Display module: data preview | Passes in a list shaped like the return value of `mod_upload_server()` (`base_data`, `base_name`, `comp_data`, `comp_name`). Expects the name outputs to contain `"InitialData"` and `"ChangedData"`, and both table outputs to be rendered (non-`NULL`). |

## Test data

The tests use the package's own datasets, so they don't depend on files outside of `dfdiffs`:

- `InitialData` and `ChangedData` (from `data/`) are the base and compare data for the select, compare, and display tests.
- `inst/extdata/csv/InitialData.csv` and `inst/extdata/csv/ChangedData.csv` are the files uploaded in the upload tests.

# Call structure

Each test targets a module's server function. The call trees below show the `dfdiffs` functions those tests exercise indirectly (for example, the compare module test also runs `create_new_data()`, `create_deleted_data()`, and `create_modified_data()`). They were generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`); only functions defined in `dfdiffs` are shown.

```{r call-tree-tests, eval=FALSE}
library(stackcallr)
call_tree_dir("R", root = "mod_upload_server")
call_tree_dir("R", root = "mod_select_server")
call_tree_dir("R", root = "mod_compare_server")
call_tree_dir("R", root = "mod_display_server")
```

```
█─mod_upload_server
└─█─upload_data
  └─load_flat_file
```

```
█─mod_select_server
└─create_join_column
```

```
█─mod_compare_server
├─%nin%
├─█─create_new_data
│ ├─rename_join_col
│ └─create_new_column
├─█─create_deleted_data
│ ├─rename_join_col
│ └─create_new_column
├─█─create_modified_data
│ ├─rename_join_col
│ └─create_new_column
└─█─create_changed_data
  ├─extract_df_tables
  ├─rename_join_col
  └─create_new_column
```

```
mod_display_server
```

# What isn't tested

The current tests cover the reactive logic of the four modules. These areas don't have tests yet:

- The UI functions (`mod_upload_ui()`, `mod_select_ui()`, `mod_compare_ui()`, `mod_display_ui()`, and `app_ui()`).
- The app launchers (`launch_app()` and the `launch_*_demo()` functions).
- Uploading Excel (`.xlsx`) files (the upload tests use CSVs).
- The compare module's row-by-row path (no `join_column`), and the report download.
- `compare_data()` and `create_comparison_report()`, which have vignettes but no tests.
