---
title: "upload-data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{upload-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  rmarkdown.html_vignette.check_title = FALSE,
  tibble.print_max = Inf)
```

# Motivation

This vignette walks through the `upload_data()` function, which loads a variety of file types into the `dfdiffs` Shiny application.

```{r pkgs, message=FALSE, warning=FALSE}
library(dfdiffs)
library(shiny)
library(data.table)
library(dplyr)
library(stringr)
library(lubridate)
library(glue)
library(purrr)
library(reactable)
library(haven)
library(readxl)
library(labelled)
library(gtsummary)
```

## External Data 

Test data can be found in the `../inst/extdata/` folder:

```{r extdata-folder, echo=FALSE}
fs::dir_tree("../inst/extdata/", recurse = TRUE)
```


## load_flat_file()

The `load_flat_file()` function imports all forms of flat data files.  

```{r load_flat_file}
load_flat_file <- function(path) {
  ext <- tools::file_ext(path)
  data <- switch(ext,
    txt = data.table::fread(path),
    csv = data.table::fread(path),
    tsv = data.table::fread(path),
    sas7bdat = haven::read_sas(data_file = path),
    sas7bcat = haven::read_sas(data_file = path),
    sav = haven::read_sav(file = path),
    dta = haven::read_dta(file = path)
  )
  return_data <- tibble::as_tibble(data)
  return(return_data)
}
```

The `upload_data()` function below also handles Excel files. If the file is an Excel file, pass the name of the sheet to `sheet`.

```{r upload_data}
upload_data <- function(path, sheet = NULL) {
  ext <- tools::file_ext(path)
  if (ext == "xlsx") {
    raw_data <- readxl::read_excel(
        path = path,
        sheet = sheet
      )
    uploaded <- tibble::as_tibble(raw_data)
  } else {
    uploaded <- load_flat_file(path = path)
  }
  return(uploaded)
}
```


## Call structure

`upload_data()` reads Excel workbooks (`.xlsx`) with `readxl::read_excel()` and hands every other file type to `load_flat_file()`, which picks the reader from the file extension. In the Shiny app, the upload module (`mod_upload_server()`) calls `upload_data()`. The call trees below were generated from the package source with [stackcallr](https://github.com/mjfrigaard/stackcallr) (`pak::pak("mjfrigaard/stackcallr")`); only functions defined in `dfdiffs` are shown.

```{r call-tree-mod_upload_server, eval=FALSE}
stackcallr::call_tree_dir("R", root = "mod_upload_server")
```

```
█─mod_upload_server
└─█─upload_data
  └─load_flat_file
```

The upload demo app (`launch_upload_demo()`) shows the module's UI and server sides together:

```{r call-tree-launch_upload_demo, eval=FALSE}
stackcallr::call_tree_dir("R", root = "launch_upload_demo")
```

```
█─launch_upload_demo
├─dfdiffs_fresh_theme
├─mod_upload_ui
└─█─mod_upload_server
  └─█─upload_data
    └─load_flat_file
```

## 2021 site roster CSVs

```{r lahman_2010_csv_paths}
roster2021_csv_paths <- list.files(path = "../inst/extdata/csv/site-roster/2021", full.names = TRUE, pattern = ".csv$")
head(roster2021_csv_paths)
```

We'll test this on `roster2021_csv_paths[4]` (the full `Roster.csv`).

```{r test-load_flat_file-1}
roster_2021 <- load_flat_file(path = roster2021_csv_paths[4])
glimpse(roster_2021)
```

## 2022 site roster CSVs

```{r lahman_2015_csv_paths}
roster2022_csv_paths <- list.files(path = "../inst/extdata/csv/site-roster/2022", full.names = TRUE, pattern = ".csv$")
head(roster2022_csv_paths)
```

We'll test this on `roster2022_csv_paths[4]`.

```{r test-master_2015}
roster_csv_2022 <- load_flat_file(path = roster2022_csv_paths[4])
glimpse(roster_csv_2022)
```

## List of 2021 CSVs

Now we import every 2021 CSV into a named list, `roster2021_csv_files`:

```{r lahman_2010_csv_files}
roster2021_csv_files <- map(.x = roster2021_csv_paths,
  .f = load_flat_file) %>%
  set_names(x = ., nm = basename(roster2021_csv_paths))
map(roster2021_csv_files, names)
```

We'll also test this with the `map_df()` function.

```{r tbl_2010_csv_files}
tbl_2021_csv_files <- roster2021_csv_paths %>%
  set_names() %>%
  map_df(.x = .,
  .f = load_flat_file, .id = "source") %>%
  mutate(source = basename(source))
tbl_2021_csv_files %>% count(source)
```


## Test on dta 

```{r dta_paths}
dta_paths <- list.files(path = "../inst/extdata/dta", 
  full.names = TRUE, pattern = ".dta$")
tbl_dta_files <- dta_paths %>% 
  set_names() %>% 
  map_df(.x = ., 
  .f = load_flat_file, .id = "source") %>% 
  mutate(source = basename(source))
tbl_dta_files %>% count(source)
```

## Test on sas7bdat 

```{r sas7bdat_paths}
sas7bdat_paths <- list.files(path = "../inst/extdata/sas7bdat", 
  full.names = TRUE, pattern = ".sas7bdat$")
tbl_sas7bdat_files <- sas7bdat_paths %>% 
  set_names() %>% 
  map_df(.x = ., 
  .f = load_flat_file, .id = "source") %>% 
  mutate(source = basename(source))
tbl_sas7bdat_files %>% count(source)
```

## Test on sav 

```{r sav_paths}
sav_paths <- list.files(path = "../inst/extdata/sav", 
  full.names = TRUE, pattern = ".sav$")
tbl_sav_files <- sav_paths %>% 
  set_names() %>% 
  map_df(.x = ., 
  .f = load_flat_file, .id = "source") %>% 
  mutate(source = basename(source))
tbl_sav_files %>% count(source)
```

## Test on tsv

```{r tsv_paths}
tsv_paths <- list.files(path = "../inst/extdata/tsv", 
  full.names = TRUE, pattern = ".tsv$")
tbl_tsv_files <- tsv_paths %>% 
  set_names() %>% 
  map_df(.x = ., 
  .f = load_flat_file, .id = "source") %>% 
  mutate(source = basename(source))
tbl_tsv_files %>% count(source)
```

## Test on txt 

```{r txt_paths}
txt_paths <- list.files(path = "../inst/extdata/txt", 
  full.names = TRUE, pattern = ".txt$")
tbl_txt_files <- txt_paths %>% 
  set_names() %>% 
  map_df(.x = ., 
  .f = load_flat_file, .id = "source") %>% 
  mutate(source = basename(source))
tbl_txt_files %>% count(source)
```
